### 2015-05-02 typeinst/typeinst.tex
cleaned code

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\usepackage{url}
\urldef{\mailsa}\path|{alfred.hofmann, ursula.barth, ingrid.haas, frank.holzwarth,|
\urldef{\mailsb}\path|anna.kramer, leonie.kunz, christine.reiss, nicole.sator,|
\urldef{\mailsc}\path|erika.siebert-cole, peter.strasser, lncs}@springer.com|    
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Lecture Notes in Computer Science:\\Authors' Instructions
for the Preparation\\of Camera-Ready
Contributions\\to LNCS/LNAI/LNBI Proceedings}

% a short form should be given in case it is too long for the running head
\titlerunning{Lecture Notes in Computer Science: Authors' Instructions}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Alfred Hofmann%
\thanks{Please note that the LNCS Editorial assumes that all authors have used
the western naming convention, with given names preceding surnames. This determines
the structure of the names in the running heads and the author index.}%
\and Ursula Barth\and Ingrid Haas\and Frank Holzwarth\and\\
Anna Kramer\and Leonie Kunz\and Christine Rei\ss\and\\
Nicole Sator\and Erika Siebert-Cole\and Peter Stra\ss er}
%
\authorrunning{Lecture Notes in Computer Science: Authors' Instructions}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Springer-Verlag, Computer Science Editorial,\\
Tiergartenstr. 17, 69121 Heidelberg, Germany\\
\mailsa\\
\mailsb\\
\mailsc\\
\url{http://www.springer.com/lncs}}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

\toctitle{Lecture Notes in Computer Science}
\tocauthor{Authors' Instructions}
\maketitle

\begin{abstract}
The abstract should summarize the contents of the paper and should
contain at least 70 and at most 150 words. It should be written using the
\emph{abstract} environment.
\keywords{We would like to encourage you to list your keywords within
the abstract section}
\end{abstract}

\section{Introduction}

Metagenomics is a direct sequencing of DNA from all organisms in an environment without culturing.  It is a culture-independent tool for studying environmental microorganisms. In addition to the information about taxonomic diversity (‘who is there’), metagenomics gives insight into the physiology of the organisms present in the environment (‘what are they doing’), through studying their genes. Two main goals to be achieved with metagenomics are finding new genes with desired biological activity (bioprospecting) and studying environmental microbes without the need to culture them.

Metagenomic data are considerably more complex. The assembly of metagenomic data is complicated due to several reasons. Few of them are: (i) widely different levels of representation for different organisms in a community; (ii) genomic variation between closely related organisms; (iii) conserved genomic regions shared by distantly related organims; and (iv) repetitive sequences within individual genomes. Genomic repeats are the major challenge when assembling isolated genomes and their effect is prominently seen in metagenomic datasets. It has been show in (cite kingsford) that the number of reconstructions grow exponentially with the number of repeats. Therefore it is impossible find one correct solution for assembly. There are several approaches such as (cite bambus2) to find repeats in metagenomic data. The main drawback of these methods is that they do not scale well as number of nodes in assembly graph increase and prove very inefficient as number of nodes reaches to about a million. To tackle this problem, we explore some of the work in social network analysis, particularly in the area of node centrality. We make use of betweenness centrality to find out repeats in assembly graph.

\section{Related Work}
In graph theory and network analysis, indicators of centrality identify the most important vertices within a graph.Several metrics to measure centrality have been proposed. We make use of betweenness centrality. For a particular node, betweenness centrality is equal to the number of shortest paths from all vertices to all others that pass through that node. Formally, for a node $v$, it is defined by following expression:

$$g(v) = \sum_{s \neq t \neq v} \frac{\sigma_{st}(v)}{\sigma_{st}}$$

where $\sigma_{st}$ is the total number of shortest paths from node $s$ to node $t$ and $\sigma_{st}(v)$ is the number of those paths passing through $v$.

Brandes(cite brandes) gives an exact algorithm for computing betweenness centrality of all the nodes that is based on solving a single source shortest path problem for each node. This algorithm takes time $\Theta(mn)$ for unit edge weight graphs and $\Theta(mn + n^{2}log(n))$ for weighted graphs. This algorithm does not scale well with large networks with millions of nodes and edges. Bader and Madduri (cite) provide a massively parallel implementation of the exact algorithm. However this approach scales to about a few million nodes.

Due to scalability limitations of exact betweenness centrality, several approximation algorithms for betweenness centrality have been proposed. Bader and Pich(cite) provide an approximation algorithm by extrapolating from a subset of $k$ starting nodes called pivots and using same aggregation strategy as exact betweeness algorithm.

Springer provides you with a complete integrated \LaTeX{} document class
(\texttt{llncs.cls}) for multi-author books such as those in the LNCS
series. Papers not complying with the LNCS style will be reformatted.
This can lead to an increase in the overall number of pages. We would
therefore urge you not to squash your paper.

Please always cancel any superfluous definitions that are
not actually used in your text. If you do not, these may conflict with
the definitions of the macro package, causing changes in the structure
of the text and leading to numerous mistakes in the proofs.

If you wonder what \LaTeX{} is and where it can be obtained, see the
``\textit{LaTeX project site}'' (\url{http://www.latex-project.org})
and especially the webpage ``\textit{How to get it}''
(\url{http://www.latex-project.org/ftp.html}) respectively.

When you use \LaTeX\ together with our document class file,
\texttt{llncs.cls},
your text is typeset automatically in Computer Modern Roman (CM) fonts.
Please do
\emph{not} change the preset fonts. If you have to use fonts other
than the preset fonts, kindly submit these with your files.

Please use the commands \verb+\label+ and \verb+\ref+ for
cross-references and the commands \verb+\bibitem+ and \verb+\cite+ for
references to the bibliography, to enable us to create hyperlinks at
these places.

For preparing your figures electronically and integrating them into
your source file we recommend using the standard \LaTeX{} \verb+graphics+ or
\verb+graphicx+ package. These provide the \verb+\includegraphics+ command.
In general, please refrain from using the \verb+\special+ command.

Remember to submit any further style files and
fonts you have used together with your source files.

\subsubsection{Headings.}

Headings should be capitalized
(i.e., nouns, verbs, and all other words
except articles, prepositions, and conjunctions should be set with an
initial capital) and should,
with the exception of the title, be aligned to the left.
Words joined by a hyphen are subject to a special rule. If the first
word can stand alone, the second word should be capitalized.

Here are some examples of headings: ``Criteria to Disprove
Context-Freeness of Collage Language", ``On Correcting the Intrusion of
Tracing Non-deterministic Programs by Software", ``A User-Friendly and
Extendable Data Distribution System", ``Multi-flip Networks:
Parallelizing GenSAT", ``Self-determinations of Man".

\subsubsection{Lemmas, Propositions, and Theorems.}

The numbers accorded to lemmas, propositions, and theorems, etc. should
appear in consecutive order, starting with Lemma 1, and not, for
example, with Lemma 11.

\subsection{Figures}

For \LaTeX\ users, we recommend using the \emph{graphics} or \emph{graphicx}
package and the \verb+\includegraphics+ command.

Please check that the lines in line drawings are not
interrupted and are of a constant width. Grids and details within the
figures must be clearly legible and may not be written one on top of
the other. Line drawings should have a resolution of at least 800 dpi
(preferably 1200 dpi). The lettering in figures should have a height of
2~mm (10-point type). Figures should be numbered and should have a
caption which should always be positioned \emph{under} the figures, in
contrast to the caption belonging to a table, which should always appear
\emph{above} the table; this is simply achieved as matter of sequence in
your source.

Please center the figures or your tabular material by using the \verb+\centering+
declaration. Short captions are centered by default between the margins
and typeset in 9-point type (Fig.~\ref{fig:example} shows an example).
The distance between text and figure is preset to be about 8~mm, the
distance between figure and caption about 6~mm.

To ensure that the reproduction of your illustrations is of a reasonable
quality, we advise against the use of shading. The contrast should be as
pronounced as possible.

If screenshots are necessary, please make sure that you are happy with
the print quality before you send the files.
\begin{figure}
\centering
\includegraphics[height=6.2cm]{eijkel2}
\caption{One kernel at $x_s$ (\emph{dotted kernel}) or two kernels at
$x_i$ and $x_j$ (\textit{left and right}) lead to the same summed estimate
at $x_s$. This shows a figure consisting of different types of
lines. Elements of the figure described in the caption should be set in
italics, in parentheses, as shown in this sample caption.}
\label{fig:example}
\end{figure}

Please define figures (and tables) as floating objects. Please avoid
using optional location parameters like ``\verb+[h]+" for ``here".

\paragraph{Remark 1.}

In the printed volumes, illustrations are generally black and white
(halftones), and only in exceptional cases, and if the author is
prepared to cover the extra cost for color reproduction, are colored
pictures accepted. Colored pictures are welcome in the electronic
version free of charge. If you send colored figures that are to be
printed in black and white, please make sure that they really are
legible in black and white. Some colors as well as the contrast of
converted colors show up very poorly when printed in black and white.

\subsection{Formulas}

Displayed equations or formulas are centered and set on a separate
line (with an extra line or halfline space above and below). Displayed
expressions should be numbered for reference. The numbers should be
consecutive within each section or within the contribution,
with numbers enclosed in parentheses and set on the right margin --
which is the default if you use the \emph{equation} environment, e.g.,
\begin{equation}
  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt \;  .
\end{equation}

Equations should be punctuated in the same way as ordinary
text but with a small space before the end punctuation mark.

\subsection{Footnotes}

The superscript numeral used to refer to a footnote appears in the text
either directly after the word to be discussed or -- in relation to a
phrase or a sentence -- following the punctuation sign (comma,
semicolon, or period). Footnotes should appear at the bottom of
the
normal text area, with a line of about 2~cm set
immediately above them.\footnote{The footnote numeral is set flush left
and the text follows with the usual word spacing.}

\subsection{Program Code}

Program listings or program commands in the text are normally set in
typewriter font, e.g., CMTT10 or Courier.

\medskip

\noindent
{\it Example of a Computer Program}
\begin{verbatim}
program Inflation (Output)
  {Assuming annual inflation rates of 7%, 8%, and 10%,...
   years};
   const
     MaxYears = 10;
   var
     Year: 0..MaxYears;
     Factor1, Factor2, Factor3: Real;
   begin
     Year := 0;
     Factor1 := 1.0; Factor2 := 1.0; Factor3 := 1.0;
     WriteLn('Year  7% 8% 10%'); WriteLn;
     repeat
       Year := Year + 1;
       Factor1 := Factor1 * 1.07;
       Factor2 := Factor2 * 1.08;
       Factor3 := Factor3 * 1.10;
       WriteLn(Year:5,Factor1:7:3,Factor2:7:3,Factor3:7:3)
     until Year = MaxYears
end.
\end{verbatim}
%
\noindent
{\small (Example from Jensen K., Wirth N. (1991) Pascal user manual and
report. Springer, New York)}

\subsection{Citations}

For citations in the text please use
square brackets and consecutive numbers: \cite{jour}, \cite{lncschap},
\cite{proceeding1} -- provided automatically
by \LaTeX 's \verb|\cite| \dots\verb|\bibitem| mechanism.

\subsection{Page Numbering and Running Heads}

There is no need to include page numbers. If your paper title is too
long to serve as a running head, it will be shortened. Your suggestion
as to how to shorten it would be most welcome.

\section{LNCS Online}

The online version of the volume will be available in LNCS Online.
Members of institutes subscribing to the Lecture Notes in Computer
Science series have access to all the pdfs of all the online
publications. Non-subscribers can only read as far as the abstracts. If
they try to go beyond this point, they are automatically asked, whether
they would like to order the pdf, and are given instructions as to how
to do so.

Please note that, if your email address is given in your paper,
it will also be included in the meta data of the online version.

\section{BibTeX Entries}

The correct BibTeX entries for the Lecture Notes in Computer Science
volumes can be found at the following Website shortly after the
publication of the book:
\url{http://www.informatik.uni-trier.de/~ley/db/journals/lncs.html}

\subsubsection*{Acknowledgments.} The heading should be treated as a
subsubsection heading and should not be assigned a number.

\section{The References Section}\label{references}

In order to permit cross referencing within LNCS-Online, and eventually
between different publishers and their online databases, LNCS will,
from now on, be standardizing the format of the references. This new
feature will increase the visibility of publications and facilitate
academic research considerably. Please base your references on the
examples below. References that don't adhere to this style will be
reformatted by Springer. You should therefore check your references
thoroughly when you receive the final pdf of your paper.
The reference section must be complete. You may not omit references.
Instructions as to where to find a fuller version of the references are
not permissible.

We only accept references written using the latin alphabet. If the title
of the book you are referring to is in Russian or Chinese, then please write
(in Russian) or (in Chinese) at the end of the transcript or translation
of the title.

The following section shows a sample reference list with entries for
journal articles \cite{jour}, an LNCS chapter \cite{lncschap}, a book
\cite{book}, proceedings without editors \cite{proceeding1} and
\cite{proceeding2}, as well as a URL \cite{url}.
Please note that proceedings published in LNCS are not cited with their
full titles, but with their acronyms!

\begin{thebibliography}{4}

\bibitem{jour} Smith, T.F., Waterman, M.S.: Identification of Common Molecular
Subsequences. J. Mol. Biol. 147, 195--197 (1981)

\bibitem{lncschap} May, P., Ehrlich, H.C., Steinke, T.: ZIB Structure Prediction Pipeline:
Composing a Complex Biological Workflow through Web Services. In: Nagel,
W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006. LNCS, vol. 4128,
pp. 1148--1158. Springer, Heidelberg (2006)

\bibitem{book} Foster, I., Kesselman, C.: The Grid: Blueprint for a New Computing
Infrastructure. Morgan Kaufmann, San Francisco (1999)

\bibitem{proceeding1} Czajkowski, K., Fitzgerald, S., Foster, I., Kesselman, C.: Grid
Information Services for Distributed Resource Sharing. In: 10th IEEE
International Symposium on High Performance Distributed Computing, pp.
181--184. IEEE Press, New York (2001)

\bibitem{proceeding2} Foster, I., Kesselman, C., Nick, J., Tuecke, S.: The Physiology of the
Grid: an Open Grid Services Architecture for Distributed Systems
Integration. Technical report, Global Grid Forum (2002)

\bibitem{url} National Center for Biotechnology Information, \url{http://www.ncbi.nlm.nih.gov}

\end{thebibliography}

\section*{Appendix: Springer-Author Discount}

LNCS authors are entitled to a 33.3\% discount off all Springer
publications. Before placing an order, the author should send an email, 
giving full details of his or her Springer publication,
to \url{orders-HD-individuals@springer.com} to obtain a so-called token. This token is a
number, which must be entered when placing an order via the Internet, in
order to obtain the discount.

\section{Checklist of Items to be Sent to Volume Editors}
Here is a checklist of everything the volume editor requires from you:

\begin{itemize}
\settowidth{\leftmargin}{{\Large$\square$}}\advance\leftmargin\labelsep
\itemsep8pt\relax
\renewcommand\labelitemi{{\lower1.5pt\hbox{\Large$\square$}}}

\item The final \LaTeX{} source files
\item A final PDF file
\item A copyright form, signed by one author on behalf of all of the
authors of the paper.
\item A readme giving the name and email address of the
corresponding author.
\end{itemize}
\end{document}

### 2015-05-02 sampling_algo.cpp

@@ -41,56 +41,44 @@ int main(int argc, char* argv[]) {
 	map<string,string> labelmap;
 	string id;
 	string label;
-	int cnt = 1;
 	while (std::getline(ifile, str)) {
-		if(cnt %4 == 0){
+		if(str.find("id") != string::npos){
 			id = str.substr(7);
 		}
-		if((cnt-1)%4 == 0 && cnt != 1) {
+		if(str.find("label") != string::npos) {
 			label = str.substr(10);
 			labelmap[id] = label;
 		}
-		cnt += 1;
 	}
 
-	cout << "Nodes: " << g.nodes().size() << endl;
-	cout << "Edges: " << g.edges().size() << endl;
+	cerr << "Nodes: " << g.nodes().size() << endl;
+	cerr << "Edges: " << g.edges().size() << endl;
 	double delta = 0.1;
 	double epsilon = 0.1;
-	//Betweenness centrality(g);
-	//ApproxBetweenness centrality(g, delta, epsilon, 2);
+	
 	ApproxBetweenness2 centrality(g,sampleNodes,false);
 	centrality.run();
 	ofstream myfile;
 	filepath = filepath.substr(0,filepath.length()-4);
-	myfile.open(filepath+".repeats");
 	vector<pair<node, double>> nodeRanks(g.nodes().size());
 	nodeRanks = centrality.ranking();
 	vector<double> scores = centrality.scores();
-	vector<double> bc_nonzero;
-	bc_nonzero = scores;
-//	for(int i = 0;i < scores.size();i++){
-//		if(scores.at(i) != 0){
-//			bc_nonzero.push_back(scores.at(i));
-//		}
-//	}
-	cout<<bc_nonzero.size()<<endl;
-	double sum = std::accumulate(bc_nonzero.begin(), bc_nonzero.end(), 0.0);
-	double mean = sum / bc_nonzero.size();
+	double sum = std::accumulate(scores.begin(), scores.end(), 0.0);
+	double mean = sum / scores.size();
 
-	double sq_sum = std::inner_product(bc_nonzero.begin(), bc_nonzero.end(), bc_nonzero.begin(), 0.0);
-	double stdev = std::sqrt(sq_sum / bc_nonzero.size() - mean * mean);
-	cout<<"Mean: "<<mean<<endl;
-	cout<<"Std Dev: "<<stdev<<endl;
-	cout<<"bound: "<<mean + filterparam*stdev<<endl;
[…]

### 2015-05-02 vc_algo.cpp

@@ -39,56 +39,45 @@ int main(int argc, char* argv[]) {
 	map<string,string> labelmap;
 	string id;
 	string label;
-	int cnt = 1;
 	while (std::getline(ifile, str)) {
-		if(cnt %4 == 0){
+		if(str.find("id") != string::npos){
 			id = str.substr(7);
 		}
-		if((cnt-1)%4 == 0 && cnt != 1) {
+		if(str.find("label") != string::npos) {
 			label = str.substr(10);
 			labelmap[id] = label;
 		}
-		cnt += 1;
 	}
 
-	cout << "Nodes: " << g.nodes().size() << endl;
-	cout << "Edges: " << g.edges().size() << endl;
+	cerr << "Nodes: " << g.nodes().size() << endl;
+	cerr << "Edges: " << g.edges().size() << endl;
 	double delta = 0.1;
 	double epsilon = 0.1;
-	//Betweenness centrality(g);
 	ApproxBetweenness centrality(g, delta, epsilon, 2);
-	//ApproxBetweenness2 centrality(g,100,false);
 	centrality.run();
 	ofstream myfile;
 	filepath = filepath.substr(0,filepath.length()-4);
-	myfile.open(filepath+".repeats");
 	vector<pair<node, double>> nodeRanks(g.nodes().size());
 	nodeRanks = centrality.ranking();
 	vector<double> scores = centrality.scores();
-	vector<double> bc_nonzero;
-	bc_nonzero = scores;
-//	for(int i = 0;i < scores.size();i++){
-//		if(scores.at(i) != 0){
-//			bc_nonzero.push_back(scores.at(i));
-//		}
-//	}
-	cout<<bc_nonzero.size()<<endl;
-	double sum = std::accumulate(bc_nonzero.begin(), bc_nonzero.end(), 0.0);
-	double mean = sum / bc_nonzero.size();
 
-	double sq_sum = std::inner_product(bc_nonzero.begin(), bc_nonzero.end(), bc_nonzero.begin(), 0.0);
-	double stdev = std::sqrt(sq_sum / bc_nonzero.size() - mean * mean);
-	cout<<"Mean: "<<mean<<endl;
-	cout<<"Std Dev: "<<stdev<<endl;
-	cout<<"bound: "<<mean + filterparam*stdev<<endl;
+	double sum = std::accumulate(scores.begin(), scores.end(), 0.0);
+	double mean = sum / scores.size();
+
[…]

### 2015-05-03 manuscript/manuscript.tex
report updated

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\usepackage{amsmath}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\newtheorem{mydef}{Definition}
\newtheorem{mylemma}{Lemma}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

\usepackage{url}
\urldef{\mailsa}\path|{alfred.hofmann, ursula.barth, ingrid.haas, frank.holzwarth,|
\urldef{\mailsb}\path|anna.kramer, leonie.kunz, christine.reiss, nicole.sator,|
\urldef{\mailsc}\path|erika.siebert-cole, peter.strasser, lncs}@springer.com|    
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Lecture Notes in Computer Science:\\Authors' Instructions
for the Preparation\\of Camera-Ready
Contributions\\to LNCS/LNAI/LNBI Proceedings}

% a short form should be given in case it is too long for the running head
\titlerunning{Lecture Notes in Computer Science: Authors' Instructions}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Alfred Hofmann%
\thanks{Please note that the LNCS Editorial assumes that all authors have used
the western naming convention, with given names preceding surnames. This determines
the structure of the names in the running heads and the author index.}%
\and Ursula Barth\and Ingrid Haas\and Frank Holzwarth\and\\
Anna Kramer\and Leonie Kunz\and Christine Rei\ss\and\\
Nicole Sator\and Erika Siebert-Cole\and Peter Stra\ss er}
%
\authorrunning{Lecture Notes in Computer Science: Authors' Instructions}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Springer-Verlag, Computer Science Editorial,\\
Tiergartenstr. 17, 69121 Heidelberg, Germany\\
\mailsa\\
\mailsb\\
\mailsc\\
\url{http://www.springer.com/lncs}}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

\toctitle{Lecture Notes in Computer Science}
\tocauthor{Authors' Instructions}
\maketitle

\begin{abstract}
The abstract should summarize the contents of the paper and should
contain at least 70 and at most 150 words. It should be written using the
\emph{abstract} environment.
\keywords{We would like to encourage you to list your keywords within
the abstract section}
\end{abstract}

\section{Introduction}

Metagenomics is a direct sequencing of DNA from all organisms in an environment without culturing.  It is a culture-independent tool for studying environmental microorganisms. In addition to the information about taxonomic diversity (‘who is there’), metagenomics gives insight into the physiology of the organisms present in the environment (‘what are they doing’), through studying their genes. Two main goals to be achieved with metagenomics are finding new genes with desired biological activity (bioprospecting) and studying environmental microbes without the need to culture them.

Metagenomic data are considerably more complex. The assembly of metagenomic data is complicated due to several reasons. Few of them are: (i) widely different levels of representation for different organisms in a community; (ii) genomic variation between closely related organisms; (iii) conserved genomic regions shared by distantly related organims; and (iv) repetitive sequences within individual genomes. Genomic repeats are the major challenge when assembling isolated genomes and their effect is prominently seen in metagenomic datasets. It has been show by Kingsford et al.\cite{kingsford} that the number of reconstructions grow exponentially with the number of repeats. Therefore it is impossible find one correct solution for assembly. There are several approaches such as \cite{bambus} to find repeats in metagenomic data. The main drawback of these methods is that they do not scale well as number of nodes in assembly graph increase and prove very inefficient as number of nodes reaches to about a million. To tackle this problem, we explore some of the work in social network analysis, particularly in the area of node centrality. We make use of betweenness centrality to find out repeats in assembly graph.

\section{Related Work}
In graph theory and network analysis, indicators of centrality identify the most important vertices within a graph.Several metrics to measure centrality have been proposed. We make use of betweenness centrality. For a particular node, betweenness centrality is equal to the number of shortest paths from all vertices to all others that pass through that node. Formally, for a node $v$, it is defined by following expression:

$$g(v) = \sum_{s \neq t \neq v} \frac{\sigma_{st}(v)}{\sigma_{st}}$$

where $\sigma_{st}$ is the total number of shortest paths from node $s$ to node $t$ and $\sigma_{st}(v)$ is the number of those paths passing through $v$.

Brandes\cite{brandes} gives an exact algorithm for computing betweenness centrality of all the nodes that is based on solving a single source shortest path problem for each node. An SSSP computation from $s$ produces a directed acyclic graph (DAG) encoding all shortest paths starting at $s$. By backward aggregation of counter values, the contributions of these paths to the betweenness counters can be computed in linear time. Brandes algorithm explains how to calculate $\sigma_{st}$ on the fly during shortest path calculations: $\sigma_{ss} = 1$ and for $s \neq t, \sigma_{st} = \sum_{v \in pred(t)} \sigma_{sv}$ where $pred(t)$ is a multiset containing the immediate predecessors of  $t$ in the shortest path DAG. In a subsequent aggregation phase, the nodes are processed in reverse topological order, i.e. by non-decreasing distance from $s$. So we have: 
$$\delta_{s}(v) = \sum_{w \in succ(v)} \frac{\sigma_{sv}}{\sigma_{sw}}(1+\delta_{s}(w))$$

where $\delta_{s}(v) = \frac{\sigma_{st}(v)}{\sigma_{st}}$ and $succ(v)$ is an immediate successor of $v$  in the shortest path DAG.   
This algorithm takes time $\Theta(mn)$ for unit edge weight graphs and $\Theta(mn + n^{2}log(n))$ for weighted graphs. This algorithm does not scale well with large networks with millions of nodes and edges. Bader and Madduri\cite{bader} provide a massively parallel implementation of the exact algorithm. However this approach scales to about a few million nodes.

Due to scalability limitations of exact betweenness centrality, several approximation algorithms for betweenness centrality have been proposed. Bader and Pich\cite{bp} provide an approximation algorithm by extrapolating from a subset of $k$ starting nodes called pivots and using same aggregation strategy as exact betweeness algorithm. However, this algorithm overestimates the centrality of some unimportant nodes which lie near pivot. The approximation algorithm proposed in \cite{sanders} comes over this drawback by changing the scheme for aggregating betweenness contributions so that nodes do not get profited just because they are near pivot. They define an estimator for betweenness centrality. This estimator is characterized by a length function $l : E \Rightarrow \mathbb{R}$ on the edges and a scaling function $f : [0,1] \Rightarrow [0,1]$. For a path $P = (e_{1},e_{2}...,e_{k})$ let $l(P) := \sum_{1 \leq i \leq k} l(e_{i})$. Let $n$ be the number of sample nodes chosen from a graph. The algorithm works as follows: In each iteration one of $2n$ shortest path searches is performed with uniform probability $1/2n$. An iteration can be a forward search in $G = (V,E)$ from a pivot $s \in V$ ($|V| = n$) or backward search from a pivot $t \in V$, that is a search from $t$ in $(V,{(v,u) : (u,v) \in E})$. For each shortest path of the form 
$$ P = (\overbrace{s,...,v}^{Q},..,t)$$ 
a scaled contribution is defined as follows:\\

\begin{equation*}
	\delta_{P}(v) := \begin{cases}
					\frac{f(l(Q)/l(P))}{\sigma_{st}} & \text{for a forward search} \\\\
					\frac{1 - f(l(Q)/l(P))}{\sigma_{st}} & \text{for a backward search} 
					\end{cases}
\end{equation*}

Overall, $v$ receives contribution

$$\delta(v) := \delta_{s}(v) := \sum_{t \in V}\sum \{\delta_{P}(v):  P \in SP_{st}(v)\}$$
due to a forward search, and

$$\delta(v) := \delta_{t}(v) := \sum_{s \in V}\sum \{\delta_{P}(v):  P \in SP_{st}(v)\}$$
due to a backward search.

Geisberger et al. \cite{sanders} have shown that $X := 2n\delta(v)$ is an unbiased estimator of centrality of v that is $E(X) := g(v)$. Therefore by averaging $k$ independent runs of this estimator, we get an approximation $\frac{X_{1}+X_{2}+..+X_{k}}{k}$ of the betweenness centrality for node $v$.

To compute the contributions of nodes to centrality, a modified scheme of Brandes algorithm is used. Contributions are calculated as follows:

$$\delta_{s}(v) = \sum_{w \in succ(v)} \frac{\mu(s,v)}{\mu(s,w)}.\frac{\sigma_{sv}}{\sigma_{sw}}(1+\delta_{s}(w))$$
where $\mu(s,w)$ is the shortest path distance from $s$ to $w$. It uses $\frac{\mu(s,v)}{\mu(s,w)}$ in aggregation instead of $1$ in Brandes algorithm.

There  is another approximation algorithm was proposed by Matteo et al. \cite{matteo} based on randomized sampling of shortest paths and offer probabilistic guarantees on the quality of approximation. This algorithm guarantees that all approximate values of betweenness for all vertices are within an additive factor $\epsilon \in (0,1)$ from the real values with probability at least $1-\delta$. To derive bounds on sample size, results from VC dimension\cite{vc} theory are used. The main advantage of this algorithm is that the sample size does not depend on the size of the graph, but only on the maximum number vertices in a shortest path which is called \textit{vertex diameter}.

The Vapnik-Chernovenkis(VC) dimensions of a class of subset is defined on a set of points is a measure of expressiveness of such class. Given a probability distribution on the set of points, a finite bound on the VC-dimension of the class of subsets implies a bound on the number of random samples required to approximate the probability of each subset in the class with its empirical average. To define this formally, let $D$ be a domain and $R$ be a collection of subsets from $D$. We call $R$ a \textit{range set} on $D$. Given $B \subseteq D$, the projection of $R$ on $B$ is the set of $P_{R}(B) = \{B \cap A : A \in R\}$.We say that the set $B$ is \textit{shattered} by $R$ if $P_{R}(B) = 2^{B}$, where $2^{B}$ is the powerset of $B$.

\begin{mydef}
The \textit{Vapnik-Chervonenkis} (VC) dimension of $R$, denoted as $VC(R)$, is the cardinality of the largest subset of $D$ that is shattered by $R$.
\end{mydef}

It has been shown that if an upper bound to the VC-dimension of $R$ is available, it is possible to build an $\epsilon$-approximation by sampling the points of the domain according to distribution $\phi$.

\begin{mydef}
Let $R$ be a range set on a domain $D$ with VC(R) $\leq d$, and let $\phi$ be a distribution on $D$. Given $\epsilon, \delta \in (0,1)$ let $S$ be a collection of $|S|$ points from $D$ sampled according to $\phi$ with
\begin{equation} \label{sample_size}
|S| = \frac{c}{\epsilon^{2}}(d + ln\frac{1}{\delta})
\end{equation}

where $c$ is a universal positive constant approximated to 0.5. Then $S$ is an $\epsilon$-approximation to $(R,\phi)$ with probability at least $(1-\delta)$.
\end{mydef}

Using these insights from VC-dimensions, we can define a range set on the shortest paths of a graph $G = (V,E)$ and present a tight upper bound to its VC-dimensions. The range set $R_{G}$ is defined on the set $\mathtt{S}_{G}$ of all shortest paths between vertices of $G$. It contains, for each vertex $v \in V$, the set $T_{v}$ of shortest paths that $v$ is internal to: 
$$R_{G} = \{T_{v} : v \in V\}.$$
Let us define for each shortest path $p$, $Int(p)$ be the set of internal vertices on that path.  Let $H(G)$ be the maximum integer $d$ such that there are $h = \left\lfloor{log_{2}(d-2)}\right\rfloor + 1$ shortest paths $\{p_{1},p_{2},....p_{h}\}$
in $\mathtt{S}_{G}$ of size at least $d$ and such that the collection $\{Int(p_{1}),....,Int(p_{h})\}$ is an anti-chain, that is a collection of sets such that for no two sets in it, one is a subset of other. It has been shown that this can be used to provide an upper bound on VC-dimension on set $R_{G}$. So we have:
$$VC(R_{G}) \leq \left\lfloor{log_{2}(H(G) - 2)}\right\rfloor + 1$$
It $VD(G)$ be the vertex diameter of graph G. It can be shown that $H(G) \leq VD(G)$. As calculating exact $H(G)$ value is hard, we can make use of $VD(G)$ to define upper bound on VC-dimension of $R_{G}$ as follows:
$$VC(R_{G}) \leq \left\lfloor{log_{2}(VD(G) - 2)}\right\rfloor + 1$$

The algorithm to calculate approximate betweenness centrality works as follows. For a graph $G(V,E)$, with vertex diameter $VD(G)$, sample size $r$ is estimated as follows:
$$ d = \left\lfloor{log_{2}(VD(G) - 2)}\right\rfloor + 1$$
Substituting this $d$ in equation\ref{sample_size}, we get
\begin{equation} \label{sample_size_algo}
r = \frac{c}{\epsilon^{2}}(\left\lfloor{log_{2}(VD(G) - 2)}\right\rfloor + 1 + ln\frac{1}{\delta})
\end{equation}
This sample size achieves desired accuracy with the desired confidence. The algorithm repeats following steps $r$ times: 1. It samples a pair of $u,v$ of distinct vertices uniformly at random. 2. It computes set $S_{uv}$ or all shortest paths between $u$ and $v$. 3. It selects path $p$ from $S_{uv}$ at random. 4. It increases by $1/r$ the betweenness estimation of each vertex in $Int(p)$. If sampled vertices $u$ and $v$ are not connected, steps 3 and 4 can be skipped.

There are two crucial things in this algorithm. First is computing approximate vertex diameter. Calculating exact VD(G) will need to solve all pair shortest  path problem. The complexity of this algorithm is $O(V^{2}logV + VE)$ by Johnson's algorithm(cite) and $\Theta(V^{3})$ by Floyd-Warshall algorithm. This would be very slow. A faster linear time 2-approximation algorithm is proposed in \cite{matteo}. It is as follows:
\begin{itemize}
\item Select $v \in V$ at random
\item Compute shortest path from $v$ to all vertices in $V$
\item Finally take $VD(G)$ be the sum of lengths of two shortest paths with maximum size from $v$ to two distinct nodes $u$ and $w$. 
\end{itemize}

Second is to sample shortest paths. They way a random shortest path from $S_{uv}$ is inspired by Brandes algorithm(cite). Assume that $u$ and $v$ are connected. For each vertex $w$, let $P_{u}(w)$ be a subset of neighbors of $w$ which are predecessors of $w$ along the shortest path from $u$ to $w$. Let $p^{*}$ be the sampled shortest path that we build backwards starting from the endpoint $v$ and by adding sampled predecessors before $v$. Initially $p^{*} = {v}$. Starting from $v$, we select it's predecessor $z \in P_{u}(v)$ using weighted random sampling. Each $z \in P_{u}(v)$ has probability $\sigma_{uz}/\sum_{w \in P_{u}(v)}\sigma_{uv} = \sigma_{uz}/\sigma_{uv}$. We add $z$ to $p^{*}$ and repeat this procedure until $z \neq u$. In the end, we have single shortest path sampled from $S_{uv}$.

\begin{algorithm}
\caption{Algorithm to compute approximate betweenness centrality}
\begin {algorithmic}[1]
\State {\textbf{Input}:Graph $G(V,E)$ with $|V| = n,\epsilon,\delta \in (0,1)$}
\State {\textbf{Output}: A set of approximate betweenness centrality for all nodes $v$ in $G$}
\For {$v \in V$}
\State {$b'(v) = 0$}
\EndFor 
\State {$VD(G) \leftarrow \mathtt{getVertexDiameter}(G)$}
\State {$r \leftarrow \frac{c}{\epsilon^{2}}(\left\lfloor{log_{2}(VD(G) - 2)}\right\rfloor + 1 + ln\frac{1}{\delta})$}
\For{$i \leftarrow 1$ to $r$}
\State {$(u,v) = \mathtt{SampleUniformVertexPair(V)}$}
\State {$S_{uv} = \mathtt{ComputeAllShortestPath(u,v)}$}
\If {$S_{uv} \neq \{p_{\phi}\}$}
\State 	{$t \leftarrow v$}
\While {$t \neq u$}
\State {sample $z \in P_{u}(t)$ with probability $\sigma_{uz}/\sigma_{ut}$}
\If {$z \neq u$}
\State {$b'(z) \leftarrow b'(z) + 1/r$}
\EndIf
\State {$t \leftarrow z$}
\EndWhile
\EndIf
\EndFor\\
\Return {$(v,b'(v)) :\forall v \in V$}
\end{algorithmic}
\end{algorithm}

Analysis of this algorithm shows that it offers probabilistic guarantees on the quality of all approximations of the betweenness centrality. Let $b(v)$ be the actual betweenness centrality of node $v$ and $b'(v)$ be the betweenness centrality returned by algorithm 1. Then following lemma holds:

\begin{mylemma}
With probability atleast $1-\delta$, all the approximations computed by the algorithm are within $\epsilon$ from their real value:
$$Pr(\exists v \in V s.t. |b(v) - b'(v)| > \epsilon) < \delta$$
\end{mylemma}

The runtime of this algorithm is dominated by the computation of shortest path at each step which takes time $O(|V| + |E|)$ if graph is unweighted (BFS) and $O(|E| + |V|log|V|)$ othewise (Dijkstra's algorithm). This is multiplied by $r$ for $r$ iterations to obtain final complexity.

\section{Methods}

\subsection*{De Bruijn Graph}

De Bruijn strings are a type of 'comprehensive' strings. For a given alphabet $\Sigma = \{A,C,T,G\}$ and a length $k$, a de Bruijn string of order $k$ contains as substrings of length $k$ over $\Sigma$. These strings can be represented in a sophisticated way in term of a graph called de Bruijn graph \cite{debruijn}. A de Bruijn graph of order $k$ is a graph that contains all strings of length $k-1$ from a given alphabets as nodes and contains edges between two nodes if the corresponding strings overlap by exactly $k-2$ letters in prefix-suffix manner. In other words, each $k$-mer is represented by two nodes in a graph connected by an edge.

We modify the de Bruijn graph generated from genomes to identify repeats in a better way. We define \textit{unipaths} as the path of nodes in which the outdegree of first node on the path is $1$ and indegree is greater than $1$, indegree of last node on the path is $1$ and outdegree is greater than $1$ and for all other nodes, both indegree and out degree is exactly $1$. We compress all the unipaths in a graph to represent in a compact way. For all the internal nodes on unipaths, we represent them by a single node with label as the combined label of all nodes. figure\ref{fig:repeatgraph} shows an example of such compression of unipaths in de Bruijn graphs.

\begin{figure}[htbp]
\centering
\includegraphics[width = \textwidth]{repeatgraph}
\caption{Example of conversion of de Bruijn graph to Repeats graph}
\label{fig:repeatgraph}
\end{figure}

\subsection*{Networkit}
We investigated the performance of approximate betweenness centrality algorithms on the metagenomic graphs. We used the implementation in Networkit\cite{networkit} library. Its a c++ library with parallel implementation of many graph algorithms. We used the implementation of approximate betweenness centrality algorithm by (cite matteo). This algorithm is implemented in parallel.We also investigated the performance of algorithm by (cite other) on graphs. This algorithm is not parallelized. It takes number of sample nodes as an input.

\subsection*{Cutoff Criteria For Repeats}
To determine which nodes in a graph should be marked as repeats, we define a criteria based on the value of centrality calculated by the algorithm. Let $x$ be the mean of centralities of all the nodes and $\sigma$ be the standard deviation. A node $v$ is marked as repeat only if the following condition holds: $b(v) \geq x + c*\sigma$ where $c$ is a constant and $b(v)$ is betweenness centrality of the node. We set value of $c$ as 3 as it works out the best.

\begin{thebibliography}{4}

\bibitem{kingsford} Kingsford, Carl, Michael C. Schatz, and Mihai Pop. "Assembly complexity of prokaryotic genomes using short reads." BMC bioinformatics 11.1 (2010).

\bibitem{bambus} Koren, Sergey, Todd J. Treangen, and Mihai Pop. "Bambus 2: scaffolding metagenomes." Bioinformatics 27.21 (2011): 2964-2971.

\bibitem{brandes} Brandes, Ulrik. "A faster algorithm for betweenness centrality*." Journal of Mathematical Sociology 25.2 (2001): 163-177.

\bibitem{bader} Madduri, Kamesh, et al. "A faster parallel algorithm and efficient multithreaded implementations for evaluating betweenness centrality on massive datasets." Parallel \& Distributed Processing, 2009. IPDPS 2009. IEEE International Symposium on. IEEE, 2009.

\bibitem{bp} Bader, David A., et al. "Approximating betweenness centrality." Algorithms and Models for the Web-Graph. Springer Berlin Heidelberg, 2007. 124-137.

\bibitem{sanders}Geisberger, Robert, Peter Sanders, and Dominik Schultes. "Better Approximation of Betweenness Centrality." ALENEX. 2008.

\bibitem{matteo} Riondato, Matteo, and Evgenios M. Kornaropoulos. "Fast approximation of betweenness centrality through sampling." Proceedings of the 7th ACM international conference on Web search and data mining. ACM, 2014.

\bibitem{vc} V. N. Vapnik and A. J. Chervonenkis. On the uniform convergence of relative frequencies of events
to their probabilities. Theory of Probability and its Applications, 16(2):264–280, 1971. doi: 10.1137/
1116025.

\bibitem{debruijn} Kleene, Stephen Cole, et al. "Introduction to metamathematics." (1952).

\bibitem{networkit} Staudt, Christian L., Aleksejs Sazonovs, and Henning Meyerhenke. "Networkit: An interactive tool suite for high-performance network analysis." arXiv preprint arXiv:1403.3005 (2014).

\end{thebibliography}
\end{document}

### 2015-05-03 manuscript/manuscript.tex
refereces moved to bib

@@
 \section{Related Work}
+
+\subsection*{Betweenness Centrality}
 In graph theory and network analysis, indicators of centrality identify the most important vertices within a graph.Several metrics to measure centrality have been proposed.
@@
 
+\subsection*{Betweenness Centrality Based on Sampling Nodes}
 Due to scalability limitations of exact betweenness centrality, several approximation algorithms for betweenness centrality have been proposed.
@@
 
+\subsection*{Betweenness Centrality Based on Sampling Shortest Paths}
 There  is another approximation algorithm was proposed by Matteo et al.
@@
 Calculating exact VD(G) will need to solve all pair shortest  path problem.
-The complexity of this algorithm is $O(V^{2}logV + VE)$ by Johnson's algorithm(cite) and $\Theta(V^{3})$ by Floyd-Warshall algorithm.
+The complexity of this algorithm is $O(V^{2}logV + VE)$ by Johnson's algorithm\cite{johnson} and $\Theta(V^{3})$ by Floyd-Warshall algorithm.
 This would be very slow.
@@
 Second is to sample shortest paths.
-They way a random shortest path from $S_{uv}$ is inspired by Brandes algorithm(cite).
+They way a random shortest path from $S_{uv}$ is inspired by Brandes algorithm\cite{brandes}.
 Assume that $u$ and $v$ are connected.
@@
 Its a c++ library with parallel implementation of many graph algorithms.
-We used the implementation of approximate betweenness centrality algorithm by (cite matteo).
+We used the implementation of approximate betweenness centrality algorithm by\cite{matteo}.
 This algorithm is implemented in parallel.We also investigated the performance of algorithm by (cite other) on graphs.
@@
 
-\begin{thebibliography}{4}
+\bibliography{manuscript}{}
+\bibliographystyle{abbrv}
 
-\bibitem{kingsford} Kingsford, Carl, Michael C. Schatz, and Mihai Pop. "Assembly complexity of prokaryotic genomes using short reads." BMC bioinformatics 11.1 (2010).
-
-\bibitem{bambus} Koren, Sergey, Todd J. Treangen, and Mihai Pop. "Bambus 2: scaffolding metagenomes." Bioinformatics 27.21 (2011): 2964-2971.
-
-\bibitem{brandes} Brandes, Ulrik. "A faster algorithm for betweenness centrality*." Journal of Mathematical Sociology 25.2 (2001): 163-177.
-
-\bibitem{bader} Madduri, Kamesh, et al.
-"A faster parallel algorithm and efficient multithreaded implementations for evaluating betweenness centrality on massive datasets." Parallel \& Distributed Processing, 2009.
-IPDPS 2009.
-IEEE International Symposium on.
-IEEE, 2009.
-
-\bibitem{bp} Bader, David A., et al. "Approximating betweenness centrality." Algorithms and Models for the Web-Graph. Springer Berlin Heidelberg, 2007. 124-137.
-
-\bibitem{sanders}Geisberger, Robert, Peter Sanders, and Dominik Schultes. "Better Approximation of Betweenness Centrality." ALENEX. 2008.
-
-\bibitem{matteo} Riondato, Matteo, and Evgenios M.
-Kornaropoulos.
-"Fast approximation of betweenness centrality through sampling." Proceedings of the 7th ACM international conference on Web search and data mining.
-ACM, 2014.
-
-\bibitem{vc} V. N. Vapnik and A. J. Chervonenkis. On the uniform convergence of relative frequencies of events
-to their probabilities. Theory of Probability and its Applications, 16(2):264–280, 1971. doi: 10.1137/
-1116025.
-
-\bibitem{debruijn} Kleene, Stephen Cole, et al. "Introduction to metamathematics." (1952).
-
-\bibitem{networkit} Staudt, Christian L., Aleksejs Sazonovs, and Henning Meyerhenke.
-"Networkit: An interactive tool suite for high-performance network analysis." arXiv preprint arXiv:1403.3005 (2014).
-
-\end{thebibliography}
 \end{document}

### 2015-05-03 evaluate.py
evaluation script added

@@ -1,0 +1,48 @@
+import sys
+
+def main():
+	fasta_file = sys.argv[1]
+	repeat_file = sys.argv[2]
+	f_file = open(fasta_file,"rb")
+	r_file = open(repeat_file,"rb")
+	genome_map = {}
+	lines = f_file.readlines()
+	key = ""
+	for line in lines:
+		line = line[:-1]
+		if line[0] == '>':
+			key = line
+		else:
+			genome_map[key] = line
+	false_positive = 0
+	true_positive = 0
+	true_negative = 0
+	false_negative = 0
+	lines = r_file.readlines()
+	for line in lines:
+		attrs = line.split("\t")
+		label = attrs[3]
+		isrepeat = int(attrs[-1][:-1])
+		label = label.strip("\"")
+		cnt = 0
+		for key in genome_map:
+			genome = genomo_map[key]
+			if label in genome:
+				cnt += 1
+		if cnt > 1 and isrepeat == 1:
+			true_positive += 1
+		if cnt > 1 and isrepeat = 0:
+			false_positive += 1
+		if cnt == 1 and isrepeat == 1:
+			false_negative += 1
+		if cnt == 1 and isrepeat == 0:
+			true_negative += 1
+	sensitivity = true_positive*1.0/(true_positive + false_negative)
+	specificity = true_negative*1.0/(true_negative + false_positive)
+
+	print "sensitivity = " + str(sensitivity)
+	print "specificity = " + str(specificity)
+
+
+if __name__ == "__main__":
+	main()

### 2015-05-03 evaluate.py
modified script

@@ -19,19 +19,20 @@ def main():
 	true_negative = 0
 	false_negative = 0
 	lines = r_file.readlines()
+	revcompl = lambda x: ''.join([{'A':'T','C':'G','G':'C','T':'A'}[B] for B in x][::-1])
 	for line in lines:
 		attrs = line.split("\t")
-		label = attrs[3]
+		label = attrs[2]
 		isrepeat = int(attrs[-1][:-1])
 		label = label.strip("\"")
 		cnt = 0
 		for key in genome_map:
-			genome = genomo_map[key]
-			if label in genome:
+			genome = genome_map[key]
+			if label in genome or revcompl(label) in genome:
 				cnt += 1
 		if cnt > 1 and isrepeat == 1:
 			true_positive += 1
-		if cnt > 1 and isrepeat = 0:
+		if cnt > 1 and isrepeat == 0:
 			false_positive += 1
 		if cnt == 1 and isrepeat == 1:
 			false_negative += 1
@@ -42,6 +43,10 @@ def main():
 
 	print "sensitivity = " + str(sensitivity)
 	print "specificity = " + str(specificity)
+	print "true positives = " + str(true_positive)
+	print "true negatives = " +str(true_negative)
+	print "false positives = " + str(false_positive)
+	print "false negatives = " + str(false_negative)
 
 
 if __name__ == "__main__":

### 2015-05-03 sampling_algo.cpp

@@ -73,10 +73,12 @@ int main(int argc, char* argv[]) {
 	cerr<<"bound: "<<mean + filterparam*stdev<<endl;
 	double bound = mean + filterparam*stdev;
 	for (int i = 0; i < g.nodes().size(); i++) {
+		int isrepeat = 0;
 		if(nodeRanks.at(i).second >= bound){
-			string String = static_cast<ostringstream*>( &(ostringstream() << nodeRanks.at(i).first) )->str();
-			cout << nodeRanks.at(i).first <<"\t"<< nodeRanks.at(i).second<<"\t"<<labelmap[String]<< endl;
+			isrepeat = 1;
 		}
+		string String = static_cast<ostringstream*>( &(ostringstream() << nodeRanks.at(i).first) )->str();
+		cout << nodeRanks.at(i).first <<"\t"<< nodeRanks.at(i).second<<"\t"<<labelmap[String]<<isrepeat<< endl;
 	}
 	cerr << "Execution Time = " << (double) (clock() - start) / CLOCKS_PER_SEC
 			<< endl;

### 2015-05-03 vc_algo.cpp

@@ -57,6 +57,7 @@ int main(int argc, char* argv[]) {
 	centrality.run();
 	ofstream myfile;
 	filepath = filepath.substr(0,filepath.length()-4);
+	myfile.open(filepath+".repeats");
 	vector<pair<node, double>> nodeRanks(g.nodes().size());
 	nodeRanks = centrality.ranking();
 	vector<double> scores = centrality.scores();
@@ -71,11 +72,13 @@ int main(int argc, char* argv[]) {
 	cerr<<"bound: "<<mean + filterparam*stdev<<endl;
 	double bound = mean + filterparam*stdev;
 	for (int i = 0; i < g.nodes().size(); i++) {
+		int isrepeat = 0;
 		if(nodeRanks.at(i).second >= bound){
-			string String = static_cast<ostringstream*>( &(ostringstream() << nodeRanks.at(i).first) )->str();
-//			myfile << nodeRanks.at(i).first <<"\t"<< nodeRanks.at(i).second<<"\t"<<labelmap[String]<< endl;
-			cout << nodeRanks.at(i).first <<"\t"<< nodeRanks.at(i).second<<"\t"<<labelmap[String]<< endl;
+			isrepeat = 1;
 		}
+		string String = static_cast<ostringstream*>( &(ostringstream() << nodeRanks.at(i).first) )->str();
+//		myfile << nodeRanks.at(i).first <<"\t"<< nodeRanks.at(i).second<<"\t"<<labelmap[String]<< endl;
+	myfile << nodeRanks.at(i).first <<"\t"<< nodeRanks.at(i).second<<"\t"<<labelmap[String]<<"\t"<<isrepeat<< endl;
 	}
 	cerr << "Execution Time = " << (double) (clock() - start) / CLOCKS_PER_SEC
 			<< endl;

### 2015-05-03 sampling_algo.cpp
fixed output

@@ -47,6 +47,8 @@ int main(int argc, char* argv[]) {
 		}
 		if(str.find("label") != string::npos) {
 			label = str.substr(10);
+			label.erase(0,1);
+			label.erase(label.length()-1);
 			labelmap[id] = label;
 		}
 	}

### 2015-05-03 vc_algo.cpp

@@ -13,6 +13,7 @@
 #include <string>
 #include <numeric>
 #include <map>
+#include <algorithm>
 
 #include <NetworKit/graph/Graph.h>
 #include <NetworKit/io/GMLGraphReader.h>
@@ -45,6 +46,8 @@ int main(int argc, char* argv[]) {
 		}
 		if(str.find("label") != string::npos) {
 			label = str.substr(10);
+			label.erase(0,1);
+			label.erase(label.length()-1);
 			labelmap[id] = label;
 		}
 	}

### 2015-05-04 evaluate.py
fixed sampling algo

@@ -38,15 +38,19 @@ def main():
 			false_negative += 1
 		if cnt == 1 and isrepeat == 0:
 			true_negative += 1
+	
+	
+        print "true positives = " + str(true_positive)
+        print "true negatives = " +str(true_negative)
+        print "false positives = " + str(false_positive)
+        print "false negatives = " + str(false_negative)
+	
 	sensitivity = true_positive*1.0/(true_positive + false_negative)
 	specificity = true_negative*1.0/(true_negative + false_positive)
 
 	print "sensitivity = " + str(sensitivity)
 	print "specificity = " + str(specificity)
-	print "true positives = " + str(true_positive)
-	print "true negatives = " +str(true_negative)
-	print "false positives = " + str(false_positive)
-	print "false negatives = " + str(false_negative)
+	
 
 
 if __name__ == "__main__":

### 2015-05-04 sampling_algo.cpp

@@ -62,6 +62,7 @@ int main(int argc, char* argv[]) {
 	centrality.run();
 	ofstream myfile;
 	filepath = filepath.substr(0,filepath.length()-4);
+	myfile.open(filepath+".repeats");
 	vector<pair<node, double>> nodeRanks(g.nodes().size());
 	nodeRanks = centrality.ranking();
 	vector<double> scores = centrality.scores();
@@ -80,7 +81,7 @@ int main(int argc, char* argv[]) {
 			isrepeat = 1;
 		}
 		string String = static_cast<ostringstream*>( &(ostringstream() << nodeRanks.at(i).first) )->str();
-		cout << nodeRanks.at(i).first <<"\t"<< nodeRanks.at(i).second<<"\t"<<labelmap[String]<<isrepeat<< endl;
+		myfile << nodeRanks.at(i).first <<"\t"<< nodeRanks.at(i).second<<"\t"<<labelmap[String]<<"\t"<<isrepeat<< endl;
 	}
 	cerr << "Execution Time = " << (double) (clock() - start) / CLOCKS_PER_SEC
 			<< endl;

### 2015-05-04 evaluate.py
evaluation fixed

@@ -33,9 +33,9 @@ def main():
 		if cnt > 1 and isrepeat == 1:
 			true_positive += 1
 		if cnt > 1 and isrepeat == 0:
-			false_positive += 1
-		if cnt == 1 and isrepeat == 1:
 			false_negative += 1
+		if cnt == 1 and isrepeat == 1:
+			false_positive += 1
 		if cnt == 1 and isrepeat == 0:
 			true_negative += 1
 	

### 2015-05-04 runnode.sh
added scripts

@@ -1,0 +1,13 @@
+#!/bin/bash
+
+# mix_graph = $1
+#cutoff = $2
+#nodes = $3
+#repeat_file = $4
+#fasta_file = $5
+
+make
+
+./sampling_algo $1 $2 $3 $4
+
+python evaluate.py $5 $4

### 2015-05-04 runpath.sh

@@ -1,0 +1,12 @@
+#!/bin/bash
+
+# mix_graph = $1
+#cutoff = $2
+#repeat_file = $3
+#fasta_file = $4
+
+make
+
+./vc_algo $1 $2 $3
+
+python evaluate.py $4 $3

### 2015-05-04 sampling_algo.cpp

@@ -32,6 +32,7 @@ int main(int argc, char* argv[]) {
 	filepath = argv[1];
 	filterparam = atoi(argv[2]);
 	sampleNodes = atoi(argv[3]);
+	string repeat_file = argv[4];
 	clock_t start = clock();
 	GMLGraphReader reader;
 	Graph g = reader.read(argv[1]);
@@ -62,7 +63,7 @@ int main(int argc, char* argv[]) {
 	centrality.run();
 	ofstream myfile;
 	filepath = filepath.substr(0,filepath.length()-4);
-	myfile.open(filepath+".repeats");
+	myfile.open(repeat_file);
 	vector<pair<node, double>> nodeRanks(g.nodes().size());
 	nodeRanks = centrality.ranking();
 	vector<double> scores = centrality.scores();

### 2015-05-04 vc_algo.cpp

@@ -31,6 +31,7 @@ int main(int argc, char* argv[]) {
 	int filterparam;
 	filepath = argv[1];
 	filterparam = atoi(argv[2]);
+	string repeat_file = argv[3];
 	clock_t start = clock();
 	GMLGraphReader reader;
 	Graph g = reader.read(argv[1]);
@@ -60,7 +61,7 @@ int main(int argc, char* argv[]) {
 	centrality.run();
 	ofstream myfile;
 	filepath = filepath.substr(0,filepath.length()-4);
-	myfile.open(filepath+".repeats");
+	myfile.open(repeat_file);
 	vector<pair<node, double>> nodeRanks(g.nodes().size());
 	nodeRanks = centrality.ranking();
 	vector<double> scores = centrality.scores();

### 2015-05-04 manuscript/manuscript.tex
evaluation script added
modified report

@@
 % first the title is needed
-\title{Lecture Notes in Computer Science:\\Authors' Instructions
-for the Preparation\\of Camera-Ready
-Contributions\\to LNCS/LNAI/LNBI Proceedings}
+\title{Identifiyng Inter-Genomic Repeats Using Fast, Approximate Betweenness Centrality}
 
@@
 \subsection*{Networkit}
-We investigated the performance of approximate betweenness centrality algorithms on the metagenomic graphs.
+We investigated the performance of approximate betweenness centrality algorithms on the metagenomic graphs to find inter-genome repeats.
 We used the implementation in Networkit\cite{networkit} library.
-Its a c++ library with parallel implementation of many graph algorithms.
-We used the implementation of approximate betweenness centrality algorithm by\cite{matteo}.
-This algorithm is implemented in parallel.We also investigated the performance of algorithm by (cite other) on graphs.
-This algorithm is not parallelized.
-It takes number of sample nodes as an input.
+It provides implementation for various graph algorithms.
+We are particularly interested in the implementation of betweenness centrality algorithms.
+We make use of two approximate betweenness centrality algorithms provided by NetworKit.
+
+NetworKit provides fast parallel implementation of approximate betweenness centrality algorithm by Matteo et al.
+\cite{matteo}.
+The implementation first approximates diameter of of graph.
+After this, it calculates the number of shortest paths to be sampled.
+It processes shortest paths in the sample in parallel in multithreaded fashion and then aggregates the centrality scores for each shortest paths once every thread finishes processing of shortest path assigned to it.
+
+We also investigate the performance of algorithm by Geisberger et al.
+\cite{sanders}.
+NetworKit provides single threaded implementation of this algorithm.
+It first creates a sample of $n$ nodes at random from a given graph.
+It then runs single source shortest path algorithm for each of those nodes.
+It calculates the betweenness scores for each node by the modified aggregation scheme proposed in the algorithm.
+
+Here are few questions we want to investigate based on these algorithms:
+
+\begin{itemize}
+\item \textbf{Scalability}: How these algorithms scale on large metagenomic repeats graph? Does parallelized algorithm provides quick results or taking sample of fewer nodes provides quick results?
+\item \textbf{Accuracy}: Which of these algorithms provides better accuracy while determining inter-genomic repeats?
+\item \textbf{Sample Size}: For the algorithm by Geisberger et al.
+\cite{sanders}, how should a sample size be defined so that it would produce fast and accurate results?
+We try to come up with a heuristic to define sample size
+\end{itemize}
 
@@
 
+\section{Results} 
+We evaluated algorithms for betweenness centrality to find inter-genome repeats on real datasets described above.
+We use the notion of sensitivity and specificity to define the accuracy of repeat detection.
+We define following terms:
+\begin{itemize}
+\item \textbf{True positives:} Number of nodes marked as repeat that is found within at least 2 genomes
+\item \textbf{False positives:} Number of nodes marked as repeat that is only repetitive within a single genome
+\item \textbf{True negatives:} Number of nodes marked as non-repeat that is only repetitive within a single genome
+\item \textbf{False negatives:} Number of nodes marked as non-repeat that is found within at least 2 genomes
+\end{itemize}
+
+A good repeat detection should have both good sensitivity and specificity.
+Sensitivity reflects how many true repeats are detected.
+Detecting to few repeats can lead to assembly errors in scaffolding.
+Specificity reflects the false positives.
+Detecting too many repeats leads to a suboptimal assembly as these contigs do not fully participate in scaffolding.
+
 \bibliography{manuscript}{}

### 2015-05-05 manuscript/manuscript.tex

--- 2015-05-04 manuscript/manuscript.tex
@@
 \begin{abstract}
-The abstract should summarize the contents of the paper and should
-contain at least 70 and at most 150 words. It should be written using the
-\emph{abstract} environment.
-\keywords{We would like to encourage you to list your keywords within
-the abstract section}
+Assembling metagenomic data is a challenging task due to several reasons.
+One of the reasons that complicates assembly is the sequences shared between the genomes of multiple organisms.
+Such regions in genomes are called repeats.
+Repeats make assembly difficult because they link the unrelated sections of genomes.
+It has been shown that the number of reconstructions of genome grows exponentially with repeats.
+Existing approaches to identify repeats are inefficient and do not scale well as size of data increases.
+We propose an application of a core concept in social network analysis called betweenness centrality to metagenomic data to identify inter-genomic repeats.
+We study how approximation algorithms for betweenness centrality can be useful to find central nodes quickly and propose a novel approach to identify inter-genomic repeats accurately.
+\keywords{Metagenomics, Algorithms, Centrality, Graph}
 \end{abstract}
@@
 
+\section{Conclusion}
+
 \bibliography{manuscript}{}

### 2015-05-05 generate_test.py
preprocessing script added

@@ -1,0 +1,41 @@
+import sys
+
+def main():
+	fasta_file = sys.argv[1]
+	repeat_file = sys.argv[2]
+	f_file = open(fasta_file,"rb")
+	r_file = open(repeat_file,"rb")
+	genome_map = {}
+	repeat_map = {}
+	lines = f_file.readlines()
+	key = ""
+	for line in lines:
+		line = line[:-1]
+		if line[0] == '>':
+			key = line
+		else:
+			genome_map[key] = line
+	
+	lines = r_file.readlines()
+	revcompl = lambda x: ''.join([{'A':'T','C':'G','G':'C','T':'A'}[B] for B in x][::-1])
+	for line in lines:
+		attrs = line.split("\t")
+		label = attrs[2]
+		isrepeat = int(attrs[-1][:-1])
+		label = label.strip("\"")
+		cnt = 0
+		for key in genome_map:
+			genome = genome_map[key]
+			if label in genome or revcompl(label) in genome:
+				cnt += 1
+		repeat_map[label] = cnt
+	
+	
+    ofile = open(sys.argv[3],"rb")
+    for key in repeat_map:
+    	ofile.write(str(key)+"\t"+str(repeat_map[key])+"\n")
+	
+
+
+if __name__ == "__main__":
+	main()

### 2015-05-05 generate_test.py
epsilon added, script modified

@@ -1,5 +1,6 @@
 import sys
 
+
 def main():
 	fasta_file = sys.argv[1]
 	repeat_file = sys.argv[2]
@@ -14,10 +15,11 @@ def main():
 		if line[0] == '>':
 			key = line
 		else:
+			print len(line)
 			genome_map[key] = line
-	
 	lines = r_file.readlines()
 	revcompl = lambda x: ''.join([{'A':'T','C':'G','G':'C','T':'A'}[B] for B in x][::-1])
+	node = 0
 	for line in lines:
 		attrs = line.split("\t")
 		label = attrs[2]
@@ -26,14 +28,15 @@ def main():
 		cnt = 0
 		for key in genome_map:
 			genome = genome_map[key]
-			if label in genome or revcompl(label) in genome:
+			if genome.find(label) == True or genome.find(revcompl(label)):
 				cnt += 1
 		repeat_map[label] = cnt
+		print node
+		node += 1
 	
-	
-    ofile = open(sys.argv[3],"rb")
-    for key in repeat_map:
-    	ofile.write(str(key)+"\t"+str(repeat_map[key])+"\n")
+    	ofile = open(sys.argv[3],"w")
+    	for key in repeat_map:
+    		ofile.write(str(key)+"\t"+str(repeat_map[key])+"\n")
 	
 
 

### 2015-05-05 runpath.sh

@@ -2,11 +2,11 @@
 
 # mix_graph = $1
 #cutoff = $2
-#repeat_file = $3
-#fasta_file = $4
+#epsilon = $3
+#repeat_file = $4
+#fasta_file = $5
 
-make
 
-./vc_algo $1 $2 $3
+./vc_algo $1 $2 $3 $4
 
-python evaluate.py $4 $3
+python evaluate.py $5 $4

### 2015-05-05 sampling_algo.cpp

@@ -76,6 +76,9 @@ int main(int argc, char* argv[]) {
 	cerr<<"Std Dev: "<<stdev<<endl;
 	cerr<<"bound: "<<mean + filterparam*stdev<<endl;
 	double bound = mean + filterparam*stdev;
+	 cerr << "Execution Time = " << (double) (clock() - start) / CLOCKS_PER_SEC
+                        << endl;
+
 	for (int i = 0; i < g.nodes().size(); i++) {
 		int isrepeat = 0;
 		if(nodeRanks.at(i).second >= bound){
@@ -84,8 +87,7 @@ int main(int argc, char* argv[]) {
 		string String = static_cast<ostringstream*>( &(ostringstream() << nodeRanks.at(i).first) )->str();
 		myfile << nodeRanks.at(i).first <<"\t"<< nodeRanks.at(i).second<<"\t"<<labelmap[String]<<"\t"<<isrepeat<< endl;
 	}
-	cerr << "Execution Time = " << (double) (clock() - start) / CLOCKS_PER_SEC
-			<< endl;
+
 	return 0;
 }
 

### 2015-05-05 vc_algo.cpp

@@ -31,7 +31,9 @@ int main(int argc, char* argv[]) {
 	int filterparam;
 	filepath = argv[1];
 	filterparam = atoi(argv[2]);
-	string repeat_file = argv[3];
+	string eps = argv[3];
+	double epsilon = (double)atof(eps.c_str());
+	string repeat_file = argv[4];
 	clock_t start = clock();
 	GMLGraphReader reader;
 	Graph g = reader.read(argv[1]);
@@ -56,7 +58,7 @@ int main(int argc, char* argv[]) {
 	cerr << "Nodes: " << g.nodes().size() << endl;
 	cerr << "Edges: " << g.edges().size() << endl;
 	double delta = 0.1;
-	double epsilon = 0.1;
+	
 	ApproxBetweenness centrality(g, delta, epsilon, 2);
 	centrality.run();
 	ofstream myfile;
@@ -82,6 +84,9 @@ int main(int argc, char* argv[]) {
 	cerr<<"bound: "<<mean + filterparam*stdev<<endl;
 	cerr<<"filter: "<<filter<<endl;
 	double bound = mean + filterparam*stdev;
+	cerr << "Execution Time = " << (double) (clock() - start) / CLOCKS_PER_SEC
+                        << endl;
+
 	for (int i = 0; i < g.nodes().size(); i++) {
 		int isrepeat = 0;
 		if(nodeRanks.at(i).second >= bound){
@@ -91,8 +96,7 @@ int main(int argc, char* argv[]) {
 //		myfile << nodeRanks.at(i).first <<"\t"<< nodeRanks.at(i).second<<"\t"<<labelmap[String]<< endl;
 	myfile << nodeRanks.at(i).first <<"\t"<< nodeRanks.at(i).second<<"\t"<<labelmap[String]<<"\t"<<isrepeat<< endl;
 	}
-	cerr << "Execution Time = " << (double) (clock() - start) / CLOCKS_PER_SEC
-			<< endl;
+
 	return 0;
 }
 

### 2015-05-05 evaluate.py
output modified

@@ -1,10 +1,12 @@
 import sys
+import os
 
 def main():
 	fasta_file = sys.argv[1]
 	repeat_file = sys.argv[2]
 	f_file = open(fasta_file,"rb")
 	r_file = open(repeat_file,"rb")
+	param = sys.argv[3]
 	genome_map = {}
 	lines = f_file.readlines()
 	key = ""
@@ -39,17 +41,12 @@ def main():
 		if cnt == 1 and isrepeat == 0:
 			true_negative += 1
 	
-	
-        print "true positives = " + str(true_positive)
-        print "true negatives = " +str(true_negative)
-        print "false positives = " + str(false_positive)
-        print "false negatives = " + str(false_negative)
-	
-	sensitivity = true_positive*1.0/(true_positive + false_negative)
-	specificity = true_negative*1.0/(true_negative + false_positive)
-
-	print "sensitivity = " + str(sensitivity)
-	print "specificity = " + str(specificity)
+	runfile = open("runtime","rb")
+	line = runfile.readline()
+#	print order nodes/epsilon	runtime	true_pos	true_neg	false_pos	false_neg	
+        line = line.strip("\n")
+	print str(param) + "\t" + line + "\t" + str(true_positive) + "\t" + str(true_negative) + "\t" + str(false_positive) + "\t" + str(false_negative)
+	os.remove("runtime")
 	
 
 

### 2015-05-05 generate_test.py

@@ -1,44 +1,0 @@
-import sys
-
-
-def main():
-	fasta_file = sys.argv[1]
-	repeat_file = sys.argv[2]
-	f_file = open(fasta_file,"rb")
-	r_file = open(repeat_file,"rb")
-	genome_map = {}
-	repeat_map = {}
-	lines = f_file.readlines()
-	key = ""
-	for line in lines:
-		line = line[:-1]
-		if line[0] == '>':
-			key = line
-		else:
-			print len(line)
-			genome_map[key] = line
-	lines = r_file.readlines()
-	revcompl = lambda x: ''.join([{'A':'T','C':'G','G':'C','T':'A'}[B] for B in x][::-1])
-	node = 0
-	for line in lines:
-		attrs = line.split("\t")
-		label = attrs[2]
-		isrepeat = int(attrs[-1][:-1])
-		label = label.strip("\"")
-		cnt = 0
-		for key in genome_map:
-			genome = genome_map[key]
-			if genome.find(label) == True or genome.find(revcompl(label)):
-				cnt += 1
-		repeat_map[label] = cnt
-		print node
-		node += 1
-	
-    	ofile = open(sys.argv[3],"w")
-    	for key in repeat_map:
-    		ofile.write(str(key)+"\t"+str(repeat_map[key])+"\n")
-	
-
-
-if __name__ == "__main__":
-	main()

### 2015-05-05 runpath.sh

@@ -9,4 +9,4 @@
 
 ./vc_algo $1 $2 $3 $4
 
-python evaluate.py $5 $4
+python evaluate.py $5 $4 $3

### 2015-05-05 sampling_algo.cpp

@@ -76,8 +76,10 @@ int main(int argc, char* argv[]) {
 	cerr<<"Std Dev: "<<stdev<<endl;
 	cerr<<"bound: "<<mean + filterparam*stdev<<endl;
 	double bound = mean + filterparam*stdev;
-	 cerr << "Execution Time = " << (double) (clock() - start) / CLOCKS_PER_SEC
-                        << endl;
+
+	ofstream runfile;
+        runfile.open("runtime");
+        runfile << (double) (clock() - start) / CLOCKS_PER_SEC << endl;
 
 	for (int i = 0; i < g.nodes().size(); i++) {
 		int isrepeat = 0;

### 2015-05-05 vc_algo.cpp

@@ -84,8 +84,9 @@ int main(int argc, char* argv[]) {
 	cerr<<"bound: "<<mean + filterparam*stdev<<endl;
 	cerr<<"filter: "<<filter<<endl;
 	double bound = mean + filterparam*stdev;
-	cerr << "Execution Time = " << (double) (clock() - start) / CLOCKS_PER_SEC
-                        << endl;
+	ofstream runfile;
+	runfile.open("runtime");
+	runfile << (double) (clock() - start) / CLOCKS_PER_SEC << endl;
 
 	for (int i = 0; i < g.nodes().size(); i++) {
 		int isrepeat = 0;

### 2015-05-05 runnode.sh
runnode modified

@@ -6,8 +6,6 @@
 #repeat_file = $4
 #fasta_file = $5
 
-make
-
 ./sampling_algo $1 $2 $3 $4
 
-python evaluate.py $5 $4
+python evaluate.py $5 $4 $3

### 2015-05-05 sampling_algo.cpp
bugs removed

@@ -26,7 +26,6 @@ using namespace NetworKit;
 int main(int argc, char* argv[]) {
 
 	string filepath;
-	cout<<argv[1]<<endl;
 	int filterparam;
 	int sampleNodes;
 	filepath = argv[1];

### 2015-05-05 vc_algo.cpp

@@ -27,7 +27,6 @@ using namespace NetworKit;
 int main(int argc, char* argv[]) {
 
 	string filepath;
-	cout<<argv[1]<<endl;
 	int filterparam;
 	filepath = argv[1];
 	filterparam = atoi(argv[2]);

### 2015-05-05 evaluate.py
minor bug fixed

@@ -41,12 +41,12 @@ def main():
 		if cnt == 1 and isrepeat == 0:
 			true_negative += 1
 	
-	runfile = open("runtime","rb")
+	runfile = open(repeat_file+".runtime","rb")
 	line = runfile.readline()
 #	print order nodes/epsilon	runtime	true_pos	true_neg	false_pos	false_neg	
         line = line.strip("\n")
 	print str(param) + "\t" + line + "\t" + str(true_positive) + "\t" + str(true_negative) + "\t" + str(false_positive) + "\t" + str(false_negative)
-	os.remove("runtime")
+	os.remove(repeat_file+".runtime")
 	
 
 

### 2015-05-05 sampling_algo.cpp

@@ -77,7 +77,7 @@ int main(int argc, char* argv[]) {
 	double bound = mean + filterparam*stdev;
 
 	ofstream runfile;
-        runfile.open("runtime");
+        runfile.open(repeat_file+".runtime");
         runfile << (double) (clock() - start) / CLOCKS_PER_SEC << endl;
 
 	for (int i = 0; i < g.nodes().size(); i++) {

### 2015-05-05 vc_algo.cpp

@@ -84,7 +84,7 @@ int main(int argc, char* argv[]) {
 	cerr<<"filter: "<<filter<<endl;
 	double bound = mean + filterparam*stdev;
 	ofstream runfile;
-	runfile.open("runtime");
+	runfile.open(repeat_file+".runtime");
 	runfile << (double) (clock() - start) / CLOCKS_PER_SEC << endl;
 
 	for (int i = 0; i < g.nodes().size(); i++) {

### 2015-05-06 runnode.sh
added scripts

@@ -4,8 +4,7 @@
 #cutoff = $2
 #nodes = $3
 #repeat_file = $4
-#fasta_file = $5
-
+#map fie = $5
 ./sampling_algo $1 $2 $3 $4
 
-python evaluate.py $5 $4 $3
+python evaluate1.py  $4 $3 $5

### 2015-05-06 runpath.sh

@@ -4,9 +4,8 @@
 #cutoff = $2
 #epsilon = $3
 #repeat_file = $4
-#fasta_file = $5
-
+#map file = $5
 
 ./vc_algo $1 $2 $3 $4
 
-python evaluate.py $5 $4 $3
+python evaluate1.py  $4 $3 $5

### 2015-05-06 evaluate.py
report modified

@@ -40,6 +40,7 @@ def main():
 			true_negative += 1
 	
 	
+#		print order nodes/epsilon	runtime	true_pos	true_neg	false_pos	false_neg
         print "true positives = " + str(true_positive)
         print "true negatives = " +str(true_negative)
         print "false positives = " + str(false_positive)

### 2015-05-06 manuscript/manuscript.tex

@@
 %
-\author{Alfred Hofmann%
-\thanks{Please note that the LNCS Editorial assumes that all authors have used
-the western naming convention, with given names preceding surnames. This determines
-the structure of the names in the running heads and the author index.}%
-\and Ursula Barth\and Ingrid Haas\and Frank Holzwarth\and\\
-Anna Kramer\and Leonie Kunz\and Christine Rei\ss\and\\
-Nicole Sator\and Erika Siebert-Cole\and Peter Stra\ss er}
+\author{Jay Ghurye%
+\and Chris Hill\and Mihai Pop}
 %
-\authorrunning{Lecture Notes in Computer Science: Authors' Instructions}
+\authorrunning{Identifiyng Inter-Genomic Repeats Using Fast,
+Approximate Betweenness Centrality}
 % (feature abused for this document to repeat the title also on left hand pages)
@@
 % unless you accept that it will be published
-\institute{Springer-Verlag, Computer Science Editorial,\\
-Tiergartenstr. 17, 69121 Heidelberg, Germany\\
+\institute{University of Maryland, College Park\\
 \mailsa\\
@@
 However, this algorithm overestimates the centrality of some unimportant nodes which lie near pivot.
-The approximation algorithm proposed in \cite{sanders} comes over this drawback by changing the scheme for aggregating betweenness contributions so that nodes do not get profited just because they are near pivot.
+The approximation algorithm proposed by Geisberger et al.
+\cite{sanders} comes over this drawback by changing the scheme for aggregating betweenness contributions so that nodes do not get profited just because they are near pivot.
 They define an estimator for betweenness centrality.
@@
 This algorithm guarantees that all approximate values of betweenness for all vertices are within an additive factor $\epsilon \in (0,1)$ from the real values with probability at least $1-\delta$.
-To derive bounds on sample size, results from VC dimension\cite{vc} theory are used.
+To derive bounds on sample size, results from VC dimension theory\cite{vc} are used.
 The main advantage of this algorithm is that the sample size does not depend on the size of the graph, but only on the maximum number vertices in a shortest path which is called \textit{vertex diameter}.
@@
 Given a probability distribution on the set of points, a finite bound on the VC-dimension of the class of subsets implies a bound on the number of random samples required to approximate the probability of each subset in the class with its empirical average.
-To define this formally, let $D$ be a domain and $R$ be a collection of subsets from $D$.
-We call $R$ a \textit{range set} on $D$.
-Given $B \subseteq D$, the projection of $R$ on $B$ is the set of $P_{R}(B) = \{B \cap A : A \in R\}$.We say that the set $B$ is \textit{shattered} by $R$ if $P_{R}(B) = 2^{B}$, where $2^{B}$ is the powerset of $B$.
-
-\begin{mydef}
-The \textit{Vapnik-Chervonenkis} (VC) dimension of $R$, denoted as $VC(R)$, is the cardinality of the largest subset of $D$ that is shattered by $R$.
-\end{mydef}
-
-It has been shown that if an upper bound to the VC-dimension of $R$ is available, it is possible to build an $\epsilon$-approximation by sampling the points of the domain according to distribution $\phi$.
-
-\begin{mydef}
-Let $R$ be a range set on a domain $D$ with VC(R) $\leq d$, and let $\phi$ be a distribution on $D$.
-Given $\epsilon, \delta \in (0,1)$ let $S$ be a collection of $|S|$ points from $D$ sampled according to $\phi$ with
-\begin{equation} \label{sample_size}
-|S| = \frac{c}{\epsilon^{2}}(d + ln\frac{1}{\delta})
-\end{equation}
-
-where $c$ is a universal positive constant approximated to 0.5. Then $S$ is an $\epsilon$-approximation to $(R,\phi)$ with probability at least $(1-\delta)$.
-\end{mydef}
-
-Using these insights from VC-dimensions, we can define a range set on the shortest paths of a graph $G = (V,E)$ and present a tight upper bound to its VC-dimensions.
-The range set $R_{G}$ is defined on the set $\mathtt{S}_{G}$ of all shortest paths between vertices of $G$.
-It contains, for each vertex $v \in V$, the set $T_{v}$ of shortest paths that $v$ is internal to:
-$$R_{G} = \{T_{v} : v \in V\}.$$
-Let us define for each shortest path $p$, $Int(p)$ be the set of internal vertices on that path.
-Let $H(G)$ be the maximum integer $d$ such that there are $h = \left\lfloor{log_{2}(d-2)}\right\rfloor + 1$ shortest paths $\{p_{1},p_{2},....p_{h}\}$
-in $\mathtt{S}_{G}$ of size at least $d$ and such that the collection $\{Int(p_{1}),....,Int(p_{h})\}$ is an anti-chain, that is a collection of sets such that for no two sets in it, one is a subset of other.
-It has been shown that this can be used to provide an upper bound on VC-dimension on set $R_{G}$.
-So we have:
-$$VC(R_{G}) \leq \left\lfloor{log_{2}(H(G) - 2)}\right\rfloor + 1$$
-It $VD(G)$ be the vertex diameter of graph G.
-It can be shown that $H(G) \leq VD(G)$.
-As calculating exact $H(G)$ value is hard, we can make use of $VD(G)$ to define upper bound on VC-dimension of $R_{G}$ as follows:
-$$VC(R_{G}) \leq \left\lfloor{log_{2}(VD(G) - 2)}\right\rfloor + 1$$
-
-The algorithm to calculate approximate betweenness centrality works as follows. For a graph $G(V,E)$, with vertex diameter $VD(G)$, sample size $r$ is estimated as follows:
-$$ d = \left\lfloor{log_{2}(VD(G) - 2)}\right\rfloor + 1$$
-Substituting this $d$ in equation\ref{sample_size}, we get
+This technique is used to derive the expressiveness of the sample of shortest paths derived.
+For graph $G(V,E)$, let $S_{all}$ be the union of $T_{v}$ for all $v \in V$, where $T_{v}$ is the set of all shortest path to which $v$ is internal.
+VC dimension technique helps in choosing sample from $S_{all}$ so that it gives probabilistic guarantees.
+It is show that the upper bound on VC dimensions of $S_{all}$ is given by:
+$$VC(S_{all}) \leq \left\lfloor{log_{2}(VD(G) - 2)}\right\rfloor + 1$$
+where $VD(G)$ is vertex diameter of $G$. Using this upper bound, sample size $r$ can be estimated as follows:
 \begin{equation} \label{sample_size_algo}
@@
 \end{equation} 
-This sample size achieves desired accuracy with the desired confidence.
-The algorithm repeats following steps $r$ times: 1.
-It samples a pair of $u,v$ of distinct vertices uniformly at random.
-2.
-It computes set $S_{uv}$ or all shortest paths between $u$ and $v$.
-3.
-It selects path $p$ from $S_{uv}$ at random.
-4.
-It increases by $1/r$ the betweenness estimation of each vertex in $Int(p)$.
-If sampled vertices $u$ and $v$ are not connected, steps 3 and 4 can be skipped.
+where $c$ is a universal constant with value $0.5$.
+It is guaranteed that betweenness centralities estimated using this sample size are within additive factor $\epsilon$ from real values with probability atleast $1-\delta$.
 
+Complete algorithm is sketched in algorithm 1.
 There are two crucial things in this algorithm.
@@
 This would be very slow.
-A faster linear time 2-approximation algorithm is proposed in \cite{matteo}.
-It is as follows:
-\begin{itemize}
-\item Select $v \in V$ at random
-\item Compute shortest path from $v$ to all vertices in $V$
-\item Finally take $VD(G)$ be the sum of lengths of two shortest paths with maximum size from $v$ to two distinct nodes $u$ and $w$. 
-\end{itemize}
+A faster linear time 2-approximation algorithm is proposed in \cite{matteo} is used to calculate approximate vertex diameter.
 
@@
 \section{Conclusion}
-
+In this work, we show how approximate betweenness centrality can be used to identify inter-genome repeats.
+We explored two algorithms, one based on sampling shortest paths in a graph and other based on sampling nodes in a graph.
+Application of both these algorithms to repeat detection problem provided significantly accurate and fast results.
+We explore how various parameters such as number of sampled nodes and filtering criteria to mark a node as repeat affects the accuracy of repeat detection.
+We are planning to use these techniques for repeat detection in our future metagenomic assembler.
 \bibliography{manuscript}{}

### 2015-05-06 runnode.sh
scripts fixed

@@ -5,6 +5,7 @@
 #nodes = $3
 #repeat_file = $4
 #map fie = $5
-./sampling_algo $1 $2 $3 $4
+#rt_file = $6
+./sampling_algo $1 $2 $3 $4 $6
 
-python evaluate1.py  $4 $3 $5
+python evaluate1.py  $4 $3 $5 $6

### 2015-05-06 sampling_algo.cpp

@@ -32,6 +32,7 @@ int main(int argc, char* argv[]) {
 	filterparam = atoi(argv[2]);
 	sampleNodes = atoi(argv[3]);
 	string repeat_file = argv[4];
+	string rt_file = argv[5];
 	clock_t start = clock();
 	GMLGraphReader reader;
 	Graph g = reader.read(argv[1]);
@@ -77,7 +78,7 @@ int main(int argc, char* argv[]) {
 	double bound = mean + filterparam*stdev;
 
 	ofstream runfile;
-        runfile.open(repeat_file+".runtime");
+        runfile.open(rt_file);
         runfile << (double) (clock() - start) / CLOCKS_PER_SEC << endl;
 
 	for (int i = 0; i < g.nodes().size(); i++) {

### 2015-05-06 vc_algo.cpp

@@ -33,6 +33,7 @@ int main(int argc, char* argv[]) {
 	string eps = argv[3];
 	double epsilon = (double)atof(eps.c_str());
 	string repeat_file = argv[4];
+	string rt_file = argv[5];
 	clock_t start = clock();
 	GMLGraphReader reader;
 	Graph g = reader.read(argv[1]);
@@ -84,7 +85,7 @@ int main(int argc, char* argv[]) {
 	cerr<<"filter: "<<filter<<endl;
 	double bound = mean + filterparam*stdev;
 	ofstream runfile;
-	runfile.open(repeat_file+".runtime");
+	runfile.open(rt_file);
 	runfile << (double) (clock() - start) / CLOCKS_PER_SEC << endl;
 
 	for (int i = 0; i < g.nodes().size(); i++) {

### 2015-05-06 manuscript/manuscript.tex
report modified

@@
 Genomic repeats are the major challenge when assembling isolated genomes and their effect is prominently seen in metagenomic datasets.
-It has been show by Kingsford et al.\cite{kingsford} that the number of reconstructions grow exponentially with the number of repeats.
+It has been shown by Kingsford et al.\cite{kingsford} that the number of reconstructions grow exponentially with the number of repeats.
 Therefore it is impossible find one correct solution for assembly.
@@
 
-Geisberger et al.
-\cite{sanders} have shown that $X := 2n\delta(v)$ is an unbiased estimator of centrality of v that is $E(X) := g(v)$.
-Therefore by averaging $k$ independent runs of this estimator, we get an approximation $\frac{X_{1}+X_{2}+..+X_{k}}{k}$ of the betweenness centrality for node $v$.
-
-To compute the contributions of nodes to centrality, a modified scheme of Brandes algorithm is used. Contributions are calculated as follows:
-
-$$\delta_{s}(v) = \sum_{w \in succ(v)} \frac{\mu(s,v)}{\mu(s,w)}.\frac{\sigma_{sv}}{\sigma_{sw}}(1+\delta_{s}(w))$$
-where $\mu(s,w)$ is the shortest path distance from $s$ to $w$. It uses $\frac{\mu(s,v)}{\mu(s,w)}$ in aggregation instead of $1$ in Brandes algorithm.
+This algorithm is efficiently implemented using two techniques.
+First one is \textit{linear scaling}, where the contribution of a node $v$ depends linearly on the distance to the sample.
+Nearer the node is to the sample, lesser is its contribution counted.
+Linear scaling is implemented using a simple variant of aggregation scheme proposed in Brandes algorithm.
+Second technique is \textit{bisection scaling} where the sample only contributes to the second half of a path.
+This means that for a vertex $v$, its contribution is considered only if it is at atleast half distance between $s$ and $t$ away from $s$ in forward search or more than half a distance away from $t$ in backward search.
 
@@
 VC dimension technique helps in choosing sample from $S_{all}$ so that it gives probabilistic guarantees.
-It is show that the upper bound on VC dimensions of $S_{all}$ is given by:
+It is shown that the upper bound on VC dimensions of $S_{all}$ is given by:
 $$VC(S_{all}) \leq \left\lfloor{log_{2}(VD(G) - 2)}\right\rfloor + 1$$

### 2015-05-06 manuscript/manuscript.tex
Fixed a few formatting issues.

@@
 \usepackage{url}
-\urldef{\mailsa}\path|{alfred.hofmann, ursula.barth, ingrid.haas, frank.holzwarth,|
-\urldef{\mailsb}\path|anna.kramer, leonie.kunz, christine.reiss, nicole.sator,|
-\urldef{\mailsc}\path|erika.siebert-cole, peter.strasser, lncs}@springer.com|    
+\urldef{\mailsa}\path|{jayg, cmhill, mpop}@cs.umd.edu|    
 \newcommand{\keywords}[1]{\par\addvspace\baselineskip
@@
 % first the title is needed
-\title{Identifiyng Inter-Genomic Repeats Using Fast, Approximate Betweenness Centrality}
+\title{Identifying Inter-Genomic Repeats Using Fast, Approximate Methods for Betweenness Centrality}
 
 % a short form should be given in case it is too long for the running head
-\titlerunning{Lecture Notes in Computer Science: Authors' Instructions}
+\titlerunning{Identifying Inter-Genomic Repeats Using Fast, Approximate Betweenness Centrality}
 
@@
 \author{Jay Ghurye%
-\and Chris Hill\and Mihai Pop}
+\and Christopher M. Hill\and Mihai Pop}
 %
 \authorrunning{Identifiyng Inter-Genomic Repeats Using Fast,
-Approximate Betweenness Centrality}
+Approximate Methods for Betweenness Centrality}
 % (feature abused for this document to repeat the title also on left hand pages)
@@
 \institute{University of Maryland, College Park\\
-\mailsa\\
-\mailsb\\
-\mailsc\\
-\url{http://www.springer.com/lncs}}
+\mailsa\\}
 
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics[width = \textwidth]{sampled_nodes}
+\caption{Example of conversion of de Bruijn graph to Repeats graph}
+\label{fig:sampled_nodes}
+\end{figure}
+
 \section{Conclusion}

### 2015-05-07 manuscript/manuscript.tex
changed results section

@@
 \item \textbf{Sample Size}: For the algorithm by Geisberger et al.
-\cite{sanders}, how should a sample size be defined so that it would produce fast and accurate results?
-We try to come up with a heuristic to define sample size
+\cite{sanders}, how should a sample size be defined so that accuray of repeat detection is maximized.
+We also stuy the effect of sample size in ghe algorithm by Matteo et al.\cite{matteo}
 \end{itemize}
@@
 We use the notion of sensitivity and specificity to define the accuracy of repeat detection.
-We define following terms:
+A good repeat detection strategy should have both high sensitivity and specificity.
+Sensitivity is the measure of proportion of actual repeats which are correctly identified by the algorithm.
+Specificity is the proportion of non-repeats correctly identified by the algorithm.
+High sensitivity is desired because we want to remove as many repeats from the graph.
+High specificity is desired becasue we do not want to remove non-repeats which are incorrectly identified as repeats by algorithm.
+To calculate sensitivity and specificity, we define following terms:
 \begin{itemize}
-\item \textbf{True positives:} Number of nodes marked as repeat that is found within at least 2 genomes
-\item \textbf{False positives:} Number of nodes marked as repeat that is only repetitive within a single genome
-\item \textbf{True negatives:} Number of nodes marked as non-repeat that is only repetitive within a single genome
-\item \textbf{False negatives:} Number of nodes marked as non-repeat that is found within at least 2 genomes
+\item \textbf{True positives(TP):} Number of nodes marked as repeat that is found within at least 2 genomes
+\item \textbf{False positives(FP):} Number of nodes marked as repeat that is only repetitive within a single genome
+\item \textbf{True negatives(TN):} Number of nodes marked as non-repeat that is only repetitive within a single genome
+\item \textbf{False negatives(FN):} Number of nodes marked as non-repeat that is found within at least 2 genomes
 \end{itemize}
 
-A good repeat detection should have both good sensitivity and specificity.
-Sensitivity reflects how many true repeats are detected.
-Detecting to few repeats can lead to assembly errors in scaffolding.
-Specificity reflects the false positives.
-Detecting too many repeats leads to a suboptimal assembly as these contigs do not fully participate in scaffolding.
+Using these quantities, we can define sensitivity and specificity as follows:
+ $$Sensitivity = \frac{TP}{TP+FN}$$
+ $$Specificity = \frac{TN}{TN+FP}$$
 

### 2015-05-07 manuscript/manuscript.tex
report changed

@@
 \section{Introduction}
-
-Metagenomics is a direct sequencing of DNA from all organisms in an environment without culturing.
-It is a culture-independent tool for studying environmental microorganisms.
-In addition to the information about taxonomic diversity (‘who is there’), metagenomics gives insight into the physiology of the organisms present in the environment (‘what are they doing’), through studying their genes.
-Two main goals to be achieved with metagenomics are finding new genes with desired biological activity (bioprospecting) and studying environmental microbes without the need to culture them.
 

### 2015-05-07 manuscript/manuscript.tex
report modified

@@
 We investigated the performance of approximate betweenness centrality algorithms on the metagenomic graphs to find inter-genome repeats.
-We used the implementation in Networkit\cite{networkit} library.
-It provides implementation for various graph algorithms.
-We are particularly interested in the implementation of betweenness centrality algorithms.
-We make use of two approximate betweenness centrality algorithms provided by NetworKit.
-
-NetworKit provides fast parallel implementation of approximate betweenness centrality algorithm by Matteo et al.
-\cite{matteo}.
-The implementation first approximates diameter of of graph.
-After this, it calculates the number of shortest paths to be sampled.
-It processes shortest paths in the sample in parallel in multithreaded fashion and then aggregates the centrality scores for each shortest paths once every thread finishes processing of shortest path assigned to it.
-
-We also investigate the performance of algorithm by Geisberger et al.
-\cite{sanders}.
-NetworKit provides single threaded implementation of this algorithm.
-It first creates a sample of $n$ nodes at random from a given graph.
-It then runs single source shortest path algorithm for each of those nodes.
-It calculates the betweenness scores for each node by the modified aggregation scheme proposed in the algorithm.
-
-Here are few questions we want to investigate based on these algorithms:
+We used the implementations provided by  Networkit\cite{networkit} library.
+NetworKit an open source library for high-performance network analysis.
+It implements effecient graph algorithms and parallelize most of them.
+It provides implementation for variety of centrality measures.
+We used the implementation of approximate network centrality.
+NetworKit provides a parallel implementation of the algorithm by Matteo et al.\cite{matteo} which is based on sampling shortest paths.
+It also provides implementation for the algorithm by Geisberger et al.\cite{sanders}.
+However, this algorithm is not parallelized.
 

### 2015-05-07 manuscript/manuscript.tex
Added sampled paths plot.
IQR added in cutoff

@@
 % a short form should be given in case it is too long for the running head
-\titlerunning{Identifying Inter-Genomic Repeats Using Fast, Approximate Betweenness Centrality}
+\titlerunning{Inter-Genomic Repeats Using Betweenness Centrality}
 
@@
 %
-\authorrunning{Identifiyng Inter-Genomic Repeats Using Fast,
-Approximate Methods for Betweenness Centrality}
+\authorrunning{Ghurye et al.}
 % (feature abused for this document to repeat the title also on left hand pages)
@@
 We used the implementations provided by  Networkit\cite{networkit} library.
-NetworKit an open source library for high-performance network analysis.
+NetworKit is an open source library for high-performance network analysis.
 It implements effecient graph algorithms and parallelize most of them.
 It provides implementation for variety of centrality measures.
-We used the implementation of approximate network centrality.
-NetworKit provides a parallel implementation of the algorithm by Matteo et al.\cite{matteo} which is based on sampling shortest paths.
+We used the implementation of approximate betweenness centrality.
+NetworKit provides parallelized implementation of the algorithm by Matteo et al.\cite{matteo} which is based on sampling shortest paths.
 It also provides implementation for the algorithm by Geisberger et al.\cite{sanders}.
 However, this algorithm is not parallelized.
-
-\begin{itemize}
-\item \textbf{Scalability}: How these algorithms scale on large metagenomic repeats graph? Does parallelized algorithm provides quick results or taking sample of fewer nodes provides quick results?
-\item \textbf{Accuracy}: Which of these algorithms provides better accuracy while determining inter-genomic repeats?
-\item \textbf{Sample Size}: For the algorithm by Geisberger et al.
-\cite{sanders}, how should a sample size be defined so that accuray of repeat detection is maximized.
-We also stuy the effect of sample size in ghe algorithm by Matteo et al.\cite{matteo}
-\end{itemize}
 
@@
 We set value of $c$ as 3 as it works out the best.
+
+Another cutoff criteria we test is based on \textit{interquartile range(IQR)}.
+Let $Q_{1}$,$Q_{2}$, and $Q_{3}$ denote lower, midle and upper quartile respectiely.
+IQR can be defined as the difference between the upper and lower quartile.
+So, $IQR = Q_{3} - Q_{1}$.
+For the values of centrality, we calculate upper quartile, middle quartile, lower quartile and IQR.
+For filtering, we mark all the nodes with centrality more than $Q_{2} + 1.5*IQR$ as repeats.
+The advantage of this method is that it does not make any assumptions about the distribution of underlying data.
 
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics[width = \textwidth]{sampled_paths}
+\caption{Example of conversion of de Bruijn graph to Repeats graph}
+\label{fig:sampled_nodes}
+\end{figure}
+
 \section{Conclusion}

### 2015-05-07 manuscript/manuscript.tex
minor bug fixed
Added example repeat graph
Added captions/table cutoff result.

@@
 
+\begin{figure*}[htbp]
+\centering
+\includegraphics[width=\textwidth]{es_mix_200kb_k21.png}
+\caption{Assembly graph of a simulated community consisting of 200 Kbp subsets of \textit{Escherichia coli} str.
+K-12 MG1655 and emph{Staphylococcus aureus}.
+Nodes are colored and sized based on their relative betweenness centrality.
+The highlighted nodes are inter-genomic repeats whose deletion would separate the graph.}
+\label{fig:sampled_nodes}
+\end{figure*}
+
 \subsection*{Betweenness Centrality Based on Sampling Shortest Paths}
@@
 Another cutoff criteria we test is based on \textit{interquartile range(IQR)}.
-Let $Q_{1}$,$Q_{2}$, and $Q_{3}$ denote lower, midle and upper quartile respectiely.
+Let $Q_{1}$,$Q_{2}$, and $Q_{3}$ denote lower, middle and upper quartile respectiely.
 IQR can be defined as the difference between the upper and lower quartile.
@@
 
+\begin{table}[h]
+\centering
+\caption[]{Statistical measures of performance using different cutoff levels for marking a node as an inter-genomic repeat on the simulated 10 genome assembly graph.
+$c=3$ is Bambus2's default cutoff level and \textsc{irq} utilizes the different quartile values.
+}
+\begin{tabular}{lllll}
+ & $c=3$        & \textsc{irq}   \\
+\hline
+sensitivity & 0.0584 & 0.344 \\
+specificity & 0.993  & 0.909 \\
+precision   & 0.250  & 0.125 \\
+\hline 
+\end{tabular}
+\end{table}
+
 \section{Conclusion}

### 2015-05-15 manuscript/manuscript.tex
Added some results.

@@
 It has been shown by Kingsford et al.\cite{kingsford} that the number of reconstructions grow exponentially with the number of repeats.
-Therefore it is impossible find one correct solution for assembly.
-There are several approaches such as \cite{bambus} to find repeats in metagenomic data.
+Therefore it is computationally infeasible to find one correct solution for assembly.
+There are several approaches, such as \cite{bambus}, to find repeats in metagenomic data.
 The main drawback of these methods is that they do not scale well as number of nodes in assembly graph increase and prove very inefficient as number of nodes reaches to about a million.
@@
 
-Brandes\cite{brandes} gives an exact algorithm for computing betweenness centrality of all the nodes that is based on solving a single source shortest path problem for each node.
+Brandes\cite{brandes} gives an exact algorithm for computing betweenness centrality of all the nodes that is based on solving a single source shortest path (SSSP) problem for each node.
 An SSSP computation from $s$ produces a directed acyclic graph (DAG) encoding all shortest paths starting at $s$.
@@
 \caption{Assembly graph of a simulated community consisting of 200 Kbp subsets of \textit{Escherichia coli} str.
-K-12 MG1655 and emph{Staphylococcus aureus}.
+K-12 MG1655 and \textit{Staphylococcus aureus}.
 Nodes are colored and sized based on their relative betweenness centrality.
@@
 
-\subsection*{De Bruijn Graph}
+\subsection*{Simulating Metagenomic Assembly Repeat Graphs}
 
-De Bruijn strings are a type of 'comprehensive' strings.
+We evaluate the efficacy of detecting inter-genomic repeats using the approximate betweenness centrality methods on simulated metagenomic assembly graphs.
+The first step is to construct the \textit{de Bruijn} graph from the set of reference genomes.
+De Bruijn strings are a type of ``comprehensive'' strings.
 For a given alphabet $\Sigma = \{A,C,T,G\}$ and a length $k$, a de Bruijn string of order $k$ contains as substrings of length $k$ over $\Sigma$.
@@
 
-We modify the de Bruijn graph generated from genomes to identify repeats in a better way.
-We define \textit{unipaths} as the path of nodes in which the outdegree of first node on the path is $1$ and indegree is greater than $1$, indegree of last node on the path is $1$ and outdegree is greater than $1$ and for all other nodes, both indegree and out degree is exactly $1$.
-We compress all the unipaths in a graph to represent in a compact way.
-For all the internal nodes on unipaths, we represent them by a single node with label as the combined label of all nodes.
-figure\ref{fig:repeatgraph} shows an example of such compression of unipaths in de Bruijn graphs.
+Next we modify the de Bruijn graph generated from genomes to identify repeats in a better way.
+We define \textit{unipaths} as the path of nodes in which the outdegree of first node on the path is $1$ and indegree is not equal to $1$, indegree of last node on the path is $1$ and outdegree is not equal to $1$ and for all other nodes along the path, both indegree and out degree is exactly $1$.
+We compress all the unipaths in a graph to represent the path in a compact way.
+We represent the unipath as a single node whose label is the concatenation of all node labels along the path.
+Nodes in the resulting graph represent uniquely resolvable sequences of length $>= k$.
+Figure~\ref{fig:repeatgraph} shows an example of such compression of unipaths in de Bruijn graphs.
 
@@
 
-\subsection*{Networkit}
-We investigated the performance of approximate betweenness centrality algorithms on the metagenomic graphs to find inter-genome repeats.
-We used the implementations provided by  Networkit\cite{networkit} library.
-NetworKit is an open source library for high-performance network analysis.
+For evaluation we simulate metagenomic assembly graphs consisting of two, five, and ten bacterial genomes.
+The composition of each simulated sample is described in Table \ref{tab:composition}.
+For the five and ten metagenome samples, random bacterial genomes were chosen from the data set provided by \cite{shakya2013comparative}.
+A $k$-mer size of 55 was chosen for initial creation of the de Bruijn graph.
+
+\begin{table}[h]
+\centering
+\caption[]{Composition of simulated metagenomic assembly graph.}
+\begin{tabular}{lllll}
+name & genomes & nodes & edges \\
+\hline
+2-genome & \textit{Escherichia coli} str. K-12, \textit{Staphylococcus aureus}  & 3,384 & 4,534 \\
+5-genome & \textit{Burkholderia xenovorans} LB400, \textit{Archaeoglobus fulgidus} DSM 4304, & 5,114 & 6,844 \\
+& \textit{Pyrobaculum calidifontis} JCM 11548, \textit{Zymomonas mobilis subsp. mobilis} ZM4, \\
+& \textit{Porphyromonas gingivalis} ATCC 33277  \\
+10-genome  & \textit{Caldicellulosiruptor bescii} DSM 6725, \textit{Bacteroides vulgatus} ATCC 8482, & 25,425 & 34,056\\
+& \textit{Bacteroides thetaiotaomicron} VPI-5482, \textit{Clostridium thermocellum} ATCC 27405, \\
+& \textit{Pyrobaculum arsenaticum} DSM 13514, \textit{Chlorobium phaeobacteroides} DSM 266, \\
+& \textit{Ruegeria pomeroyi} DSS-3, \textit{Treponema denticola} ATCC 35405, \\
+& \textit{Porphyromonas gingivalis} ATCC 33277, \textit{Nostoc sp.} PCC 7120  \\
+\hline 
+\end{tabular}
+\label{tab:composition}
+\end{table}
+
+\subsection*{\textsc{NetworKit}}
+We investigate the performance of approximate betweenness centrality algorithms on the metagenomic graphs to find inter-genome repeats.
+We used the implementations provided by \textsc{NetworKit}\cite{networkit} library.
+\textsc{NetworKit} is an open source library for high-performance network analysis.
 It implements effecient graph algorithms and parallelize most of them.
@@
 We used the implementation of approximate betweenness centrality.
-NetworKit provides parallelized implementation of the algorithm by Matteo et al.\cite{matteo} which is based on sampling shortest paths.
+\textsc{NetworKit} provides parallelized implementation of the algorithm by Matteo et al.\cite{matteo} which is based on sampling shortest paths.
 It also provides implementation for the algorithm by Geisberger et al.\cite{sanders}.
@@
 For the values of centrality, we calculate upper quartile, middle quartile, lower quartile and IQR.
-For filtering, we mark all the nodes with centrality more than $Q_{2} + 1.5*IQR$ as repeats.
+For filtering, we mark all the nodes with centrality more than $Q_{3} + 1.5*IQR$ as repeats.
 The advantage of this method is that it does not make any assumptions about the distribution of underlying data.
@@
 \section{Results} 
-We evaluated algorithms for betweenness centrality to find inter-genome repeats on real datasets described above.
-We use the notion of sensitivity and specificity to define the accuracy of repeat detection.
-A good repeat detection strategy should have both high sensitivity and specificity.
+We evaluate the effect of using approximation algorithms for betweenness centrality on finding inter-genome repeats using the simulated datasets described above.
+%We use the notion of Sensitivity, specificity, and precision to define the quality of repeat detection.
+A good repeat detection strategy should have high sensitivity, specificity, and precision.
 Sensitivity is the measure of proportion of actual repeats which are correctly identified by the algorithm.
 Specificity is the proportion of non-repeats correctly identified by the algorithm.
+Precision is the proportion of repeats identified that are correct.
 High sensitivity is desired because we want to remove as many repeats from the graph.
-High specificity is desired becasue we do not want to remove non-repeats which are incorrectly identified as repeats by algorithm.
-To calculate sensitivity and specificity, we define following terms:
+High specificity is desired because we do not want to remove non-repeats which are incorrectly identified as repeats by algorithm.
+High precision helps us determine the usefulness of our methods.
+To calculate sensitivity, specificity, and precision, we define following terms:
 \begin{itemize}
-\item \textbf{True positives(TP):} Number of nodes marked as repeat that is found within at least 2 genomes
-\item \textbf{False positives(FP):} Number of nodes marked as repeat that is only repetitive within a single genome
-\item \textbf{True negatives(TN):} Number of nodes marked as non-repeat that is only repetitive within a single genome
-\item \textbf{False negatives(FN):} Number of nodes marked as non-repeat that is found within at least 2 genomes
+\item \textbf{True positives(TP):} Number of nodes marked as a repeat that is found within at least 2 genomes
+\item \textbf{False positives(FP):} Number of nodes marked as a repeat that is only found within a single genome
+\item \textbf{True negatives(TN):} Number of nodes marked as a non-repeat that is only found within a single genome
+\item \textbf{False negatives(FN):} Number of nodes marked as a non-repeat that is found within at least 2 genomes
 \end{itemize}
@@
  $$Specificity = \frac{TN}{TN+FP}$$
+ $$Precision = \frac{TP}{TP + FP}$$
+
+For evaluating the effect of node sampling on inter-genomic repeat detection, we sample between 10 to 10,000 nodes per trial.
+For evaluating path sampling, we select epsilon parameters between 0.01 and 0.20.
+The result of each parameter is averaged over 10 trials.
+
+\subsection*{Approximate Betweenness Centrality Retains Detection Power of Inter-Genomic Repeats}
+
+Across all data sets, the more nodes and paths sampled, the better the inter-genomic repeat detection (Figures \ref{fig:sampled_nodes} and \ref{fig:sampled_paths}).
+Furthermore, the runtime increases linearly with the number of nodes sampled (Figure \ref{fig:sampled_nodes}).
+In the 2-genome assembly graph, all inter-genomic repeats were found by sampling as few as ten nodes, an order of magnitude less time than calculating the betweenness centrality using the full graph.
+In the 5-genome assembly graph, sampling 10 nodes has a sensitivity of 86\% and reaches 100\% at 500 nodes which is approximately one-tenth the size of the assembly graph.
+The sensitivity drops off drastically in the 10-genome assembly graph, oscillating around 6\% across all node sample sizes.
+While the specificity remains consistent across data sets (97\%-99\%), the precision increases from ~2.5\% to ~25\% when going from the 2-genome to the 10-genome data set.
+
+For the range of epsilon parameters used, the statistical measures of repeat detection remain constant (Figure \ref{fig:sampled_paths}).
+Similarly, the runtime for each epsilon parameter remains relatively unchanged except for a slight change in the 10-genome sample.
+For the 10-genome sample, the statistical measures and runtime are equivalent to when 100-500 nodes are sampled.
 
@@
 \includegraphics[width = \textwidth]{sampled_nodes}
-\caption{Example of conversion of de Bruijn graph to Repeats graph}
+\caption{Statistical measures of repeat detection quality for the simulated metagenomic assembly graphs. Betweenness centrality was approximated by randomly sampling nodes in the graph.}
 \label{fig:sampled_nodes}
@@
 \includegraphics[width = \textwidth]{sampled_paths}
-\caption{Example of conversion of de Bruijn graph to Repeats graph}
-\label{fig:sampled_nodes}
+\caption{Statistical measures of repeat detection quality for the simulated metagenomic assembly graphs. Betweenness centrality was approximated within epsilon of their true value.}
+\label{fig:sampled_paths}
 \end{figure}
+
+\subsection*{Alternate Betweenness Centrality Cutoff Increases Sensitivity}
+
+Decreasing the betweenness centrality cutoff requirement results in a trade-off between the different statistical measures.
+Instead of using the default cutoff of $3\sigma$ suggested by Bambus2, we mark all the nodes with centrality more than $Q_{3} + 1.5*IQR$ as an inter-genomic repeats.
+In the 10-genome assembly graph, the sensitivity increases from 5.8\% to 34.4\% when using the \textsc{iqr} cutoff method (Table \ref{tab:cutoffs}).
+Conversely, the specificity and and precision drops from 99.3\% to 90.9\% and 25\% to 12.5\%, respectively.
 
@@
 \caption[]{Statistical measures of performance using different cutoff levels for marking a node as an inter-genomic repeat on the simulated 10 genome assembly graph.
-$c=3$ is Bambus2's default cutoff level and \textsc{irq} utilizes the different quartile values.
-}
+$c=3$ is Bambus2's default cutoff level and \textsc{irq} utilizes the different quartile values.}
 \begin{tabular}{lllll}
- & $c=3$        & \textsc{irq}   \\
+ & $c=3$        & \textsc{iqr}   \\
 \hline
@@
 \end{tabular}
+\label{tab:cutoffs}
 \end{table}
+
+\section{Discussion}
+Betweenness centrality is a proven method for finding inter-genomic repeats in metagenomic assemblies.
+The size of typical metagenomic assembly graphs often make it infeasible to calculate the exact betweenness centrality of each node.
+We have taken the first steps to show how recent approximation algorithms for calculating betweenness centrality can be used retain the same repeat detection power as the exact algorithms in a fraction of the time.
+In the future, larger and more complex metagenomic assembly graphs must be used to further evaluate the efficacy of the approximation algorithms.
 
@@
 We explored two algorithms, one based on sampling shortest paths in a graph and other based on sampling nodes in a graph.
-Application of both these algorithms to repeat detection problem provided significantly accurate and fast results.
+Application of both these algorithms to repeat detection problem provided accurate and fast results.
 We explore how various parameters such as number of sampled nodes and filtering criteria to mark a node as repeat affects the accuracy of repeat detection.

### 2015-05-15 manuscript/manuscript.tex
description changed

@@
 \subsection*{Betweenness Centrality}
-In graph theory and network analysis, indicators of centrality identify the most important vertices within a graph.Several metrics to measure centrality have been proposed.
+In graph theory and network analysis, indicators of centrality identify the most important vertices within a graph.
+:Several metrics to measure centrality have been proposed.
 We make use of betweenness centrality.
@@
 An SSSP computation from $s$ produces a directed acyclic graph (DAG) encoding all shortest paths starting at $s$.
-By backward aggregation of counter values, the contributions of these paths to the betweenness counters can be computed in linear time.
-Brandes algorithm explains how to calculate $\sigma_{st}$ on the fly during shortest path calculations: $\sigma_{ss} = 1$ and for $s \neq t, \sigma_{st} = \sum_{v \in pred(t)} \sigma_{sv}$ where $pred(t)$ is a multiset containing the immediate predecessors of  $t$ in the shortest path DAG.
-In a subsequent aggregation phase, the nodes are processed in reverse topological order, i.e.
-by non-decreasing distance from $s$.
+The contributions of these paths to the betweenness counters can be computed in linear time using backward aggregation.
+Brandes algorithm explains how to calculate $\sigma_{st}$ on simultaneously with shortest path calculations: $\sigma_{ss} = 1$ and for $s \neq t, \sigma_{st} = \sum_{v \in pred(t)} \sigma_{sv}$ where $pred(t)$ is a collection of the immediate predecessors of  $t$ on every shortes path ending at $t$.
+In a subsequent aggregation phase, nodes are procesed by non-decreasing distance from $s$.
 So we have:
@@
 Bader and Madduri\cite{bader} provide a massively parallel implementation of the exact algorithm.
-However this approach scales to about a few million nodes.
+However this approach scales to about a few million nodes and can not be fast enough for larger assembly graphs.
 
@@
 Due to scalability limitations of exact betweenness centrality, several approximation algorithms for betweenness centrality have been proposed.
-Bader and Pich\cite{bp} provide an approximation algorithm by extrapolating from a subset of $k$ starting nodes called pivots and using same aggregation strategy as exact betweeness algorithm.
-However, this algorithm overestimates the centrality of some unimportant nodes which lie near pivot.
+Bader and Pich\cite{bp} provide an approximation algorithm by choosing a subset of $k$ starting nodes called pivots and using same aggregation strategy as exact betweeness centrality algorithm.
+However, this algorithm overestimates the centrality of some unimportant nodes which are nearer to pivots.
 The approximation algorithm proposed by Geisberger et al.
-\cite{sanders} comes over this drawback by changing the scheme for aggregating betweenness contributions so that nodes do not get profited just because they are near pivot.
-They define an estimator for betweenness centrality.
-This estimator is characterized by a length function $l : E \Rightarrow \mathbb{R}$ on the edges and a scaling function $f : [0,1] \Rightarrow [0,1]$.
-For a path $P = (e_{1},e_{2}...,e_{k})$ let $l(P) := \sum_{1 \leq i \leq k} l(e_{i})$.
-Let $n$ be the number of sample nodes chosen from a graph.
-The algorithm works as follows: In each iteration one of $2n$ shortest path searches is performed with uniform probability $1/2n$.
-An iteration can be a forward search in $G = (V,E)$ from a pivot $s \in V$ ($|V| = n$) or backward search from a pivot $t \in V$, that is a search from $t$ in $(V,{(v,u) : (u,v) \in E})$.
-For each shortest path of the form
-$$ P = (\overbrace{s,...,v}^{Q},..,t)$$ 
-a scaled contribution is defined as follows:\\
-
-\begin{equation*}
-	\delta_{P}(v) := \begin{cases}
-					\frac{f(l(Q)/l(P))}{\sigma_{st}} & \text{for a forward search} \\\\
-					\frac{1 - f(l(Q)/l(P))}{\sigma_{st}} & \text{for a backward search} 
-					\end{cases}
-\end{equation*}
-
-Overall, $v$ receives contribution
+\cite{sanders} comes over this drawback by changing the scheme for aggregating betweenness contributions so that nodes nearer to pivots are not unduly profited.
+The algorithm starts with the sample of $k$ nodes from $|V| = n$ nodes in graph.
+In each iteration, it chooses one of $2n$ shortest path at random.
+Here there are $2n$ sample shortest paths because algorithm can do forward search from $s \in V$ or backward search from $t \in V$ and each $s$ and $t$ can have $n$ values.
+Let $\delta_{P}(v)$ be the contribution received by node $v$ due to shortest path $P$.
+It can be shown that \cite{sanders} due to all shortes paths $P$ on which $v$ lies, $v$ receives contribution
 
@@
 First one is \textit{linear scaling}, where the contribution of a node $v$ depends linearly on the distance to the sample.
-Nearer the node is to the sample, lesser is its contribution counted.
+Nearer the node is to the sample, lesser is the contribution of that sample for the that node is counted.
 Linear scaling is implemented using a simple variant of aggregation scheme proposed in Brandes algorithm.
 Second technique is \textit{bisection scaling} where the sample only contributes to the second half of a path.
-This means that for a vertex $v$, its contribution is considered only if it is at atleast half distance between $s$ and $t$ away from $s$ in forward search or more than half a distance away from $t$ in backward search.
+This means that for a vertex $v$, the contribution of sample is considered only if it is at atleast half distance between $s$ and $t$ away from $s$ in forward search or more than half a distance away from $t$ in backward search where $s$ and $t$ are nodes in sample.
 
@@
 \subsection*{Betweenness Centrality Based on Sampling Shortest Paths}
-There  is another approximation algorithm was proposed by Matteo et al.
+Another approximation algorithm was proposed by Matteo et al.
 \cite{matteo} based on randomized sampling of shortest paths and offer probabilistic guarantees on the quality of approximation.
@@
 The main advantage of this algorithm is that the sample size does not depend on the size of the graph, but only on the maximum number vertices in a shortest path which is called \textit{vertex diameter}.
-
-The Vapnik-Chernovenkis(VC) dimensions of a class of subset is defined on a set of points is a measure of expressiveness of such class.
+The Vapnik-Chernovenkis(VC) dimensions of a class of subset is defined on a set of points is a representation of expressiveness of such class.
 Given a probability distribution on the set of points, a finite bound on the VC-dimension of the class of subsets implies a bound on the number of random samples required to approximate the probability of each subset in the class with its empirical average.
-This technique is used to derive the expressiveness of the sample of shortest paths derived.
+This technique is used to derive the expressiveness of the sample of shortest paths.
 For graph $G(V,E)$, let $S_{all}$ be the union of $T_{v}$ for all $v \in V$, where $T_{v}$ is the set of all shortest path to which $v$ is internal.
-VC dimension technique helps in choosing sample from $S_{all}$ so that it gives probabilistic guarantees.
+VC dimension technique helps in choosing appropriate sample from $S_{all}$.
 It is shown that the upper bound on VC dimensions of $S_{all}$ is given by:
 $$VC(S_{all}) \leq \left\lfloor{log_{2}(VD(G) - 2)}\right\rfloor + 1$$
-where $VD(G)$ is vertex diameter of $G$. Using this upper bound, sample size $r$ can be estimated as follows:
+where $VD(G)$ is vertex diameter of $G$. Using this upper bound, sample size $r$ can be derived as follows:
 \begin{equation} \label{sample_size_algo}
@@
 \end{equation} 
-where $c$ is a universal constant with value $0.5$.
+where $c$ is a constant with value $0.5$.
 It is guaranteed that betweenness centralities estimated using this sample size are within additive factor $\epsilon$ from real values with probability atleast $1-\delta$.
 
-Complete algorithm is sketched in algorithm 1.
-There are two crucial things in this algorithm.
-First is computing approximate vertex diameter.
+There are two crucial things needed to be handled.
+First is computing vertex diameter.
 Calculating exact VD(G) will need to solve all pair shortest  path problem.
 The complexity of this algorithm is $O(V^{2}logV + VE)$ by Johnson's algorithm\cite{johnson} and $\Theta(V^{3})$ by Floyd-Warshall algorithm.
-This would be very slow.
-A faster linear time 2-approximation algorithm is proposed in \cite{matteo} is used to calculate approximate vertex diameter.
+There exits a faster linear time 2-approximation algorithm to calculate approximate vertex diameter for unweighted graph.
+For weighted graph, there is no relationship between VD(G) and $diam$(G) as shortest path with maximum size may not have maximum weight.
+For weighted graphs, VD(G) is approximated by calculating the size of largest Weekly Connected Componet(WCC).
+This can be done in $O(m+n)$ using Breadth First Search.
 
 Second is to sample shortest paths.
-They way a random shortest path from $S_{uv}$ is inspired by Brandes algorithm\cite{brandes}.
-Assume that $u$ and $v$ are connected.
-For each vertex $w$, let $P_{u}(w)$ be a subset of neighbors of $w$ which are predecessors of $w$ along the shortest path from $u$ to $w$.
-Let $p^{*}$ be the sampled shortest path that we build backwards starting from the endpoint $v$ and by adding sampled predecessors before $v$.
-Initially $p^{*} = {v}$.
-Starting from $v$, we select it's predecessor $z \in P_{u}(v)$ using weighted random sampling.
-Each $z \in P_{u}(v)$ has probability $\sigma_{uz}/\sum_{w \in P_{u}(v)}\sigma_{uv} = \sigma_{uz}/\sigma_{uv}$.
-We add $z$ to $p^{*}$ and repeat this procedure until $z \neq u$.
-In the end, we have single shortest path sampled from $S_{uv}$.
-
-\begin{algorithm}
-\caption{Algorithm to compute approximate betweenness centrality}
-\begin {algorithmic}[1]
-\State {\textbf{Input}:Graph $G(V,E)$ with $|V| = n,\epsilon,\delta \in (0,1)$}
-\State {\textbf{Output}: A set of approximate betweenness centrality for all nodes $v$ in $G$}
-\For {$v \in V$}
-\State {$b'(v) = 0$}
-\EndFor 
-\State {$VD(G) \leftarrow \mathtt{getVertexDiameter}(G)$}
-\State {$r \leftarrow \frac{c}{\epsilon^{2}}(\left\lfloor{log_{2}(VD(G) - 2)}\right\rfloor + 1 + ln\frac{1}{\delta})$}
-\For{$i \leftarrow 1$ to $r$}
-\State {$(u,v) = \mathtt{SampleUniformVertexPair(V)}$}
-\State {$S_{uv} = \mathtt{ComputeAllShortestPath(u,v)}$}
-\If {$S_{uv} \neq \{p_{\phi}\}$}
-\State 	{$t \leftarrow v$}
-\While {$t \neq u$}
-\State {sample $z \in P_{u}(t)$ with probability $\sigma_{uz}/\sigma_{ut}$}
-\If {$z \neq u$}
-\State {$b'(z) \leftarrow b'(z) + 1/r$}
-\EndIf
-\State {$t \leftarrow z$}
-\EndWhile
-\EndIf
-\EndFor\\
-\Return {$(v,b'(v)) :\forall v \in V$}
-\end{algorithmic}
-\end{algorithm}
-
-Analysis of this algorithm shows that it offers probabilistic guarantees on the quality of all approximations of the betweenness centrality.
-Let $b(v)$ be the actual betweenness centrality of node $v$ and $b'(v)$ be the betweenness centrality returned by algorithm 1.
-Then following lemma holds:
-
-\begin{mylemma}
-With probability atleast $1-\delta$, all the approximations computed by the algorithm are within $\epsilon$ from their real value:
-$$Pr(\exists v \in V s.t. |b(v) - b'(v)| > \epsilon) < \delta$$
-\end{mylemma}
+Assume that two vertices $u$ and $v$ are connected.
+Shortest paths are sampled from backward using weighted random sampling.
+Let $S$ we the path we want to sample.
+Initially, $S = \{v\}$.
+For all nodes that are predecessors of $v$ on shortest paths from $u$ to $v$, each node will have different probability to get sampled.
+This probability can be calculated using method similar to aggregation method used in Brandes algorithm.
+We keep on adding nodes to $S$ until we hit the source node $u$.
+In this way a single shortest path is sampled from all shortest paths between $u$ and $v$.
 

### 2015-05-16 manuscript/manuscript.tex
modified tables

@@
 \usepackage{algorithm}
+\usepackage{tabularx}
 \usepackage[noend]{algpseudocode}
@@
 We propose an application of a core concept in social network analysis called betweenness centrality to metagenomic data to identify inter-genomic repeats.
-We study how approximation algorithms for betweenness centrality can be useful to find central nodes quickly and propose a novel approach to identify inter-genomic repeats accurately.
+We show how approximation algorithms for betweenness centrality can be useful to design a faster approach to identify inter-genomic repeats accurately.
 \keywords{Metagenomics, Algorithms, Centrality, Graph}
@@
 Therefore it is computationally infeasible to find one correct solution for assembly.
-There are several approaches, such as \cite{bambus}, to find repeats in metagenomic data.
-The main drawback of these methods is that they do not scale well as number of nodes in assembly graph increase and prove very inefficient as number of nodes reaches to about a million.
-To tackle this problem, we explore some of the work in social network analysis, particularly in the area of node centrality.
-We make use of betweenness centrality to find out repeats in assembly graph.
+To tackle this problem, several assemblers have used the notion of network centrality to find repeats.
+For example, Bambus2\cite{bambus} uses betweenness centrality to find central nodes in graphs.
+The drawback of their approach is that they calculate exact betweenness centrality by computing all pair shortest paths.
+This approach is very slow when applied to large assembly graphs.
+To overcome this drawback, we explore how can we apply approximate betweenness centrality to assembly graphs and find inter-genomic repeats accurately.
+The reason for using approximation algorithms is that they scale well as the size of graphs increase.
+We also investigate how parallelized implementation of these algorithms provide speedups in detecting repeats.
 
@@
 In graph theory and network analysis, indicators of centrality identify the most important vertices within a graph.
-:Several metrics to measure centrality have been proposed.
+Several metrics to measure centrality have been proposed.
 We make use of betweenness centrality.
@@
 \caption[]{Composition of simulated metagenomic assembly graph.}
-\begin{tabular}{lllll}
-name & genomes & nodes & edges \\
+\begin{tabularx}{\linewidth}{|l|X|l|l|}
+\hline
+\textbf{Name} & \textbf{Genomes} & \textbf{Nodes} & \textbf{Edges} \\
 \hline
 2-genome & \textit{Escherichia coli} str. K-12, \textit{Staphylococcus aureus}  & 3,384 & 4,534 \\
-5-genome & \textit{Burkholderia xenovorans} LB400, \textit{Archaeoglobus fulgidus} DSM 4304, & 5,114 & 6,844 \\
-& \textit{Pyrobaculum calidifontis} JCM 11548, \textit{Zymomonas mobilis subsp. mobilis} ZM4, \\
-& \textit{Porphyromonas gingivalis} ATCC 33277  \\
-10-genome  & \textit{Caldicellulosiruptor bescii} DSM 6725, \textit{Bacteroides vulgatus} ATCC 8482, & 25,425 & 34,056\\
-& \textit{Bacteroides thetaiotaomicron} VPI-5482, \textit{Clostridium thermocellum} ATCC 27405, \\
-& \textit{Pyrobaculum arsenaticum} DSM 13514, \textit{Chlorobium phaeobacteroides} DSM 266, \\
-& \textit{Ruegeria pomeroyi} DSS-3, \textit{Treponema denticola} ATCC 35405, \\
-& \textit{Porphyromonas gingivalis} ATCC 33277, \textit{Nostoc sp.} PCC 7120  \\
 \hline
-\end{tabular}
+5-genome & \textit{Burkholderia xenovorans} LB400, \textit{Archaeoglobus fulgidus} DSM 4304,  \textit{Pyrobaculum calidifontis} JCM 11548, \textit{Zymomonas mobilis subsp.
+mobilis} ZM4, \textit{Porphyromonas gingivalis} ATCC 33277 & 5,114 & 6,844  \\
+\hline
+10-genome  & \textit{Caldicellulosiruptor bescii} DSM 6725, \textit{Bacteroides vulgatus} ATCC 8482, 
+ \textit{Bacteroides thetaiotaomicron} VPI-5482, \textit{Clostridium thermocellum} ATCC 27405, 
+ \textit{Pyrobaculum arsenaticum} DSM 13514, \textit{Chlorobium phaeobacteroides} DSM 266, 
+ \textit{Ruegeria pomeroyi} DSS-3, \textit{Treponema denticola} ATCC 35405, 
+ \textit{Porphyromonas gingivalis} ATCC 33277, \textit{Nostoc sp.} PCC 7120  & 25,425 & 34,056\\
+\hline 
+\end{tabularx}
 \label{tab:composition}
@@
 We used the implementation of approximate betweenness centrality.
-\textsc{NetworKit} provides parallelized implementation of the algorithm by Matteo et al.\cite{matteo} which is based on sampling shortest paths.
-It also provides implementation for the algorithm by Geisberger et al.\cite{sanders}.
-However, this algorithm is not parallelized.
+For our experiments, we use the parallelized implementation for the algorithm by Matteo et al.\cite{matteo} and serial implementation for the algorithm by Geisberger et al.\cite{sanders}.
 
@@
 $c=3$ is Bambus2's default cutoff level and \textsc{irq} utilizes the different quartile values.}
-\begin{tabular}{lllll}
- & $c=3$        & \textsc{iqr}   \\
+\begin{tabular}{|l|l|l|}
+\hline
+Measure & $c=3$ & \textsc{iqr}   \\
 \hline

### 2015-05-17 manuscript/manuscript.tex
updated few changes suggested by Mihai

@@
 Genomic repeats are the major challenge when assembling isolated genomes and their effect is prominently seen in metagenomic datasets.
-It has been shown by Kingsford et al.\cite{kingsford} that the number of reconstructions grow exponentially with the number of repeats.
+It has been shown that the number of reconstructions grow exponentially with the number of repeats\cite{kingsford}.
 Therefore it is computationally infeasible to find one correct solution for assembly.
@@
 The contributions of these paths to the betweenness counters can be computed in linear time using backward aggregation.
-Brandes algorithm explains how to calculate $\sigma_{st}$ on simultaneously with shortest path calculations: $\sigma_{ss} = 1$ and for $s \neq t, \sigma_{st} = \sum_{v \in pred(t)} \sigma_{sv}$ where $pred(t)$ is a collection of the immediate predecessors of  $t$ on every shortes path ending at $t$.
+Brandes algorithm calculates $\sigma_{st}$  simultaneously with shortest path calculations: $\sigma_{ss} = 1$ and for $s \neq t, \sigma_{st} = \sum_{v \in pred(t)} \sigma_{sv}$ where $pred(t)$ is a collection of the immediate predecessors of  $t$ on every shortes path ending at $t$.
 In a subsequent aggregation phase, nodes are procesed by non-decreasing distance from $s$.
@@
 Bader and Madduri\cite{bader} provide a massively parallel implementation of the exact algorithm.
-However this approach scales to about a few million nodes and can not be fast enough for larger assembly graphs.
+However this approach scales to about a few million nodes and may not be fast enough for larger assembly graphs.
 
@@
 The approximation algorithm proposed by Geisberger et al.
-\cite{sanders} comes over this drawback by changing the scheme for aggregating betweenness contributions so that nodes nearer to pivots are not unduly profited.
+\cite{sanders} overcomes this drawback by changing the scheme for aggregating betweenness contributions so that nodes nearer to pivots are not unduly profited.
 The algorithm starts with the sample of $k$ nodes from $|V| = n$ nodes in graph.
@@
 Let $\delta_{P}(v)$ be the contribution received by node $v$ due to shortest path $P$.
-It can be shown that \cite{sanders} due to all shortes paths $P$ on which $v$ lies, $v$ receives contribution
+It can be shown that \cite{sanders} due to all shortest paths $P$ on which $v$ lies, $v$ receives contribution
 
@@
 
-This algorithm is efficiently implemented using two techniques.
-First one is \textit{linear scaling}, where the contribution of a node $v$ depends linearly on the distance to the sample.
-Nearer the node is to the sample, lesser is the contribution of that sample for the that node is counted.
-Linear scaling is implemented using a simple variant of aggregation scheme proposed in Brandes algorithm.
-Second technique is \textit{bisection scaling} where the sample only contributes to the second half of a path.
-This means that for a vertex $v$, the contribution of sample is considered only if it is at atleast half distance between $s$ and $t$ away from $s$ in forward search or more than half a distance away from $t$ in backward search where $s$ and $t$ are nodes in sample.
+This algorithm is efficiently implemented by modifying original aggregation scheme of Brandes algorithm so that the nodes near the sampled node will not receive high centrality score.
+Runtime of the algorithm largely depends on the size of the sample we choose.
+However,there is a tradeoff between size of sample and accuracy of centrality detection.
 
@@
 The main advantage of this algorithm is that the sample size does not depend on the size of the graph, but only on the maximum number vertices in a shortest path which is called \textit{vertex diameter}.
-The Vapnik-Chernovenkis(VC) dimensions of a class of subset is defined on a set of points is a representation of expressiveness of such class.
-Given a probability distribution on the set of points, a finite bound on the VC-dimension of the class of subsets implies a bound on the number of random samples required to approximate the probability of each subset in the class with its empirical average.
-This technique is used to derive the expressiveness of the sample of shortest paths.
-For graph $G(V,E)$, let $S_{all}$ be the union of $T_{v}$ for all $v \in V$, where $T_{v}$ is the set of all shortest path to which $v$ is internal.
-VC dimension technique helps in choosing appropriate sample from $S_{all}$.
-It is shown that the upper bound on VC dimensions of $S_{all}$ is given by:
-$$VC(S_{all}) \leq \left\lfloor{log_{2}(VD(G) - 2)}\right\rfloor + 1$$
-where $VD(G)$ is vertex diameter of $G$. Using this upper bound, sample size $r$ can be derived as follows:
+The sample size $r$ of shortest paths is found out with the help of $VD(G)$ as follows:
 \begin{equation} \label{sample_size_algo}
@@
 
-There are two crucial things needed to be handled.
-First is computing vertex diameter.
-Calculating exact VD(G) will need to solve all pair shortest  path problem.
-The complexity of this algorithm is $O(V^{2}logV + VE)$ by Johnson's algorithm\cite{johnson} and $\Theta(V^{3})$ by Floyd-Warshall algorithm.
-There exits a faster linear time 2-approximation algorithm to calculate approximate vertex diameter for unweighted graph.
-For weighted graph, there is no relationship between VD(G) and $diam$(G) as shortest path with maximum size may not have maximum weight.
-For weighted graphs, VD(G) is approximated by calculating the size of largest Weekly Connected Componet(WCC).
-This can be done in $O(m+n)$ using Breadth First Search.
-
-Second is to sample shortest paths.
-Assume that two vertices $u$ and $v$ are connected.
-Shortest paths are sampled from backward using weighted random sampling.
-Let $S$ we the path we want to sample.
-Initially, $S = \{v\}$.
-For all nodes that are predecessors of $v$ on shortest paths from $u$ to $v$, each node will have different probability to get sampled.
-This probability can be calculated using method similar to aggregation method used in Brandes algorithm.
-We keep on adding nodes to $S$ until we hit the source node $u$.
-In this way a single shortest path is sampled from all shortest paths between $u$ and $v$.
-
-The runtime of this algorithm is dominated by the computation of shortest path at each step which takes time $O(|V| + |E|)$ if graph is unweighted (BFS) and $O(|E| + |V|log|V|)$ othewise (Dijkstra's algorithm).
+The algorithm estimates the vertex diameter of the graph using a simple $2-$approximation algorithm.
+The shortest paths are sampled using weighted random sampling.
+The runtime of this algorithm is dominated by the computation of shortest path at each step which takes time $O(|V| + |E|)$ if graph is unweighted (BFS) and $O(|E| + |V|log|V|)$ otherwise (Dijkstra's algorithm).
 This is multiplied by $r$ for $r$ iterations to obtain final complexity.
@@
 
-Next we modify the de Bruijn graph generated from genomes to identify repeats in a better way.
+Next we modify the de Bruijn graph generated from genomes to identify repeats.
 We define \textit{unipaths} as the path of nodes in which the outdegree of first node on the path is $1$ and indegree is not equal to $1$, indegree of last node on the path is $1$ and outdegree is not equal to $1$ and for all other nodes along the path, both indegree and out degree is exactly $1$.
@@
 Nodes in the resulting graph represent uniquely resolvable sequences of length $>= k$.
-Figure~\ref{fig:repeatgraph} shows an example of such compression of unipaths in de Bruijn graphs.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width = \textwidth]{repeatgraph}
-\caption{Example of conversion of de Bruijn graph to Repeats graph}
-\label{fig:repeatgraph}
-\end{figure}
 
@@
 \subsection*{Cutoff Criteria For Repeats}
-To determine which nodes in a graph should be marked as repeats, we define a criteria based on the value of centrality calculated by the algorithm.
-Let $x$ be the mean of centralities of all the nodes and $\sigma$ be the standard deviation.
-A node $v$ is marked as repeat only if the following condition holds: $b(v) \geq x + c*\sigma$ where $c$ is a constant and $b(v)$ is betweenness centrality of the node.
-We set value of $c$ as 3 as it works out the best.
+Bambus2\cite{bambus} uses exact betweenness centrality to find repeats.
+To filter out repeats, it calculates mean $\mu$ and standard deviation $\sigma$ over the centrality scores for all the nodes in graph.
+It then marks all the nodes with centrality score greater than $\mu + 3*\sigma$ as repeats.
+To filter out repeats based on approximate centrality scores, we use the same method as Bambus2.
 
@@
 Instead of using the default cutoff of $3\sigma$ suggested by Bambus2, we mark all the nodes with centrality more than $Q_{3} + 1.5*IQR$ as an inter-genomic repeats.
-In the 10-genome assembly graph, the sensitivity increases from 5.8\% to 34.4\% when using the \textsc{iqr} cutoff method (Table \ref{tab:cutoffs}).
+In the 10-genome assembly graph, the sensitivity increases from 5.8\% to 34.4\% when using the \textsc{iqr} cutoff method.
 Conversely, the specificity and and precision drops from 99.3\% to 90.9\% and 25\% to 12.5\%, respectively.
-
-\begin{table}[h]
-\centering
-\caption[]{Statistical measures of performance using different cutoff levels for marking a node as an inter-genomic repeat on the simulated 10 genome assembly graph.
-$c=3$ is Bambus2's default cutoff level and \textsc{irq} utilizes the different quartile values.}
-\begin{tabular}{|l|l|l|}
-\hline
-Measure & $c=3$ & \textsc{iqr}   \\
-\hline
-sensitivity & 0.0584 & 0.344 \\
-specificity & 0.993  & 0.909 \\
-precision   & 0.250  & 0.125 \\
-\hline 
-\end{tabular}
-\label{tab:cutoffs}
-\end{table}
 

### 2015-05-17 manuscript/manuscript.tex
benchmakrs added:

@@
 \usepackage{tabularx}
+
 \usepackage[noend]{algpseudocode}
@@
 
+\subsection*{Benchmarking Runtime of Sampling Algorithms}
+We test how the runtime of both the algorithms change as we change size of the sample.
+We use the k-mer graph generated from the genome of \textit{Escherichia coli} str.
+K-12 with 311137 nodes and 279290 edges.
+For the algorithm based on node sampling, we tested runtime on the sample sizes 10, 100, 1000, 10000 and 50000.
+It is observed that runtime increases linearly with sample size(Figure \ref{fig:bm_nodes}).
+For the algorithm based on shortest path sampling, we varied the value of $\epsilon$ parameter which controls the sample size.
+Lesser the $\epsilon$ is, higher is the sample size.
+It is observed that as we increased epsilon, runtime decreased(Figure \ref{fig:bm_paths}).
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=8cm,height=5cm]{benchmarknodes}
+\caption{Plot of runtime against number of sampled nodes. It can be seen that there is a linear relation between these two.}
+\label{fig:bm_nodes}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=8cm,height=5cm]{benchmarkpaths}
+\caption{Plot of runtime against shortest path sample size. It can be seen that runtime increases with higher size sample.}
+\label{fig:bm_paths}
+\end{figure}
+
 \subsection*{Approximate Betweenness Centrality Retains Detection Power of Inter-Genomic Repeats}

### 2015-05-18 manuscript/manuscript.tex
Revised intro.

@@
 
-Metagenomic data are considerably more complex.
-The assembly of metagenomic data is complicated due to several reasons.
-Few of them are: (i) widely different levels of representation for different organisms in a community; (ii) genomic variation between closely related organisms; (iii) conserved genomic regions shared by distantly related organims; and (iv) repetitive sequences within individual genomes.
-Genomic repeats are the major challenge when assembling isolated genomes and their effect is prominently seen in metagenomic datasets.
-It has been shown that the number of reconstructions grow exponentially with the number of repeats\cite{kingsford}.
-Therefore it is computationally infeasible to find one correct solution for assembly.
-To tackle this problem, several assemblers have used the notion of network centrality to find repeats.
-For example, Bambus2\cite{bambus} uses betweenness centrality to find central nodes in graphs.
-The drawback of their approach is that they calculate exact betweenness centrality by computing all pair shortest paths.
-This approach is very slow when applied to large assembly graphs.
-To overcome this drawback, we explore how can we apply approximate betweenness centrality to assembly graphs and find inter-genomic repeats accurately.
-The reason for using approximation algorithms is that they scale well as the size of graphs increase.
-We also investigate how parallelized implementation of these algorithms provide speedups in detecting repeats.
+The assembly of metagenomics data, the sequences of DNA recovered from an environment without culturing, is considerably more complex than that of individual genomes.
+%Challenges include: (i) repetitive sequences within individual genomes; (ii) varying abundances of different organisms in the community; (iii) variation between closely related organisms; and (iv) conserved genomic segments shared by distantly related organisms.
+Repetitive sequences pose a major challenge when assembling isolated genomes.
+It has been shown that the number of reconstructions grow exponentially with the number of repeats making it computationally infeasible to find one correct solution for assembly\cite{kingsford}.
+The effect of repeats on assembly is exacerbated in metagenomic data sets due to the presence of conserved regions of the genome linking together distantly related organisms.
+To tackle the problem of identifying these inter-genomic repeats, several assemblers have used the notion of network centrality to find repeats within the assembly graph.
+For example, Bambus2\cite{bambus} uses betweenness centrality to find \emph{central} nodes in the graph.
+The drawback of their approach is the expensive calculation of betweenness centrality by computing all-pair shortest paths.
+This computation is impractical when applied to the large assembly graphs typically encountered.
+To overcome this drawback, we explore the application of approximate and parallel betweenness centrality algorithms on metagenomic assembly graphs and evaluate their effectiveness in identifying inter-genomic repeats.
 
@@
 Nodes are colored and sized based on their relative betweenness centrality.
-The highlighted nodes are inter-genomic repeats whose deletion would separate the graph.}
+The highlighted nodes are inter-genomic repeats whose deletion would separate the graph.
+Note that the betweenness centrality measure correctly identifies these nodes.}
 \label{fig:sampled_nodes}
@@
 Another cutoff criteria we test is based on \textit{interquartile range(IQR)}.
-Let $Q_{1}$,$Q_{2}$, and $Q_{3}$ denote lower, middle and upper quartile respectiely.
+Let $Q_{1}$,$Q_{2}$, and $Q_{3}$ denote the lower, middle and upper quartiles respectively.
 IQR can be defined as the difference between the upper and lower quartile.
@@
 Instead of using the default cutoff of $3\sigma$ suggested by Bambus2, we mark all the nodes with centrality more than $Q_{3} + 1.5*IQR$ as an inter-genomic repeats.
-In the 10-genome assembly graph, the sensitivity increases from 5.8\% to 34.4\% when using the \textsc{iqr} cutoff method.
+In the 10-genome assembly graph, the sensitivity increases from 5.8\% to 34.4\% when using the IQR cutoff method.
 Conversely, the specificity and and precision drops from 99.3\% to 90.9\% and 25\% to 12.5\%, respectively.

### 2015-05-18 manuscript/manuscript.tex
removed benchmark

--- 2015-05-17 manuscript/manuscript.tex
@@
 
+Apart from simulated metagenomic graphs, we study how approximation algorithms for betweenness centrality work on real datasets.
+We evaluated the runtime  of these algorithms on metagenomic stool sample of a premature infant.
+This graph has 3797425 nodes and 2970379 edges.
+We compared the performance of Bambus2's repeat detection module on this data with the performance of approximate betweenness centrality algorithms.
+
 \subsection*{\textsc{NetworKit}}
@@
 
+\subsection*{Results On Real Metagenomic Datasets}
+We evaluated the runtime of Bambus2's repeat detection module on metagonic stool sample. It
+
 \section{Discussion}

### 2015-05-18 manuscript/manuscript.tex

--- 2015-05-18 manuscript/manuscript.tex
@@
 
-\subsection*{Benchmarking Runtime of Sampling Algorithms}
-We test how the runtime of both the algorithms change as we change size of the sample.
-We use the k-mer graph generated from the genome of \textit{Escherichia coli} str.
-K-12 with 311137 nodes and 279290 edges.
-For the algorithm based on node sampling, we tested runtime on the sample sizes 10, 100, 1000, 10000 and 50000.
-It is observed that runtime increases linearly with sample size(Figure \ref{fig:bm_nodes}).
-For the algorithm based on shortest path sampling, we varied the value of $\epsilon$ parameter which controls the sample size.
-Lesser the $\epsilon$ is, higher is the sample size.
-It is observed that as we increased epsilon, runtime decreased(Figure \ref{fig:bm_paths}).
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=8cm,height=5cm]{benchmarknodes}
-\caption{Plot of runtime against number of sampled nodes. It can be seen that there is a linear relation between these two.}
-\label{fig:bm_nodes}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=8cm,height=5cm]{benchmarkpaths}
-\caption{Plot of runtime against shortest path sample size. It can be seen that runtime increases with higher size sample.}
-\label{fig:bm_paths}
-\end{figure}
-
 \subsection*{Approximate Betweenness Centrality Retains Detection Power of Inter-Genomic Repeats}

### 2015-05-18 manuscript/manuscript.tex
some changes:

@@
 Due to scalability limitations of exact betweenness centrality, several approximation algorithms for betweenness centrality have been proposed.
-Bader and Pich\cite{bp} provide an approximation algorithm by choosing a subset of $k$ starting nodes called pivots and using same aggregation strategy as exact betweeness centrality algorithm.
+Bader and Pich\cite{bp} provide an approximation algorithm by choosing a subset of $k$ starting nodes called pivots and using same aggregation strategy as Brandes algorithm.
 However, this algorithm overestimates the centrality of some unimportant nodes which are nearer to pivots.
@@
 where $c$ is a constant with value $0.5$.
-It is guaranteed that betweenness centralities estimated using this sample size are within additive factor $\epsilon$ from real values with probability atleast $1-\delta$.
 
@@
 The shortest paths are sampled using weighted random sampling.
-The runtime of this algorithm is dominated by the computation of shortest path at each step which takes time $O(|V| + |E|)$ if graph is unweighted (BFS) and $O(|E| + |V|log|V|)$ otherwise (Dijkstra's algorithm).
+The runtime of this algorithm is dominated by the computation of shortest path at each step which takes time $O(|V| + |E|)$ for unweighted graphs(BFS) and $O(|E| + |V|log|V|)$ otherwise (Dijkstra's algorithm).
 This is multiplied by $r$ for $r$ iterations to obtain final complexity.
@@
 Apart from simulated metagenomic graphs, we study how approximation algorithms for betweenness centrality work on real datasets.
-We evaluated the runtime  of these algorithms on metagenomic stool sample of a premature infant.
+We evaluated the runtime  of these algorithms on metagenomic stool sample of a premature infant\cite{morowitz2011strain}.
 This graph has 3797425 nodes and 2970379 edges.
@@
 
-\subsection*{\textsc{NetworKit}}
+\subsection*{Implementation}
 We investigate the performance of approximate betweenness centrality algorithms on the metagenomic graphs to find inter-genome repeats.

### 2015-05-18 manuscript/manuscript.tex
Cleaned up b.c. section.

--- 2015-05-18 manuscript/manuscript.tex
@@
 The effect of repeats on assembly is exacerbated in metagenomic data sets due to the presence of conserved regions of the genome linking together distantly related organisms.
-To tackle the problem of identifying these inter-genomic repeats, several assemblers have used the notion of network centrality to find repeats within the assembly graph.
+To tackle the problem of identifying these inter-genomic repeats, several assemblers have used the notion of network centrality to find repeats within the assembly graph (Figure \ref{fig:es_mix}).
 For example, Bambus2\cite{bambus} uses betweenness centrality to find \emph{central} nodes in the graph.
@@
 
+\begin{figure*}[htbp!]
+\centering
+\includegraphics[width=0.60\textwidth]{es_mix_200kb_k21.png}
+\caption{Assembly graph of a simulated community consisting of 200 Kbp subsets of \textit{Escherichia coli} str.
+K-12 MG1655 and \textit{Staphylococcus aureus}.
+Nodes are colored and sized based on their relative betweenness centrality with larger, green nodes indicating a higher centrality.
+The highlighted nodes are inter-genomic repeats whose deletion would separate the graph.
+Note that the betweenness centrality measure correctly identifies these nodes.}
+\label{fig:es_mix}
+\end{figure*}
+
 \section{Related Work}
@@
 \subsection*{Betweenness Centrality}
-In graph theory and network analysis, indicators of centrality identify the most important vertices within a graph.
-Several metrics to measure centrality have been proposed.
-We make use of betweenness centrality.
-For a particular node, betweenness centrality is equal to the number of shortest paths from all vertices to all others that pass through that node.
+In network analysis, metrics of centrality are used to identify the most important nodes within a graph.
+Several metrics to measure centrality have been proposed, but in this work, we use betweenness centrality.
+The betweenness centrality of a particular node is equal to the number of shortest paths from all nodes to all others that pass through that node.
 Formally, for a node $v$, it is defined by following expression:
@@
 
-Brandes\cite{brandes} gives an exact algorithm for computing betweenness centrality of all the nodes that is based on solving a single source shortest path (SSSP) problem for each node.
-An SSSP computation from $s$ produces a directed acyclic graph (DAG) encoding all shortest paths starting at $s$.
+Brandes\cite{brandes} gives an exact algorithm for computing betweenness centrality of all the nodes that is based on solving a single-source shortest path (SSSP) problem for each node.
+A SSSP computation from $s$ produces a directed acyclic graph (DAG) encoding all shortest paths starting at $s$.
 The contributions of these paths to the betweenness counters can be computed in linear time using backward aggregation.
 Brandes algorithm calculates $\sigma_{st}$  simultaneously with shortest path calculations: $\sigma_{ss} = 1$ and for $s \neq t, \sigma_{st} = \sum_{v \in pred(t)} \sigma_{sv}$ where $pred(t)$ is a collection of the immediate predecessors of  $t$ on every shortes path ending at $t$.
-In a subsequent aggregation phase, nodes are procesed by non-decreasing distance from $s$.
+In a subsequent aggregation phase, nodes are processed by non-decreasing distance from $s$.
 So we have:
@@
 This algorithm does not scale well with large networks with millions of nodes and edges.
-Bader and Madduri\cite{bader} provide a massively parallel implementation of the exact algorithm.
-However this approach scales to about a few million nodes and may not be fast enough for larger assembly graphs.
+Bader and Madduri\cite{bader} provide a massively parallel implementation of the exact algorithm, but scales in practice to only a few million nodes, limiting its application to larger metagenomic assembly graphs that exceed million of nodes.
 
@@
 Due to scalability limitations of exact betweenness centrality, several approximation algorithms for betweenness centrality have been proposed.
-Bader and Pich\cite{bp} provide an approximation algorithm by choosing a subset of $k$ starting nodes called pivots and using same aggregation strategy as exact betweeness centrality algorithm.
+Bader and Pich\cite{bp} provide an approximation algorithm by choosing a subset of $k$ starting nodes called pivots and using same aggregation strategy as exact betweenness centrality algorithm.
 However, this algorithm overestimates the centrality of some unimportant nodes which are nearer to pivots.
@@
 However,there is a tradeoff between size of sample and accuracy of centrality detection.
-
-\begin{figure*}[htbp]
-\centering
-\includegraphics[width=\textwidth]{es_mix_200kb_k21.png}
-\caption{Assembly graph of a simulated community consisting of 200 Kbp subsets of \textit{Escherichia coli} str.
-K-12 MG1655 and \textit{Staphylococcus aureus}.
-Nodes are colored and sized based on their relative betweenness centrality.
-The highlighted nodes are inter-genomic repeats whose deletion would separate the graph.
-Note that the betweenness centrality measure correctly identifies these nodes.}
-\label{fig:sampled_nodes}
-\end{figure*}
 
@@
 
-Apart from simulated metagenomic graphs, we study how approximation algorithms for betweenness centrality work on real datasets.
-We evaluated the runtime  of these algorithms on metagenomic stool sample of a premature infant.
-This graph has 3797425 nodes and 2970379 edges.
-We compared the performance of Bambus2's repeat detection module on this data with the performance of approximate betweenness centrality algorithms.
-
 \subsection*{\textsc{NetworKit}}
@@
 
-\subsection*{Results On Real Metagenomic Datasets}
-We evaluated the runtime of Bambus2's repeat detection module on metagonic stool sample. It
-
 \section{Discussion}

### 2015-05-18 manuscript/manuscript.tex
changed implementation
Some changes

--- 2015-05-18 manuscript/manuscript.tex
@@
 
-\subsection*{Implementation}
-We investigate the performance of approximate betweenness centrality algorithms on the metagenomic graphs to find inter-genome repeats.
-We used the implementations provided by \textsc{NetworKit}\cite{networkit} library.
-\textsc{NetworKit} is an open source library for high-performance network analysis.
-It implements effecient graph algorithms and parallelize most of them.
-It provides implementation for variety of centrality measures.
-We used the implementation of approximate betweenness centrality.
-For our experiments, we use the parallelized implementation for the algorithm by Matteo et al.\cite{matteo} and serial implementation for the algorithm by Geisberger et al.\cite{sanders}.
-
-\subsection*{Cutoff Criteria For Repeats}
-Bambus2\cite{bambus} uses exact betweenness centrality to find repeats.
-To filter out repeats, it calculates mean $\mu$ and standard deviation $\sigma$ over the centrality scores for all the nodes in graph.
-It then marks all the nodes with centrality score greater than $\mu + 3*\sigma$ as repeats.
-To filter out repeats based on approximate centrality scores, we use the same method as Bambus2.
-
-Another cutoff criteria we test is based on \textit{interquartile range(IQR)}.
-Let $Q_{1}$,$Q_{2}$, and $Q_{3}$ denote the lower, middle and upper quartiles respectively.
-IQR can be defined as the difference between the upper and lower quartile.
-So, $IQR = Q_{3} - Q_{1}$.
-For the values of centrality, we calculate upper quartile, middle quartile, lower quartile and IQR.
-For filtering, we mark all the nodes with centrality more than $Q_{3} + 1.5*IQR$ as repeats.
-The advantage of this method is that it does not make any assumptions about the distribution of underlying data.
-
-\section{Results} 
+\subsection*{Evaluation}
 We evaluate the effect of using approximation algorithms for betweenness centrality on finding inter-genome repeats using the simulated datasets described above.
@@
 The result of each parameter is averaged over 10 trials.
+
+\subsection*{Cutoff Criteria For Repeats}
+Bambus2\cite{bambus} uses exact betweenness centrality to find repeats.
+To filter out repeats, it calculates mean $\mu$ and standard deviation $\sigma$ over the centrality scores for all the nodes in graph.
+It then marks all the nodes with centrality score greater than $\mu + 3*\sigma$ as repeats.
+To filter out repeats based on approximate centrality scores, we use the same method as Bambus2.
+
+Another cutoff criteria we test is based on \textit{interquartile range(IQR)}.
+Let $Q_{1}$,$Q_{2}$, and $Q_{3}$ denote the lower, middle and upper quartiles respectively.
+IQR can be defined as the difference between the upper and lower quartile.
+So, $IQR = Q_{3} - Q_{1}$.
+For the values of centrality, we calculate upper quartile, middle quartile, lower quartile and IQR.
+For filtering, we mark all the nodes with centrality more than $Q_{3} + 1.5*IQR$ as repeats.
+The advantage of this method is that it does not make any assumptions about the distribution of underlying data.
+
+\subsection*{Implementation}
+We investigate the performance of approximate betweenness centrality algorithms on the metagenomic graphs to find inter-genome repeats.
+We carried out experiments on a 64-bit Linux machine with 12 CPUs and 50 GB memory.
+We used the implementations provided by \textsc{NetworKit}\cite{networkit} library.
+It implements effecient graph algorithms and parallelize most of them.
+It provides implementation for variety of centrality measures.
+We used the implementation of approximate betweenness centrality.
+For our experiments, we use the parallelized implementation for the algorithm by Matteo et al.\cite{matteo} and serial implementation for the algorithm by Geisberger et al.\cite{sanders}.
+
+\section{Results}
 

### 2015-05-18 manuscript/manuscript.tex
methods updated
added bambus2 result

@@
 \centering
-\includegraphics[width=\textwidth]{es_mix_200kb_k21.png}
+\includegraphics[width=0.60\textwidth]{es_mix_200kb_k21.png}
 \caption{Assembly graph of a simulated community consisting of 200 Kbp subsets of \textit{Escherichia coli} str.
@@
 \subsection*{Results On Real Metagenomic Datasets}
-We evaluated the runtime of Bambus2's repeat detection module on metagonic stool sample. It
+We evaluated the performance of approximate betweenness centrality algorithms on metagenomic stool sample.
+We compared the performance of these algorithms with Bambus2 repeat detection module.
+Bambus2 repeat detection module took over 50 hours to find repeats in this graph.
+The algorithm based on sampling shortest paths took 312.69 seconds for $\epsilon = 0.1$ to find out repeats.
+The algorithm based on sampling nodes took 469 seconds for the sample of 1000 nodes and around 5000 seconds(1.3 hours) for the sample of 10000 nodes.
+From these results it can be concluded that we can use approximation algorithms for repeat finding and get significant speedups.
 

### 2015-05-18 manuscript/manuscript.tex
Added discussion.

--- 2015-05-18 manuscript/manuscript.tex
@@
 It has been shown that the number of reconstructions grow exponentially with the number of repeats making it computationally infeasible to find one correct solution for assembly\cite{kingsford}.
-The effect of repeats on assembly is exacerbated in metagenomic data sets due to the presence of conserved regions of the genome linking together distantly related organisms.
+The effect of repeats on assembly is exacerbated in metagenomic datasets due to the presence of conserved regions of the genome linking together distantly related organisms.
 To tackle the problem of identifying these inter-genomic repeats, several assemblers have used the notion of network centrality to find repeats within the assembly graph.
@@
 To overcome this drawback, we explore the application of approximate and parallel betweenness centrality algorithms on metagenomic assembly graphs and evaluate their effectiveness in identifying inter-genomic repeats.
+
+\begin{figure*}[htbp]
+\centering
+\includegraphics[width=0.60\textwidth]{es_mix_200kb_k21.png}
+\caption{Assembly graph of a simulated community consisting of 200 Kbp subsets of \textit{Escherichia coli} str.
+K-12 MG1655 and \textit{Staphylococcus aureus}.
+Nodes are colored and sized based on their relative betweenness centrality.
+The highlighted nodes are inter-genomic repeats whose deletion would separate the graph.
+Note that the betweenness centrality measure correctly identifies these nodes.}
+\label{fig:sampled_nodes}
+\end{figure*}
 
@@
 
-\begin{figure*}[htbp]
-\centering
-\includegraphics[width=\textwidth]{es_mix_200kb_k21.png}
-\caption{Assembly graph of a simulated community consisting of 200 Kbp subsets of \textit{Escherichia coli} str.
-K-12 MG1655 and \textit{Staphylococcus aureus}.
-Nodes are colored and sized based on their relative betweenness centrality.
-The highlighted nodes are inter-genomic repeats whose deletion would separate the graph.
-Note that the betweenness centrality measure correctly identifies these nodes.}
-\label{fig:sampled_nodes}
-\end{figure*}
-
 \subsection*{Betweenness Centrality Based on Sampling Shortest Paths}
@@
 The shortest paths are sampled using weighted random sampling.
-The runtime of this algorithm is dominated by the computation of shortest path at each step which takes time $O(|V| + |E|)$ for unweighted graphs(BFS) and $O(|E| + |V|log|V|)$ otherwise (Dijkstra's algorithm).
+The runtime of this algorithm is dominated by the computation of shortest path at each step which takes time $O(|V| + |E|)$ for unweighted graphs (BFS) and $O(|E| + |V|log|V|)$ otherwise (Dijkstra's algorithm).
 This is multiplied by $r$ for $r$ iterations to obtain final complexity.
@@
 
-Apart from simulated metagenomic graphs, we study how approximation algorithms for betweenness centrality work on real datasets.
-We evaluated the runtime  of these algorithms on metagenomic stool sample of a premature infant\cite{morowitz2011strain}.
-This graph has 3797425 nodes and 2970379 edges.
-We compared the performance of Bambus2's repeat detection module on this data with the performance of approximate betweenness centrality algorithms.
+In addition to the simulated metagenomic assembly graphs, we study how approximation algorithms for betweenness centrality perform on real datasets.
+We evaluated the runtime  of these algorithms on a metagenomic stool sample of a premature infant\cite{morowitz2011strain}.
+This graph has 3,797,425 nodes and 2,970,379 edges.
+We compared the runtime of Bambus2's repeat detection module on this data with the runtime of approximate betweenness centrality algorithms.
 
@@
 
-Across all data sets, the more nodes and paths sampled, the better the inter-genomic repeat detection (Figures \ref{fig:sampled_nodes} and \ref{fig:sampled_paths}).
-Furthermore, the runtime increases linearly with the number of nodes sampled (Figure \ref{fig:sampled_nodes}).
+Across all datasets, the more nodes and paths sampled, the better the inter-genomic repeat detection (Figs. \ref{fig:sampled_nodes} and \ref{fig:sampled_paths}).
+Furthermore, the runtime increases linearly with the number of nodes sampled (Fig. \ref{fig:sampled_nodes}).
 In the 2-genome assembly graph, all inter-genomic repeats were found by sampling as few as ten nodes, an order of magnitude less time than calculating the betweenness centrality using the full graph.
@@
 The sensitivity drops off drastically in the 10-genome assembly graph, oscillating around 6\% across all node sample sizes.
-While the specificity remains consistent across data sets (97\%-99\%), the precision increases from ~2.5\% to ~25\% when going from the 2-genome to the 10-genome data set.
+While the specificity remains consistent across datasets (97\%-99\%), the precision increases from ~2.5\% to ~25\% when going from the 2-genome to the 10-genome data set.
 
-For the range of epsilon parameters used, the statistical measures of repeat detection remain constant (Figure \ref{fig:sampled_paths}).
+For the range of epsilon parameters used, the statistical measures of repeat detection remain constant (Fig. \ref{fig:sampled_paths}).
 Similarly, the runtime for each epsilon parameter remains relatively unchanged except for a slight change in the 10-genome sample.
@@
 We have taken the first steps to show how recent approximation algorithms for calculating betweenness centrality can be used retain the same repeat detection power as the exact algorithms in a fraction of the time.
+
+The advantage of using the node sampling algorithm is that it appears to scale linearly in practice.
+Thus, if runtime is a concern, small increasing subsets of nodes can be sampled and the runtime of a larger sample can be extrapolated.
+We can pair this information with how the subset of nodes marked as inter-genomic repeats changes as we increase the sample size.
+If the same set of nodes are being marked as inter-genomic repeats, then the algorithm can be terminated before sampling a larger set of nodes.
+Conversely, if there is high variability between nodes marked as repeats in subsequent samplings, then a larger set of nodes should be sampled.
+This is the main advantage of the shortest path sampling approach.
+While the shortest path sampling approach takes a fair bit longer in our approach then small node sample sizes, it has a probabilistic guarantee associated with the betweenness centrality computation.
+
 In the future, larger and more complex metagenomic assembly graphs must be used to further evaluate the efficacy of the approximation algorithms.

### 2015-05-18 manuscript/manuscript.tex
Modified bambus2 results.

--- 2015-05-18 manuscript/manuscript.tex
@@
 \subsection*{Results On Real Metagenomic Datasets}
-We evaluated the runtime of Bambus2's repeat detection module on metagonic stool sample. It
+We evaluated the performance of approximate betweenness centrality algorithms on metagenomic stool sample.
+We compared the performance of these algorithms with Bambus2's repeat detection module.
+Bambus2's repeat detection module took over 50 hours to find repeats in the assembly graph.
+The algorithm based on sampling shortest paths took 312.69 seconds for $\epsilon = 0.1$.
+The algorithm based on sampling nodes took 469 seconds for the sample of 1,000 nodes and approximately 1.3 hours for the 10,000 node sample.
+Therefore, we can conclude that the approximation algorithms offers significant speedup for repeat finding.
 

### 2015-05-18 manuscript/manuscript.tex
tmpcommit

--- 2015-05-18 manuscript/manuscript.tex
@@
 It has been shown that the number of reconstructions grow exponentially with the number of repeats making it computationally infeasible to find one correct solution for assembly\cite{kingsford}.
-The effect of repeats on assembly is exacerbated in metagenomic datasets due to the presence of conserved regions of the genome linking together distantly related organisms.
+The effect of repeats on assembly is exacerbated in metagenomic data sets due to the presence of conserved regions of the genome linking together distantly related organisms.
 To tackle the problem of identifying these inter-genomic repeats, several assemblers have used the notion of network centrality to find repeats within the assembly graph.
@@
 To overcome this drawback, we explore the application of approximate and parallel betweenness centrality algorithms on metagenomic assembly graphs and evaluate their effectiveness in identifying inter-genomic repeats.
-
-\begin{figure*}[htbp]
-\centering
-\includegraphics[width=0.60\textwidth]{es_mix_200kb_k21.png}
-\caption{Assembly graph of a simulated community consisting of 200 Kbp subsets of \textit{Escherichia coli} str.
-K-12 MG1655 and \textit{Staphylococcus aureus}.
-Nodes are colored and sized based on their relative betweenness centrality.
-The highlighted nodes are inter-genomic repeats whose deletion would separate the graph.
-Note that the betweenness centrality measure correctly identifies these nodes.}
-\label{fig:sampled_nodes}
-\end{figure*}
 
@@
 
+\begin{figure*}[htbp]
+\centering
+\includegraphics[width=0.60\textwidth]{es_mix_200kb_k21.png}
+\caption{Assembly graph of a simulated community consisting of 200 Kbp subsets of \textit{Escherichia coli} str.
+K-12 MG1655 and \textit{Staphylococcus aureus}.
+Nodes are colored and sized based on their relative betweenness centrality.
+The highlighted nodes are inter-genomic repeats whose deletion would separate the graph.
+Note that the betweenness centrality measure correctly identifies these nodes.}
+\label{fig:sampled_nodes}
+\end{figure*}
+
 \subsection*{Betweenness Centrality Based on Sampling Shortest Paths}
@@
 The shortest paths are sampled using weighted random sampling.
-The runtime of this algorithm is dominated by the computation of shortest path at each step which takes time $O(|V| + |E|)$ for unweighted graphs (BFS) and $O(|E| + |V|log|V|)$ otherwise (Dijkstra's algorithm).
+The runtime of this algorithm is dominated by the computation of shortest path at each step which takes time $O(|V| + |E|)$ for unweighted graphs(BFS) and $O(|E| + |V|log|V|)$ otherwise (Dijkstra's algorithm).
 This is multiplied by $r$ for $r$ iterations to obtain final complexity.
@@
 We represent the unipath as a single node whose label is the concatenation of all node labels along the path.
-Nodes in the resulting graph represent uniquely resolvable sequences of length $>= k$.
+Nodes in the resulting graph represent uniquely resolvable sequences of length $\geq k$.
 
@@
 
-In addition to the simulated metagenomic assembly graphs, we study how approximation algorithms for betweenness centrality perform on real datasets.
-We evaluated the runtime  of these algorithms on a metagenomic stool sample of a premature infant\cite{morowitz2011strain}.
-This graph has 3,797,425 nodes and 2,970,379 edges.
-We compared the runtime of Bambus2's repeat detection module on this data with the runtime of approximate betweenness centrality algorithms.
+Apart from simulated metagenomic graphs, we study how approximation algorithms for betweenness centrality work on real datasets.
+We evaluated the runtime  of these algorithms on metagenomic stool sample of a premature infant\cite{morowitz2011strain}.
+This graph has 3797425 nodes and 2970379 edges.
+We compared the performance of Bambus2's repeat detection module on this data with the performance of approximate betweenness centrality algorithms.
 
@@
 
-Across all datasets, the more nodes and paths sampled, the better the inter-genomic repeat detection (Figs. \ref{fig:sampled_nodes} and \ref{fig:sampled_paths}).
-Furthermore, the runtime increases linearly with the number of nodes sampled (Fig. \ref{fig:sampled_nodes}).
+Across all data sets, the more nodes and paths sampled, the better the inter-genomic repeat detection (Figures \ref{fig:sampled_nodes} and \ref{fig:sampled_paths}).
+Furthermore, the runtime increases linearly with the number of nodes sampled (Figure \ref{fig:sampled_nodes}).
 In the 2-genome assembly graph, all inter-genomic repeats were found by sampling as few as ten nodes, an order of magnitude less time than calculating the betweenness centrality using the full graph.
@@
 The sensitivity drops off drastically in the 10-genome assembly graph, oscillating around 6\% across all node sample sizes.
-While the specificity remains consistent across datasets (97\%-99\%), the precision increases from ~2.5\% to ~25\% when going from the 2-genome to the 10-genome data set.
+While the specificity remains consistent across data sets (97\%-99\%), the precision increases from ~2.5\% to ~25\% when going from the 2-genome to the 10-genome data set.
 
-For the range of epsilon parameters used, the statistical measures of repeat detection remain constant (Fig. \ref{fig:sampled_paths}).
+For the range of epsilon parameters used, the statistical measures of repeat detection remain constant (Figure \ref{fig:sampled_paths}).
 Similarly, the runtime for each epsilon parameter remains relatively unchanged except for a slight change in the 10-genome sample.
@@
 We have taken the first steps to show how recent approximation algorithms for calculating betweenness centrality can be used retain the same repeat detection power as the exact algorithms in a fraction of the time.
-
-The advantage of using the node sampling algorithm is that it appears to scale linearly in practice.
-Thus, if runtime is a concern, small increasing subsets of nodes can be sampled and the runtime of a larger sample can be extrapolated.
-We can pair this information with how the subset of nodes marked as inter-genomic repeats changes as we increase the sample size.
-If the same set of nodes are being marked as inter-genomic repeats, then the algorithm can be terminated before sampling a larger set of nodes.
-Conversely, if there is high variability between nodes marked as repeats in subsequent samplings, then a larger set of nodes should be sampled.
-This is the main advantage of the shortest path sampling approach.
-While the shortest path sampling approach takes a fair bit longer in our approach then small node sample sizes, it has a probabilistic guarantee associated with the betweenness centrality computation.
-
 In the future, larger and more complex metagenomic assembly graphs must be used to further evaluate the efficacy of the approximation algorithms.

### 2015-05-18 manuscript/manuscript.tex
Modified discussion.

--- 2015-05-18 manuscript/manuscript.tex
@@
 If the same set of nodes are being marked as inter-genomic repeats, then the algorithm can be terminated before sampling a larger set of nodes.
-Conversely, if there is high variability between nodes marked as repeats in subsequent samplings, then a larger set of nodes should be sampled.
+Conversely, if there is high variability between nodes marked as repeats in subsequent samplings, then a larger set of nodes should be sampled until the results stabilize.
 This is the main advantage of the shortest path sampling approach.
-While the shortest path sampling approach takes a fair bit longer in our approach then small node sample sizes, it has a probabilistic guarantee associated with the betweenness centrality computation.
+While the shortest path sampling approach takes a fair bit longer in our evaluation then sampling a small number of nodes, it has a probabilistic guarantee associated with the betweenness centrality computation.
 

### 2015-05-18 manuscript/manuscript.tex
Small edits.

@@
 \subsection*{Evaluation}
-We evaluate the effect of using approximation algorithms for betweenness centrality on finding inter-genome repeats using the simulated datasets described above.
+%We evaluate the effect of using approximation algorithms for betweenness centrality on finding inter-genome repeats using the simulated datasets described above.
 %We use the notion of Sensitivity, specificity, and precision to define the quality of repeat detection.
@@
 
-For evaluating the effect of node sampling on inter-genomic repeat detection, we sample between 10 to 10,000 nodes per trial.
-For evaluating path sampling, we select epsilon parameters between 0.01 and 0.20.
+We evaluate the two approaches for calculating betweenness centrality based on sampling nodes and sampling shortest paths using the above measures.
+For evaluating the effect of sampling nodes on inter-genomic repeat detection, we sample between 10 to 10,000 nodes per trial.
+For evaluating shortest paths sampling, we select epsilon parameters between 0.01 and 0.20.
 The result of each parameter is averaged over 10 trials.
@@
 \subsection*{Implementation}
-We investigate the performance of approximate betweenness centrality algorithms on the metagenomic graphs to find inter-genome repeats.
+We investigate the performance of approximate betweenness centrality algorithms on the metagenomic assembly graphs using the implementations provided by \textsc{NetworKit}\cite{networkit} library.
+\textsc{NetworKit} contains efficient and parallel implementations of a collection of graph algorithms.
+It provides implementations for various centrality measures, but here we only use the implementations related to betweenness centrality.
+For our experiments, we use the parallelized implementation for the algorithm by Matteo et al.\cite{matteo} and serial implementation for the algorithm by Geisberger et al.\cite{sanders}.
 We carried out experiments on a 64-bit Linux machine with 12 CPUs and 50 GB memory.
-We used the implementations provided by \textsc{NetworKit}\cite{networkit} library.
-It implements effecient graph algorithms and parallelize most of them.
-It provides implementation for variety of centrality measures.
-We used the implementation of approximate betweenness centrality.
-For our experiments, we use the parallelized implementation for the algorithm by Matteo et al.\cite{matteo} and serial implementation for the algorithm by Geisberger et al.\cite{sanders}.
 
@@
 
-The advantage of using the node sampling algorithm is that it appears to scale linearly in practice.
+The advantage of calculating betweenness centrality using node sampling is that it appears to scale linearly in practice.
 Thus, if runtime is a concern, small increasing subsets of nodes can be sampled and the runtime of a larger sample can be extrapolated.

### 2015-05-18 manuscript/manuscript.tex
edited IQR

--- 2015-05-18 manuscript/manuscript.tex
@@
 Decreasing the betweenness centrality cutoff requirement results in a trade-off between the different statistical measures.
-Instead of using the default cutoff of $3\sigma$ suggested by Bambus2, we mark all the nodes with centrality more than $Q_{3} + 1.5*IQR$ as an inter-genomic repeats.
-In the 10-genome assembly graph, the sensitivity increases from 5.8\% to 34.4\% when using the IQR cutoff method.
+We observed that for 10-genome assembly graph, the sensitivity was very less according to the cut off criteria in Bambus2.
+This was due to the fact that cut off was strict and was eliminating actual repeats.
+Instead of using the cut off criteria suggested by Bambus2, we mark all the nodes with centrality more than $Q_{3} + 1.5*IQR$ as an inter-genomic repeats.
+When we use this criteria on 10-genome assembly graph, the sensitivity increases from 5.8\% to 34.4\% when using the IQR cutoff method.
 Conversely, the specificity and and precision drops from 99.3\% to 90.9\% and 25\% to 12.5\%, respectively.
+We could achieve better specificity at by not losing much on sensitivity.
 
@@
 If the same set of nodes are being marked as inter-genomic repeats, then the algorithm can be terminated before sampling a larger set of nodes.
-Conversely, if there is high variability between nodes marked as repeats in subsequent samplings, then a larger set of nodes should be sampled until the results stabilize.
+Conversely, if there is high variability between nodes marked as repeats in subsequent samplings, then a larger set of nodes should be sampled.
 This is the main advantage of the shortest path sampling approach.
-While the shortest path sampling approach takes a fair bit longer in our evaluation then sampling a small number of nodes, it has a probabilistic guarantee associated with the betweenness centrality computation.
+While the shortest path sampling approach takes a fair bit longer in our approach then small node sample sizes, it has a probabilistic guarantee associated with the betweenness centrality computation.
 

### 2015-05-19 manuscript/manuscript.tex
Added back previous edits.

@@
 The effect of repeats on assembly is exacerbated in metagenomic datasets due to the presence of conserved regions of the genome linking together distantly related organisms.
-To tackle the problem of identifying these inter-genomic repeats, several assemblers have used the notion of network centrality to find repeats within the assembly graph.
+To tackle the problem of identifying these inter-genomic repeats, several assemblers have used the notion of network centrality to find repeats within the assembly graph (Figure \ref{fig:es_mix}).
 For example, Bambus2\cite{bambus} uses betweenness centrality to find \emph{central} nodes in the graph.
@@
 \centering
-\includegraphics[width=0.60\textwidth]{es_mix_200kb_k21.png}
+\includegraphics[width=0.70\textwidth]{es_mix_200kb_k21.png}
 \caption{Assembly graph of a simulated community consisting of 200 Kbp subsets of \textit{Escherichia coli} str.

### 2015-05-19 manuscript/manuscript.tex
Fixed figure ref.

@@
 The effect of repeats on assembly is exacerbated in metagenomic datasets due to the presence of conserved regions of the genome linking together distantly related organisms.
-To tackle the problem of identifying these inter-genomic repeats, several assemblers have used the notion of network centrality to find repeats within the assembly graph (Figure \ref{fig:es_mix}).
+To tackle the problem of identifying these inter-genomic repeats, several assemblers have used the notion of network centrality to find repeats within the assembly graph (Fig. \ref{fig:es_mix}).
 For example, Bambus2\cite{bambus} uses betweenness centrality to find \emph{central} nodes in the graph.
@@
 Note that the betweenness centrality measure correctly identifies these nodes.}
-\label{fig:sampled_nodes}
+\label{fig:es_mix}
 \end{figure*}

### 2015-05-19 manuscript/manuscript.tex
Edited methods.

@@
 These strings can be represented in a sophisticated way in term of a graph called de Bruijn graph \cite{debruijn}.
-A de Bruijn graph of order $k$ is a graph that contains all strings of length $k-1$ from a given alphabets as nodes and contains edges between two nodes if the corresponding strings overlap by exactly $k-2$ letters in prefix-suffix manner.
+A de Bruijn graph of order $k$ is a directed graph that contains all strings of length $k-1$ from a given alphabets as nodes and contains edges between two nodes if the corresponding strings overlap by exactly $k-2$ letters in prefix-suffix manner.
 In other words, each $k$-mer is represented by two nodes in a graph connected by an edge.
 
-Next we modify the de Bruijn graph generated from genomes to identify repeats.
+Next we modify the de Bruijn graph generated from the collection of genomes to identify repeats $>=k$.
 We define \textit{unipaths} as the path of nodes in which the outdegree of first node on the path is $1$ and indegree is not equal to $1$, indegree of last node on the path is $1$ and outdegree is not equal to $1$ and for all other nodes along the path, both indegree and out degree is exactly $1$.
@@
 
-In addition to the simulated metagenomic assembly graphs, we study how approximation algorithms for betweenness centrality perform on real datasets.
-We evaluated the runtime  of these algorithms on a metagenomic stool sample of a premature infant\cite{morowitz2011strain}.
+In addition to the simulated metagenomic assembly graphs, we evaluate the runtimes of the two approximation algorithms for betweenness centrality on a metagenomic stool sample of a premature infant\cite{morowitz2011strain}.
 This graph has 3,797,425 nodes and 2,970,379 edges.

### 2015-05-19 manuscript/manuscript.tex
Small edits.

@@
 
-For the range of epsilon parameters used, the statistical measures of repeat detection remain constant (Fig. \ref{fig:sampled_paths}).
+The statistical measures of repeat detection remain constant across the range of epsilon parameters used in the shortest path sampling approach (Fig. \ref{fig:sampled_paths}).
 Similarly, the runtime for each epsilon parameter remains relatively unchanged except for a slight change in the 10-genome sample.
-For the 10-genome sample, the statistical measures and runtime are equivalent to when 100-500 nodes are sampled.
+The measures and runtime of the shortest path sampling approach are equivalent to when roughly 100-500 nodes are sampled in the 10-genome sample.
+%For the 10-genome sample, the statistical measures and runtime are equivalent to when 100-500 nodes are sampled.
 

### 2015-05-19 manuscript/manuscript.tex
equation changed

--- 2015-05-19 manuscript/manuscript.tex
@@
 The sample size $r$ of shortest paths is found out with the help of $VD(G)$ as follows:
-\begin{equation} \label{sample_size_algo}
+$$
 r = \frac{c}{\epsilon^{2}}(\left\lfloor{log_{2}(VD(G) - 2)}\right\rfloor + 1 + ln\frac{1}{\delta})
-\end{equation} 
+$$ 
 where $c$ is a constant with value $0.5$.

### 2015-05-19 manuscript/manuscript.tex
Small edits.

@@
 Bambus2's repeat detection module took over 50 hours to find repeats in the assembly graph.
-The algorithm based on sampling shortest paths took 312.69 seconds for $\epsilon = 0.1$.
+The algorithm based on sampling shortest paths took 313 seconds for $\epsilon = 0.1$.
 The algorithm based on sampling nodes took 469 seconds for the sample of 1,000 nodes and approximately 1.3 hours for the 10,000 node sample.
-Therefore, we can conclude that the approximation algorithms offers significant speedup for repeat finding.
+Thus, the approximation algorithms offer significant speedups for repeat finding.
 

### 2015-05-19 manuscript/manuscript.tex
Updated abstract.

@@
 
+% I'll redo the abstract on Monday, but broadly the points are:
+% 1. inter-/intra-genomic repeats are the main reason assembly is complicated
+% 2. in a metagenomic context repeats are not easy to find due to uneven coverage and confounding effect of strain variants
+% 3. centrality has been proposed as a solution to this problem, but the original algorithms used are inefficient
+% 4. Here we explore the use of approximate centrality measures and show that repeats can be accurately and efficiently detected.
+
 \begin{abstract}
-Assembling metagenomic data is a challenging task due to several reasons.
-One of the reasons that complicates assembly is the sequences shared between the genomes of multiple organisms.
-Such regions in genomes are called repeats.
-Repeats make assembly difficult because they link the unrelated sections of genomes.
-It has been shown that the number of reconstructions of genome grows exponentially with repeats.
-Existing approaches to identify repeats are inefficient and do not scale well as size of data increases.
-We propose an application of a core concept in social network analysis called betweenness centrality to metagenomic data to identify inter-genomic repeats.
-We show how approximation algorithms for betweenness centrality can be useful to design a faster approach to identify inter-genomic repeats accurately.
+Assembling genomic data is challenging due in large part to the presence of repetitive regions.
+This issue is compounded in metagenomic datasets due to uneven coverages and the presence of conserved genomic regions found in distantly related organisms.
+Centrality has been proposed as a solution to the problem of identify inter-genomic repeats, but the original algorithms used are inefficient.
+Here we explore the use of approximate centrality measures and show that repeats can be accurately and efficiently detected.
 \keywords{Metagenomics, Algorithms, Centrality, Graph}

### 2015-05-19 manuscript/manuscript.tex
Updated abstract.
Incorporated mihai's changes.

--- 2015-05-19 manuscript/manuscript.tex
@@
 \begin{abstract}
-Assembling metagenomic data is a challenging task due to several reasons.
-One of the reasons that complicates assembly is the sequences shared between the genomes of multiple organisms.
-Such regions in genomes are called repeats.
-Repeats make assembly difficult because they link the unrelated sections of genomes.
-It has been shown that the number of reconstructions of genome grows exponentially with repeats.
-Existing approaches to identify repeats are inefficient and do not scale well as size of data increases.
-We propose an application of a core concept in social network analysis called betweenness centrality to metagenomic data to identify inter-genomic repeats.
-We show how approximation algorithms for betweenness centrality can be useful to design a faster approach to identify inter-genomic repeats accurately.
+Genomic repeats are the most important challenge in genomic assembly.
+While for single genomes the effect of repeats is largely addressed by modern long-read sequencing technologies, in metagenomic data intra-genome and, more importantly, inter-genome repeats continue to be a significant impediment to effective genome reconstruction.
+Detecting repeats in metagenomic samples is complicated by characteristic features of these data, primarily uneven depths of coverage and the presence of genomic polymorphisms.
+The scaffolder Bambus 2 introduced a new strategy for repeat detection based on the betweenness centrality measure -- a concept originally used in social network analysis.
+The exact computation of the betweenness centrality measure is, however, computationally intensive and impractical in large metagenomic datasets.
+Here we explore the effectiveness of approximate algorithms for network centrality to accurately detect genomic repeats within metagenomic samples.
+We show that approximate measures of centrality achieve much higher computational efficiencies with a minimal loss in the accuracy of detecting repeats in metagenomic data.
 \keywords{Metagenomics, Algorithms, Centrality, Graph}
@@
 
-The assembly of metagenomics data, the sequences of DNA recovered from an environment without culturing, is considerably more complex than that of individual genomes.
-%Challenges include: (i) repetitive sequences within individual genomes; (ii) varying abundances of different organisms in the community; (iii) variation between closely related organisms; and (iv) conserved genomic segments shared by distantly related organisms.
-Repetitive sequences pose a major challenge when assembling isolated genomes.
-It has been shown that the number of reconstructions grow exponentially with the number of repeats making it computationally infeasible to find one correct solution for assembly\cite{kingsford}.
-The effect of repeats on assembly is exacerbated in metagenomic datasets due to the presence of conserved regions of the genome linking together distantly related organisms.
-To tackle the problem of identifying these inter-genomic repeats, several assemblers have used the notion of network centrality to find repeats within the assembly graph (Fig. \ref{fig:es_mix}).
-For example, Bambus2\cite{bambus} uses betweenness centrality to find \emph{central} nodes in the graph.
-The drawback of their approach is the expensive calculation of betweenness centrality by computing all-pair shortest paths.
-This computation is impractical when applied to the large assembly graphs typically encountered.
-To overcome this drawback, we explore the application of approximate and parallel betweenness centrality algorithms on metagenomic assembly graphs and evaluate their effectiveness in identifying inter-genomic repeats.
+Genomic repeats are the most important challenge in genomic assembly even for isolate genomes.
+When reads are shorter than the repeats (a common situation until the recent development of long read sequencing technologies) it can be shown that the number of genome reconstructions consistent with the read data grows exponentially with the number of repeats~\cite{kingsford}.
+The use of additional information to constrain the one genome reconstruction representing the actual genome being assembled leads to computationally intractable problems.
+In other words, when reads are shorter than repeats the correct and complete reconstruction of a genome is impossible.
+In the case of isolate genomes, long read technologies have largely addressed this challenge, at least for bacteria where the majority of genomic repeats fall within the range of achievable read lengths~\cite{koren2015one}.
+In metagenomics, however, the problem is compounded by the fact that microbial mixtures often include multiple closely-related genomes differing in just a few locations.
+The genomic segments shared by closely related organisms -- inter-genomic repeats -- are substantially larger than intra-genomic repeats and cannot be fully resolved even if long read data were available.
+Instead, the best hope is to identify and flag these repeats  in order to avoid mis-assemblies that incorrectly span across genomes.
+
+Repeat detection strategies have been employed in genome assembly from the early days of the field, and numerous approaches have been developed for this purpose.
+In metagenomics, however, the problem is exacerbated by uneven coverage across contigs and by the fact that both inter- and intra-genomic repeats exist in the data, factors impacting statistical approaches for repeat detection.
+An alternate approach based on the social network concept of betweenness centrality was introduced in Bambus 2~\cite{bambus}.
+An example of the effectiveness of this approach in a simple community composed of two genomes is shown in Fig.~\ref{fig:es_mix}.
+The full implementation of betweenness centrality requires an all-pairs shortest path computation which is computationally too intensive for typical metagtenomic datasets.
+In Bambus 2, for example, repeat finding in a typical stool sample requires days of computation.
+To overcome this limitation, we explore the use of parallel and approximate betweenness centrality algorithms for repeat detection in metagenomic assembly graphs, and the effect of the level of approximation on the efficiency and accuracy of these algorithms.
 
@@
 K-12 MG1655 and \textit{Staphylococcus aureus}.
-Nodes are colored and sized based on their relative betweenness centrality.
+Nodes are colored and sized based on their relative betweenness centrality with larger, green nodes indicating a higher centrality.
 The highlighted nodes are inter-genomic repeats whose deletion would separate the graph.
@@
 The betweenness centrality of a particular node is equal to the number of shortest paths from all nodes to all others that pass through that node.
-Formally, for a node $v$, it is defined by following expression:
+Intuitively, a node that is frequently found on paths connecting other nodes is ``important''.
+In assembly, we can likewise infer that nodes that are on the paths connecting multiple distinct segments of a genomic assembly likely represent repetitive sequences.
+Formally, for a node $v$, centrality is defined by following expression:
 
@@
 The contributions of these paths to the betweenness counters can be computed in linear time using backward aggregation.
-Brandes algorithm calculates $\sigma_{st}$  simultaneously with shortest path calculations: $\sigma_{ss} = 1$ and for $s \neq t, \sigma_{st} = \sum_{v \in pred(t)} \sigma_{sv}$ where $pred(t)$ is a collection of the immediate predecessors of  $t$ on every shortes path ending at $t$.
+Brandes' algorithm calculates $\sigma_{st}$  simultaneously with the shortest path calculations: $\sigma_{ss} = 1$ and for $s \neq t, \sigma_{st} = \sum_{v \in pred(t)} \sigma_{sv}$ where $pred(t)$ is a collection of the immediate predecessors of  $t$ on every shortes path ending at $t$.
 In a subsequent aggregation phase, nodes are processed by non-decreasing distance from $s$.
@@
 This algorithm takes time $\Theta(mn)$ for unit edge weight graphs and $\Theta(mn + n^{2}log(n))$ for weighted graphs.
-This algorithm does not scale well with large networks with millions of nodes and edges.
-Bader and Madduri\cite{bader} provide a massively parallel implementation of the exact algorithm, but scales in practice to only a few million nodes, limiting its application to larger metagenomic assembly graphs that exceed million of nodes.
+This algorithm does not scale well in large networks with millions of nodes and edges.
+Bader and Madduri\cite{bader} provide a massively parallel implementation of the exact algorithm, but even the parallel implementation only scales in practice to  a few million nodes, limiting its application to larger metagenomic assembly graphs.
 
-\subsection*{Betweenness Centrality Based on Sampling Nodes}
+\subsection*{Approximate Betweenness Centrality Based on Node Sampling}
 Due to scalability limitations of exact betweenness centrality, several approximation algorithms for betweenness centrality have been proposed.
-Bader and Pich\cite{bp} provide an approximation algorithm by choosing a subset of $k$ starting nodes called pivots and using same aggregation strategy as Brandes algorithm.
-However, this algorithm overestimates the centrality of some unimportant nodes which are nearer to pivots.
-The approximation algorithm proposed by Geisberger et al.
-\cite{sanders} overcomes this drawback by changing the scheme for aggregating betweenness contributions so that nodes nearer to pivots are not unduly profited.
+Bader and Pich\cite{bp} provide an approximation algorithm by choosing a subset of $k$ starting nodes called pivots and apply the Brandes algorithm to just the chosen nodes.
+This algorithm was shown to overestimate the centrality of some unimportant nodes which are close to the pivots.
+Geisberger et al.
+\cite{sanders} sole this issue by changing the scheme for aggregating betweenness contributions so that nodes close to the pivots are not unduly profited.
 The algorithm starts with the sample of $k$ nodes from $|V| = n$ nodes in graph.
 In each iteration, it chooses one of $2n$ shortest path at random.
-Here there are $2n$ sample shortest paths because algorithm can do forward search from $s \in V$ or backward search from $t \in V$ and each $s$ and $t$ can have $n$ values.
+Here there are $2n$ sample shortest paths because the algorithm can perform a forward search from $s \in V$ or a backward search from $t \in V$ and each $s$ and $t$ can have $n$ values.
 Let $\delta_{P}(v)$ be the contribution received by node $v$ due to shortest path $P$.
@@
 
-This algorithm is efficiently implemented by modifying original aggregation scheme of Brandes algorithm so that the nodes near the sampled node will not receive high centrality score.
-Runtime of the algorithm largely depends on the size of the sample we choose.
-However,there is a tradeoff between size of sample and accuracy of centrality detection.
+The runtime and accuracy of these algorithms depend on the number of nodes sampled from the graph.
 
-\subsection*{Betweenness Centrality Based on Sampling Shortest Paths}
-Another approximation algorithm was proposed by Matteo et al.
-\cite{matteo} based on randomized sampling of shortest paths and offer probabilistic guarantees on the quality of approximation.
+\subsection*{Approximate Betweenness Centrality Based on Sampling Shortest Paths}
+A different approximation strategy was proposed by Matteo et al.
+\cite{matteo} based on randomized sampling of shortest paths (rather than nodes), approach which offers probabilistic guarantees on the quality of approximation.
 This algorithm guarantees that all approximate values of betweenness for all vertices are within an additive factor $\epsilon \in (0,1)$ from the real values with probability at least $1-\delta$.
 To derive bounds on sample size, results from VC dimension theory\cite{vc} are used.
-The main advantage of this algorithm is that the sample size does not depend on the size of the graph, but only on the maximum number vertices in a shortest path which is called \textit{vertex diameter}.
+The main advantage of this algorithm is that the sample size does not depend on the size of the graph, but only on the maximum number vertices in a shortest path, measure called the \textit{vertex diameter} of the graph.
 The sample size $r$ of shortest paths is found out with the help of $VD(G)$ as follows:
@@
 The runtime of this algorithm is dominated by the computation of shortest path at each step which takes time $O(|V| + |E|)$ for unweighted graphs (BFS) and $O(|E| + |V|log|V|)$ otherwise (Dijkstra's algorithm).
-This is multiplied by $r$ for $r$ iterations to obtain final complexity.
+This value is multiplied by the number of iterations to obtain the final complexity.
 
@@
 
-\subsection*{Simulating Metagenomic Assembly Repeat Graphs}
+\subsection*{Metagenomic Data}
 
 We evaluate the efficacy of detecting inter-genomic repeats using the approximate betweenness centrality methods on simulated metagenomic assembly graphs.
-The first step is to construct the \textit{de Bruijn} graph from the set of reference genomes.
-De Bruijn strings are a type of ``comprehensive'' strings.
-For a given alphabet $\Sigma = \{A,C,T,G\}$ and a length $k$, a de Bruijn string of order $k$ contains as substrings of length $k$ over $\Sigma$.
-These strings can be represented in a sophisticated way in term of a graph called de Bruijn graph \cite{debruijn}.
-A de Bruijn graph of order $k$ is a directed graph that contains all strings of length $k-1$ from a given alphabets as nodes and contains edges between two nodes if the corresponding strings overlap by exactly $k-2$ letters in prefix-suffix manner.
-In other words, each $k$-mer is represented by two nodes in a graph connected by an edge.
+We start with a collection of genomes from which we construct idealized De Bruijn graphs (assuming perfect coverage and no sequencing errors).
+Next we modify the De Bruijn graph generated from the collection of genomes to identify repeats $>=k$.
+We define \textit{unipaths} as maximal paths comprising nodes with in- and out-degrees equal to $1$.
+We compress all the unipaths into single nodes representing the concatenated labels of the collapsed nodes.
+The nodes in the resulting graph represent uniquely resolvable sequences of length $\geq k$.
 
-Next we modify the de Bruijn graph generated from the collection of genomes to identify repeats $>=k$.
-We define \textit{unipaths} as the path of nodes in which the outdegree of first node on the path is $1$ and indegree is not equal to $1$, indegree of last node on the path is $1$ and outdegree is not equal to $1$ and for all other nodes along the path, both indegree and out degree is exactly $1$.
-We compress all the unipaths in a graph to represent the path in a compact way.
-We represent the unipath as a single node whose label is the concatenation of all node labels along the path.
-Nodes in the resulting graph represent uniquely resolvable sequences of length $\geq k$.
-
-For evaluation we simulate metagenomic assembly graphs consisting of two, five, and ten bacterial genomes.
+We simulated metagenomic assembly graphs consisting of two, five, and ten bacterial genomes.
 The composition of each simulated sample is described in Table \ref{tab:composition}.
 For the five and ten metagenome samples, random bacterial genomes were chosen from the data set provided by \cite{shakya2013comparative}.
-A $k$-mer size of 55 was chosen for initial creation of the de Bruijn graph.
+A $k$-mer size of 55 was chosen for the initial creation of the De Bruijn graph.
 
@@
 
-In addition to the simulated metagenomic assembly graphs, we evaluate the runtimes of the two approximation algorithms for betweenness centrality on a metagenomic stool sample of a premature infant\cite{morowitz2011strain}.
+In addition to the simulated metagenomic assembly graphs, we evaluate the runtimes of the two approximation algorithms for betweenness centrality on a human stool metagenomic sample.
 This graph has 3,797,425 nodes and 2,970,379 edges.
@@
 For evaluating the effect of sampling nodes on inter-genomic repeat detection, we sample between 10 to 10,000 nodes per trial.
-For evaluating shortest paths sampling, we select epsilon parameters between 0.01 and 0.20.
-The result of each parameter is averaged over 10 trials.
+For evaluating shortest paths sampling, we select epsilon parameters between 0.01 and 0.20, corresponding to a level of error between $1\%$ and $20\%$.
+We repeat each algorithm 10 times and report the averge results.
 
-\subsection*{Cutoff Criteria For Repeats}
-Bambus2\cite{bambus} uses exact betweenness centrality to find repeats.
-To filter out repeats, it calculates mean $\mu$ and standard deviation $\sigma$ over the centrality scores for all the nodes in graph.
-It then marks all the nodes with centrality score greater than $\mu + 3*\sigma$ as repeats.
-To filter out repeats based on approximate centrality scores, we use the same method as Bambus2.
-
-Another cutoff criteria we test is based on \textit{interquartile range(IQR)}.
+\subsection*{Cutoff Criteria For Identifying Repeats}
+Centrality algorithms provide numeric estimates for the centrality of each node, values from which we can infer whether a graph node represents a repeat.
+In Bambus2~\cite{bambus} repeats were defined as nodes with a centrality score larger than three standard deviations from the mean centrality of the nodes in the graph.
+This strategy implicitly assumes the distribution of centrality values in the graph to follow a normal distribution.
+We use the same definition here and also explore the use of an alternate statistic -- the  \textit{interquartile range(IQR)}.
 Let $Q_{1}$,$Q_{2}$, and $Q_{3}$ denote the lower, middle and upper quartiles respectively.
@@
 So, $IQR = Q_{3} - Q_{1}$.
-For the values of centrality, we calculate upper quartile, middle quartile, lower quartile and IQR.
-For filtering, we mark all the nodes with centrality more than $Q_{3} + 1.5*IQR$ as repeats.
-The advantage of this method is that it does not make any assumptions about the distribution of underlying data.
+To identify repeats we simply mark  all the nodes with centrality values larger than $Q_{3} + 1.5*IQR$.
+This latter approach no longer assumes that the centrality values are normally distributed.
 
@@
 \textsc{NetworKit} contains efficient and parallel implementations of a collection of graph algorithms.
-It provides implementations for various centrality measures, but here we only use the implementations related to betweenness centrality.
 For our experiments, we use the parallelized implementation for the algorithm by Matteo et al.\cite{matteo} and serial implementation for the algorithm by Geisberger et al.\cite{sanders}.
-We carried out experiments on a 64-bit Linux machine with 12 CPUs and 50 GB memory.
+We carried out all experiments on a 64-bit Linux machine with 12 CPUs and 50 GB memory.
 
@@
 
-\subsection*{Approximate Betweenness Centrality Retains Detection Power of Inter-Genomic Repeats}
+\subsection*{Approximate Betweenness Centrality Measures Increase Efficiency without a Significant Loss in Accuracy}
 
-Across all datasets, the more nodes and paths sampled, the better the inter-genomic repeat detection (Figs. \ref{fig:sampled_nodes} and \ref{fig:sampled_paths}).
-Furthermore, the runtime increases linearly with the number of nodes sampled (Fig. \ref{fig:sampled_nodes}).
-In the 2-genome assembly graph, all inter-genomic repeats were found by sampling as few as ten nodes, an order of magnitude less time than calculating the betweenness centrality using the full graph.
-In the 5-genome assembly graph, sampling 10 nodes has a sensitivity of 86\% and reaches 100\% at 500 nodes which is approximately one-tenth the size of the assembly graph.
-The sensitivity drops off drastically in the 10-genome assembly graph, oscillating around 6\% across all node sample sizes.
-While the specificity remains consistent across datasets (97\%-99\%), the precision increases from ~2.5\% to ~25\% when going from the 2-genome to the 10-genome data set.
+Across all datasets, the efficiency of the approach is correlated with the level of approximation for both node sampling and path sampling approaches (Figs.
+\ref{fig:sampled_nodes} and \ref{fig:sampled_paths}).
+Accuracy is inversely proportional with the level of approximation.
+The effectiveness of these methods is also higher in simpler communities.
+In the 2-genome community,  all inter-genomic repeats were found by sampling as few as ten nodes, an order of magnitude less time than calculating the betweenness centrality using the full graph.
+For the 5-genome graph, sampling as few as ten nodes yields a sensitivity of $86\%$ and full sensitivity is achieved with just 500 nodes, or approximately one tenth of the entire size of the assembly graph.
 
-The statistical measures of repeat detection remain constant across the range of epsilon parameters used in the shortest path sampling approach (Fig. \ref{fig:sampled_paths}).
-Similarly, the runtime for each epsilon parameter remains relatively unchanged except for a slight change in the 10-genome sample.
-The measures and runtime of the shortest path sampling approach are equivalent to when roughly 100-500 nodes are sampled in the 10-genome sample.
-%For the 10-genome sample, the statistical measures and runtime are equivalent to when 100-500 nodes are sampled.
+The 10-genome graph is substantially more complex yielding much lower sensitivity for both node sampling and path sampling approaches.
+The use of the inter-quartile range as a decision criterion improves the sensitivity from about $6\%$ to $35\%$ in the 10-genome graph at the cost of a reduction in specificity from $99.3\%$ to $90.9\%$.
+Also, note that the path sampling approach is marginally more efficient than the node sampling procedure  - the worst runtime (at an error setting of $1\%$) is roughly equivalent to sampling about 1000 nodes.
 
@@
 
-\subsection*{Alternate Betweenness Centrality Cutoff Increases Sensitivity}
-
-Decreasing the betweenness centrality cutoff requirement results in a trade-off between the different statistical measures.
-We observed that for 10-genome assembly graph, the sensitivity was very less according to the cut off criteria in Bambus2.
-This was due to the fact that cut off was strict and was eliminating actual repeats.
-Instead of using the cut off criteria suggested by Bambus2, we mark all the nodes with centrality more than $Q_{3} + 1.5*IQR$ as an inter-genomic repeats.
-When we use this criteria on 10-genome assembly graph, the sensitivity increases from 5.8\% to 34.4\% when using the IQR cutoff method.
-Conversely, the specificity and and precision drops from 99.3\% to 90.9\% and 25\% to 12.5\%, respectively.
-We could achieve better specificity at by not losing much on sensitivity.
-
 \subsection*{Results On Real Metagenomic Datasets}
-We evaluated the performance of approximate betweenness centrality algorithms on metagenomic stool sample.
-We compared the performance of these algorithms with Bambus2's repeat detection module.
-Bambus2's repeat detection module took over 50 hours to find repeats in the assembly graph.
-The algorithm based on sampling shortest paths took 313 seconds for $\epsilon = 0.1$.
-The algorithm based on sampling nodes took 469 seconds for the sample of 1,000 nodes and approximately 1.3 hours for the 10,000 node sample.
-Thus, the approximation algorithms offer significant speedups for repeat finding.
+We evaluated the performance of approximate betweenness centrality algorithms on an actual metagenomic stool sample and compared the runtime to that of the MarkRepeats procedure from Bambus 2.
+On this dataset, comprising $3,797,425$ nodes and $2,970,379$ edges, Bambus 2 had not completed after 60 hours at which point we terminated the execution.
+The path sampling approach with an error setting of $1\%$ ( $\epsilon = 0.1$) took $312.69$ seconds, while the node sampling approach toolk $469$ seconds for a $1,000$ node sample and 1.3 hours for a $10,000$  node sample.
 
@@
 Betweenness centrality is a proven method for finding inter-genomic repeats in metagenomic assemblies.
-The size of typical metagenomic assembly graphs often make it infeasible to calculate the exact betweenness centrality of each node.
-We have taken the first steps to show how recent approximation algorithms for calculating betweenness centrality can be used retain the same repeat detection power as the exact algorithms in a fraction of the time.
+The size of typical metagenomic assembly graphs often make it infeasible to calculate the exact betweenness centrality of each node, as demonstrated by the slow runtime of the MarkRepeats procedure from Bambus 2.
+Here we have demonstrated the effectiveness of approximate measures of centrality able to identify repeats in a fraction of the time previously required.
+We explored two alternative approximation strategies, based on node sampling and path sampling, respectively.
+The runtime of node-sampling approaches can be more effectively estimated as it increases roughly linearly in the size of the sample selected.
+Conversely, path sampling approaches can guarantee the level of approximation but runtime cannot be easily estimated.
 
-The advantage of calculating betweenness centrality using node sampling is that it appears to scale linearly in practice.
-Thus, if runtime is a concern, small increasing subsets of nodes can be sampled and the runtime of a larger sample can be extrapolated.
-We can pair this information with how the subset of nodes marked as inter-genomic repeats changes as we increase the sample size.
-If the same set of nodes are being marked as inter-genomic repeats, then the algorithm can be terminated before sampling a larger set of nodes.
-Conversely, if there is high variability between nodes marked as repeats in subsequent samplings, then a larger set of nodes should be sampled until the results stabilize.
-This is the main advantage of the shortest path sampling approach.
-While the shortest path sampling approach takes a fair bit longer in our evaluation then sampling a small number of nodes, it has a probabilistic guarantee associated with the betweenness centrality computation.
-
-In the future, larger and more complex metagenomic assembly graphs must be used to further evaluate the efficacy of the approximation algorithms.
+Despite promising results, our work has also revealed limitations of the approximate approaches - in the more complex communities the sensitivity of detection dropped significantly, though it was partly rescued by the use of a decision cut-off based on inter-quartile ranges.
+We are currently exploring this phenomenon and whether it will have a significant impact on metagenomic assembly.
+It is likely that many of the repeats missed by the approximate procedure are small and local in nature and could be resolved through other means.
+It is also possible that other approaches for outlier detection would be more effective in restoring the sensitivity of detection.
 
 \section{Conclusion}
-In this work, we show how approximate betweenness centrality can be used to identify inter-genome repeats.
-We explored two algorithms, one based on sampling shortest paths in a graph and other based on sampling nodes in a graph.
-Application of both these algorithms to repeat detection problem provided accurate and fast results.
-We explore how various parameters such as number of sampled nodes and filtering criteria to mark a node as repeat affects the accuracy of repeat detection.
-We are planning to use these techniques for repeat detection in our future metagenomic assembler.
+Here we have demonstrated that approximate betweenness centrality approaches can be effectively used to detect repeats in metagenomic assemblies.
+These approaches run in a fraction of the time needed to exactly compute betweenness centrality with only a small loss of accuracy in repeat detection.
+We currently plan to incorporate such algorithms in the Bambus scaffolding software in order to improve its efficiency, and also plan to further develop the repeat detection algorithms in order to improve their sensitivity in complex graphs.
 \bibliography{manuscript}{}

### 2015-05-19 manuscript/manuscript.tex
added edits

@@
 This graph has 3,797,425 nodes and 2,970,379 edges.
-We compared the runtime of Bambus2's repeat detection module on this data with the runtime of approximate betweenness centrality algorithms.
+We compared the runtime of Bambus 2's repeat detection module on this data with the runtime of approximate betweenness centrality algorithms.
 
@@
 Centrality algorithms provide numeric estimates for the centrality of each node, values from which we can infer whether a graph node represents a repeat.
-In Bambus2~\cite{bambus} repeats were defined as nodes with a centrality score larger than three standard deviations from the mean centrality of the nodes in the graph.
+In Bambus 2~\cite{bambus} repeats were defined as nodes with a centrality score larger than three standard deviations from the mean centrality of the nodes in the graph.
 This strategy implicitly assumes the distribution of centrality values in the graph to follow a normal distribution.

### 2015-05-19 manuscript/manuscript.tex
added changes

@@
 
-Brandes\cite{brandes} gives an exact algorithm for computing betweenness centrality of all the nodes that is based on solving a single-source shortest path (SSSP) problem for each node.
-A SSSP computation from $s$ produces a directed acyclic graph (DAG) encoding all shortest paths starting at $s$.
-The contributions of these paths to the betweenness counters can be computed in linear time using backward aggregation.
-Brandes' algorithm calculates $\sigma_{st}$  simultaneously with the shortest path calculations: $\sigma_{ss} = 1$ and for $s \neq t, \sigma_{st} = \sum_{v \in pred(t)} \sigma_{sv}$ where $pred(t)$ is a collection of the immediate predecessors of  $t$ on every shortes path ending at $t$.
+Brandes\cite{brandes} proposed an exact algorithm for computing betweenness centrality of all the nodes based on single source shortest path approach.
+A SSSP computation from $s$ yields a directed acyclic graph (DAG) representing all shortest paths starting at $s$.
+The contributions of these paths to the betweenness counters is computed in linear time using backward aggregation.
+Brandes' algorithm calculates $\sigma_{st}$  simultaneously with the shortest path calculations: $\sigma_{ss} = 1$ and for $s \neq t, \sigma_{st} = \sum_{v \in pred(t)} \sigma_{sv}$ where $pred(t)$ is a set of the immediate predecessors of  $t$ on every shortest path ending at $t$.
 In a subsequent aggregation phase, nodes are processed by non-decreasing distance from $s$.
